---
title: "Arousal and Valence analysis"
author: "LC"
date: "2024-11-28"
output:
  html_document:
    self_contained: true
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
    theme: cerulean 
---

```{r, message=FALSE}
library(tidyverse)
library(ggstatsplot)

bd <- "/data00/leonardo/RSA/analyses/RATINGS"

subs_file <- "/data00/leonardo/RSA/sub_list.txt"
subs <- sprintf("%02d",readLines(subs_file) %>% as.numeric)
```


# Load data
```{r, message=FALSE}
# IMPORTANT: REMOVE NEUTRAL !!

df_A <- read_csv(paste0(bd,"/","arousal_ratings.csv")) %>% 
  filter(emotion != "neutral") %>% 
  select(sub,video,r_arousal,emotion)


# Subtract 5 for valence only
df_V <- read_csv(paste0(bd,"/","valence_ratings.csv")) %>%
  filter(emotion != "neutral") %>% 
  select(sub,video,r_valence,emotion) %>% 
  mutate(r_valence = r_valence - 5)


df_AV <- inner_join(df_A, df_V, by = c("sub","video","emotion")) %>% 
  select(sub, r_arousal, r_valence,emotion)
```






# Arousal 
NB: there are so many sig pairwise comparisons, that it is actually cleaner to display only those
which are __not__ significant
```{r, warning=FALSE}
# Arousal
df_A %>%
  select(-video) %>% 
  group_by(sub,emotion) %>% 
  reframe(
    mean_arousal = mean(r_arousal)
  ) %>% 
  mutate(emotion = fct_reorder(emotion, mean_arousal)) %>%
  ggstatsplot::ggwithinstats(
    title = "Arousal ratings",
    x = emotion,
    y = mean_arousal,
    type = "parametric",
    bf.message = TRUE,
    p.adjust.method = "fdr",
    pairwise.display = "ns" # "s", "ns" or "none"
  )
```









# Valence
Again we display only __non__significant pairwise comparisons.

```{r, warning=FALSE}
# Valence
df_V %>%
  select(-video) %>% 
  group_by(sub,emotion) %>% 
  reframe(
    mean_valence = mean(r_valence)
  ) %>% 
  mutate(emotion = fct_reorder(emotion, mean_valence)) %>%
  ggstatsplot::ggwithinstats(
      title = "Valence ratings",
      x = emotion,
      y = mean_valence,
      type = "parametric",
      p.adjust.method = "fdr",
      pairwise.display = "ns" # "s", "ns" or "none"
    )

```

# Very high top values of valence. Is it a single subject?

Yes, it turns out that sub 15 (which is among the subs with highest congruence) gave
on average much higher valence

```{r}

df_V %>%
  select(-video) %>% 
  group_by(sub,emotion) %>% 
  reframe(
    mean_valence = mean(r_valence)
  ) %>% 
  ungroup %>% 
  group_by(emotion) %>% 
  arrange(desc(mean_valence)) %>% 
  group_split()
  
```

This is how the violin plot looks like if we exclude sub 15

```{r}
# Valence
df_V %>%
  filter(sub != "15") %>% 
  select(-video) %>% 
  group_by(sub,emotion) %>% 
  reframe(
    mean_valence = mean(r_valence)
  ) %>% 
  mutate(emotion = fct_reorder(emotion, mean_valence)) %>%
  ggstatsplot::ggwithinstats(
      title = "Valence ratings (w/out sub15)",
      x = emotion,
      y = mean_valence,
      type = "parametric",
      p.adjust.method = "fdr",
      pairwise.display = "ns" # "s", "ns" or "none"
    )
```

# Arousal and valence for high and low intensity movies

# Arousal
```{r, message=FALSE}

df_A <- read_csv(paste0(bd,"/","arousal_ratings.csv")) %>% 
  filter(emotion != "neutral") %>% 
  mutate(high_low_code = str_match(high_low_code, "high|low")) %>% 
  select(sub, r_arousal, emotion, high_low_code) %>% 
  group_by(sub, emotion, high_low_code) %>% 
  reframe(
    r_arousal = mean(r_arousal)
  ) %>% 
  ungroup %>% 
  group_by(emotion, high_low_code) %>% 
  reframe(
    mean_r_arousal = round(mean(r_arousal),2),
    sem = round(sd(r_arousal) / sqrt(n()),2)
  ) 

df_A %>% 
  ggplot(aes(x = emotion, y = mean_r_arousal, fill = high_low_code)) +
  geom_bar(stat = "identity", position = position_dodge(), color = "black") +
  geom_errorbar(aes(
    ymin = mean_r_arousal - sem, 
    ymax = mean_r_arousal + sem
  ), 
  position = position_dodge(0.9),  # Ensures alignment with bars
  width = 0.2                     # Controls the width of the error bars
  ) +
  labs(
    title = "Arousal Ratings by Emotion and High/Low Code",
    x = "Emotion",
    y = "Mean Arousal (±SEM)",
    fill = "High/Low Code"
  ) +
  theme_minimal()


```

# Valence
```{r, message=FALSE}

df_A <- read_csv(paste0(bd,"/","valence_ratings.csv")) %>% 
  filter(emotion != "neutral") %>% 
  mutate(r_valence = r_valence - 5) %>%   # REMOVE 5 FROM VALENCE
  mutate(high_low_code = str_match(high_low_code, "high|low")) %>% 
  select(sub, r_valence, emotion, high_low_code) %>% 
  group_by(sub, emotion, high_low_code) %>% 
  reframe(
    r_valence = mean(r_valence)
  ) %>% 
  ungroup %>% 
  group_by(emotion, high_low_code) %>% 
  reframe(
    mean_r_valence = round(mean(r_valence),2),
    sem = round(sd(r_valence) / sqrt(n()),2)
  ) 

df_A %>% 
  ggplot(aes(x = emotion, y = mean_r_valence, fill = high_low_code)) +
  geom_bar(stat = "identity", position = position_dodge(), color = "black") +
  geom_errorbar(aes(
    ymin = mean_r_valence - sem, 
    ymax = mean_r_valence + sem
  ), 
  position = position_dodge(0.9),  # Ensures alignment with bars
  width = 0.2                     # Controls the width of the error bars
  ) +
  labs(
    title = "Valence Ratings by Emotion and High/Low Code",
    x = "Emotion",
    y = "Mean valence (±SEM)",
    fill = "High/Low Code"
  ) +
  theme_minimal()


```











