---
title: "Emotion Ratings Analysis V4"
author: "Leonardo Cerliani"
date: "2024-11-27"
output:
  html_document:
    self_contained: true
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
    theme: cerulean 
---

# Version notes

Now we estimate congruence only with the LOO correlation method between each subject's ratings for each movie (vector of length 6) and the mean of all other subs.

Once obtained the congruence (r) scores, we detect outliers in the following way:

- Transfom all the r values to Z scores (Fisher) and average them per sub across movies. This will yield 26 Z scores. At this point, detect the outlier(s) as those who lie > 2 std from the mean.

- Repeat this with R^2 scores. This the conversion to Z scores will be achieved with the `scale` function on the R^2 scores.

- Ascertain the quality of being an outlier with boxplot (1.5 IQR)

Also, congruence will be plot as R^2

NB: A lot of previous stuff is not present anymore - e.g. the initial calculation
of congruence using % of correctly detected emotion in movies. These can be found
in previous V1..V3 of this notebook.

```{r, message=FALSE}
library(tidyverse)
library(RColorBrewer)

bd <- "/data00/leonardo/RSA/analyses/RATINGS"

subs_file <- "/data00/leonardo/RSA/sub_list.txt"
subs <- sprintf("%02d",readLines(subs_file) %>% as.numeric)

# Load the preprocessed emotion ratings file
# which was prepared with the 02_prepare_ratings_EMOTION.Rmd
# located in /data00/leonardo/RSA/prep_scripts/05_prepare_ratings
df <- read_csv(paste0(bd,"/","emotion_ratings.csv"))

```


# Examine congruence

We need to do the following for each movie, but let's explain the procedure
for one movie only since it's simpler:

- given one movie ratings (vector of 6), for each subject take the mean of the ratings of all other subs (also a vector of 6)
- then correlate the ratings of that sub with the mean of all other sub for that movie
- repeat for each sub
- now apply to each movie

In the end, you will have 56 values (one per movie) for each sub.
Each value shows the congruence (expressed as correlation) of the ratings of that
sub for a specific movie, with respect to the mean rating of all other subs for
that movie.


## Building and testing the fn on one sub across movies
```{r}
# Let's start with only one movie
df_one_movie <- df %>% 
  select(sub, video, starts_with("r_"), emotion) %>% 
  filter(video == "AK_A_32.mp4")

df_one_movie

# Get the name of the columns with ratings
rating_cols <- df_one_movie %>% select(starts_with("r_")) %>% colnames

# Function to calculate correlation for one subject
calculate_correlation <- function(sub_id, data, rating_cols) {
  # Separate the target subject
  target <- data %>% filter(sub == sub_id) %>% select(starts_with("r_"))
  
  # Calculate the mean ratings excluding the target subject
  others_mean <- data %>%
    filter(sub != sub_id) %>%
    summarise(across(all_of(rating_cols), mean, na.rm = TRUE))
  
  # Calculate the correlation
  corr <- cor(
    unlist(target),
    unlist(others_mean),
    use = "complete.obs"
  )
  
  # Return as a tibble
  tibble(
    sub = sub_id,
    # movie = unique(data$video),
    corr = corr
  )
}


# # Try the function with one sub (and one movie)
# sub_id = "03"
# data = df_one_movie
# calculate_correlation(sub_id = "03", data = df_one_movie, rating_cols)


# Map the function to all subjects
unique(df$sub) %>% map_dfr( ~ calculate_correlation(.x, df_one_movie, rating_cols))


```


## Carry out the calculation for each sub across movies

```{r}
# Now use the full dataset of ratings with all movies.
# We will therefore need to write two nested loops
#
# NB: in the loop below, x = nested_df for each row of 'data'

congruence <- df %>%
  
  # Remove neutral first!
  filter(emotion != "neutral") %>%
  
  select(sub, video, starts_with("r_")) %>%
  group_by(video) %>%
  nest() %>%
  mutate(
    correlations = map(data, function(x) {
      # Extract the column names for ratings
      rating_cols <- x %>% select(starts_with("r_")) %>% colnames()
      
      # Iterate over the subjects in the nested data and calculate correlation
      map_dfr(x$sub, function(sub_id) {
        calculate_correlation(sub_id, x, rating_cols)
      })
    })
  ) %>%
  select(video, correlations) %>%
  unnest(correlations) %>% 
  ungroup 
  
  # NB: in some cases the rating was all zeros, thereby leading to no sd and
  # NA in the corr calculation. Replace these with zeros
  congruence <- congruence %>% 
    mutate(corr = ifelse(is.na(corr), 0, corr))
  
  # congruence is the column named corr
  congruence
```








# Visualization of congruence (R)

```{r}
# Calculate the median correlation for each subject
median_corrs <- congruence %>%
  group_by(sub) %>%
  summarise(median_corr = median(corr, na.rm = TRUE)) %>%
  arrange(median_corr)

# Create the plot
congruence %>%
  ggplot(aes(x = sub, y = corr)) +
  geom_boxplot(outlier.shape = NA) +  # Boxplot for each subject
  geom_jitter(width = 0.2, alpha = 0.6) +  # Dots inside the boxplot
  scale_x_discrete(limits = median_corrs$sub) +  # Order subjects by their median correlation
  geom_line(data = median_corrs, aes(x = sub, y = median_corr, group = 1), color = "red", linewidth = 0.8) +  # Corrected geom_line
  labs(
    title = "Congruence of each Subject Ratings with Mean across Subjects (R LOO)",
    subtitle = ("Each dot represents the correlation of that subject's emotion ratings (vector of 6) 
for a given movie with the mean ratings of all other subjects for the same movie"),
    x = "Subject ID",
    y = "Correlation"
  ) +
  theme_minimal() +  # Use a minimal theme for a cleaner look
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for better readability
    plot.title = element_text(hjust = 0.5)  # Center the title
  )

```

## Same plot but with R^2
```{r}
congruence %>%
  mutate(corr = corr^2) %>% 
  ggplot(aes(x = sub, y = corr)) +
  geom_boxplot(outlier.shape = NA) +  # Boxplot for each subject
  geom_jitter(width = 0.2, alpha = 0.6) +  # Dots inside the boxplot
  scale_x_discrete(limits = median_corrs$sub) +  # Order subjects by their median correlation
  geom_line(data = median_corrs, aes(x = sub, y = median_corr, group = 1), color = "red", linewidth = 0.8) +  # Corrected geom_line
  labs(
    title = "Congruence of each Subject Ratings with Mean across Subjects (LOO R^2)",
    subtitle = ("Each dot represents the correlation of that subject's emotion ratings (vector of 6) 
for a given movie with the mean ratings of all other subjects for the same movie"),
    x = "Subject ID",
    y = "Correlation"
  ) +
  theme_minimal() +  # Use a minimal theme for a cleaner look
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for better readability
    plot.title = element_text(hjust = 0.5)  # Center the title
  )
```




## Detecting Outliers

We detect as outliers those subjects whose congruence score (either R or R^2) 
across movies are < 2 std in the distribution of the mean congruence score 
per subject.

```{r}

# 1. r-to-Z Fisher transform of the congruence scores (R)
# Look for z_scaled values < 2
congruence %>% 
  mutate(fisher_z = 0.5 * log((1 + corr) / (1 - corr))) %>%
  select(sub, fisher_z) %>% 
  group_by(sub) %>% 
  reframe(
    mean_z = mean(fisher_z)
  ) %>% 
  mutate(z_scaled = scale(mean_z)) %>% 
  arrange(z_scaled)

# 2. Z-scoring the R^2 values 
# Look for z_scaled values < 2
mean_r2 <- congruence %>%
  mutate(r2 = corr^2) %>% 
  group_by(sub) %>% 
  reframe(
    mean_r2 = mean(r2)
  ) %>% 
  mutate(z_scaled = scale(mean_r2)) %>% 
  arrange(z_scaled)

mean_r2


# 3. sub(s) <> 1.5 IQR (efficiently with boxplot)
# boxplot(mean_r2$z_scaled)

is_outlier <- function(x) {
  return(x < quantile(x, 0.25) - 1.5 * IQR(x) | x > quantile(x, 0.75) + 1.5 * IQR(x))
}

mean_r2 %>% 
  mutate(outlier = ifelse(is_outlier(z_scaled), z_scaled, as.numeric(NA))) %>% 
  ggplot(aes(y = 1, x = z_scaled)) +  # Switch x and y aesthetics
  geom_boxplot() +
  geom_point() +
  geom_text(aes(label = sub), na.rm = TRUE, vjust = -0.5) +  # Use vjust for vertical adjustment
  theme_minimal() +
  labs(y = "Group", x = "z_scaled")  # Update axis labels


```






























