---
title: "Prepare Arousal and Valence for JASP"
author: "LC"
date: "2024-11-27"
output: html_document
---


NB: since the analysis is carried out in JASP, it can be found at [GDrive]/RSA/arousal_valence_anova2


# Load the rating files
```{r, message=FALSE}
library(tidyverse)

bd <- "/data00/leonardo/RSA/analyses/RATINGS"

subs_file <- "/data00/leonardo/RSA/sub_list.txt"
subs <- sprintf("%02d",readLines(subs_file) %>% as.numeric)

# Load the preprocessed ratings file
# which was prepared with the 02_prepare_ratings_EMOTION.Rmd
# located in /data00/leonardo/RSA/prep_scripts/05_prepare_ratings

# In both cases, REMOVE THE NEUTRALS!

df_arousal <- read_csv(paste0(bd,"/","arousal_ratings.csv")) %>% 
  filter(emotion != "neutral")

# Remove 5 from valence ratings (they were actually from -5 to 5)
df_valence <- read_csv(paste0(bd,"/","valence_ratings.csv")) %>%
  filter(emotion != "neutral") %>% 
  mutate(r_valence = r_valence - 5)

```


# Save in the format required by JASP

This is a 6x2 RM ANOVA, therefore we need 12 columns with nsub rows. E.g. Anger_A (Arousal), Anger_V (Valence), Pain_A, Pain_V and so on. 
Note that we have 8 ratings for each combination of subject/movie/rating (e.g. sub 02, anger, valence), so I will take the average rating of that subject for that movie for that rating type (arousal or valence).

```{r}

df_arousal <- df_arousal %>% select(sub, video, emotion, r_arousal)

df_valence <- df_valence %>% select(sub, video, emotion, r_valence)


df <- inner_join(df_arousal, df_valence, by = c("sub","video","emotion")) %>% 
  rename(arousal = r_arousal, valence = r_valence)


df_4_JASP <- df %>% 
  select(-video) %>% 
  
  # Take the mean rating for each sub/movie/rating type (A or V)
  group_by(sub,emotion) %>% 
  reframe(
    arousal = mean(arousal),
    valence = mean(valence)
  ) %>%
  
  # Put in the format required by JASP for RM ANOVA
  pivot_wider(names_from = "emotion", values_from = c("arousal","valence"))

df_4_JASP
  
write_location <- paste0(bd, "/ANALYSES_OF_RATINGS")
write_filename <- "arousal_valence_4_JASP.csv"

write_csv(df_4_JASP, paste0(write_location, "/", write_filename))

```

