---
title: "RSA viz dev"
author: "LC"
date: "2024-05-15"
output: html_document
---

# Motivation
Here we develop the logic for post-hoc viz of RSA results.
By viz we mean the MDS and reordered matrices for the fmri RDMs
which are commonly found in Kriegekorte papers.

Specifically, the code below is meant to be used in the following way:

- Carry out the RSA on the desired mask
- Identify regions for which you would like to have the viz
- Define an ROI - e.g. in fslview - and save it
- Feed the ROI in the code below

Basically, this code carries out the first part of the ROI-based RSA, that is, 
the calculation of the fmri RDMs.

Once we have the RDMs for each sub for ROI, we take the average across subs
and then we replot the MDS/UMAP and reordered matrix. 
The mask is also used to extract the RSA values from the saved NIFTI - in the main
RSA analysis - and plot the corresponding barplots

This means that the code below must contain functions to call the fmri RDMs
construction mutliple times (for emotion, arousal, valence)


```{r load_libraries, message=F}
library(tidyverse)
library(future)
library(furrr)
library(tictoc)
library(RNifti)
library(proxy) # distances
library(profvis)
library(DT)
library(formattable)

source("funs_V8_ROIs.R")

bd="/data00/leonardo/RSA/analyses"

# can be emotion, arousal, valence
ratings_type <- "emotion"

# the following is our contrast of interest and it's always the same
copes_type <- "one_ev_per_movie"

# Choose the distance metric to use for fmri RDM
# Supported methods: 'pearson', 'spearman', 'euclidean', 'cosine' or 'mahalanobis'
dist_method_rating <- "euclidean"
dist_method_fmri = "euclidean"
dist_method_rsa = "pearson"

RDM_type = "tril"     

# Vector of zeropadded sub_ids
subs_file <- "/data00/leonardo/RSA/sub_list.txt"
subs <- sprintf("%02d",readLines(subs_file) %>% as.numeric)

# ---------- ONLY TOP RATERS BELOW ------------
subs <- c("02","03","12","11","22","09","29","28","26","32","23","15","20","19")

```


# Aux functions
```{r}
# plot_heatmap(D %>% as.matrix)
plot_heatmap <- function(M) {
  heatmap(M, Rowv = NA, Colv = NA, symm=T, revC = T)
}
```


# Create a df_path_copes with the location of the 56 cope niis from the one_movie_per_ev model
Extract the pathname of all copes using the `list.files()` function.
Also define a copes_numba vector with all the copes numbers.

NB: The cope numbers in the `cope` column are NOT zeropadded since this is how they come out from FSL Feat
```{r, message=FALSE}

copes_location_csv <- paste0(bd,"/rsa/copes_location.csv")

df_path_copes <- if (file.exists(copes_location_csv) ) {
  df_path_copes <- read_csv(copes_location_csv)
} else {
  df_path_copes <- import_df_path_copes(bd, copes_type)
}

copes_numba <- df_path_copes$cope %>% unique
ncopes <- length(copes_numba)

```



# Read in the ROI(s) for which you want to calculate RDMs
NB: there can be multiple ROIs in one nii

```{r}

bd_ROI <- "/data00/leonardo/RSA/analyses/rsa/rsa_viz/"
ROI_filename <- "roi_099.nii.gz"

ROI_nii <- readNifti(paste0(bd_ROI, "/", ROI_filename))
# view(ROI_nii)

region_labels <- ROI_nii[ROI_nii > 0] %>% unique %>% sort

```


# Calculate RDMs
```{r}

# 5 workers give the best performance
plan(multisession, workers = 5)

tic()
RDMs_fmri <- subs %>% future_map_dfr(~{
  
  paste0("Calculating RDMs for sub ",.x,"\n") %>% cat
  
  df_copes <- load_sub_copes(.x, copes_numba, df_path_copes)
  
  one_RDM_fmri <- calculate_fmri_RDMs(
    df_copes,
    ROI_nii,
    dist_method_fmri,
    RDM_type = RDM_type
  )

  one_RDM_fmri %>% mutate(sub = .x) %>% relocate(sub)
  
})
toc()


plan(sequential)
```




```{r}

# Function to reconstruct full similarity matrix
reconstruct_similarity_matrix <- function(lower_triangular_data) {
  n <- ncopes
  full_matrix <- matrix(0, nrow = n, ncol = n)
  full_matrix[lower.tri(full_matrix)] <- lower_triangular_data
  full_matrix <- t(full_matrix) + full_matrix
  diag(full_matrix) <- 0  # Set diagonal to 1 if needed
  return(full_matrix)
}


RDMS_FULL <- RDMs_fmri %>% 
  group_by(sub) %>% 
  nest %>%
  mutate(D = map(data, ~ reconstruct_similarity_matrix( unlist(.x) )))


# Step 1: Extract the full_matrix column
full_matrices <- RDMS_FULL$D

# Step 2: Sum up all the matrices
summed_matrix <- reduce(full_matrices, `+`)

# Step 3: Divide by the number of rows to get the average
average_matrix <- summed_matrix / nrow(RDMS_FULL)

plot_heatmap(average_matrix)

```



```{r}

nROI <- 1

RDMs_fmri %>% 
  select(sub, paste0("RDM_region_", nROI)) %>% 
  rename_with(~ "ROI_tril", starts_with("RDM_region_")) %>% 
  group_by(sub) %>% 
  summarise(
    mean_tril = mean(ROI_tril),
    sd_tril = sd(ROI_tril)
  )


# Function to reconstruct full similarity matrix
reconstruct_similarity_matrix <- function(tril,ncopes) {
  n <- ncopes
  full_matrix <- matrix(0, nrow = n, ncol = n)
  full_matrix[lower.tri(full_matrix)] <- tril
  full_matrix <- t(full_matrix) + full_matrix
  diag(full_matrix) <- 0  # Set diagonal to 1 if needed
  return(full_matrix)
}

reconstruct_similarity_matrix()


RDMs_fmri %>% 
  group_by(sub) %>% 
  mutate(row_numba = row_number()) %>% 
  ungroup %>% 
  group_by(row_numba) %>% 
  mutate(mean = mean(RDM_region_1))
  

```


























