---
title: "RSA ROI analysis"
subtitle: "rating and fmri RDMs for a given roi.nii.gz"
author: "Leonardo Cerliani"
date: "23/6/2023"
output:
  html_document:
    self_contained: true
    code_folding: hide
    highlight: pygments
    theme: cerulean
    css: styles.css
---


# Initialization

## Load libraries
```{r message=F}
library(tidyverse)
library(future)
library(furrr)
library(tictoc)
library(RNifti)
library(proxy) # distances
library(profvis)
library(DT)
library(formattable)

source("funs_V8_ROIs.R")

bd="/data00/leonardo/RSA/analyses"

bd_ratings = paste0(bd,"/RATINGS")

# Choose the distance metric to use for fmri RDM
# Supported methods: 'pearson', 'spearman', 'euclidean', 'cosine' or 'mahalanobis'
dist_method_rating <- "euclidean"
dist_method_fmri = "euclidean"
dist_method_rsa = "pearson"
```


## Read in the ROI(s) for which you want to calculate RDMs
NB: there can be multiple ROIs in one nii

```{r}

bd_ROI <- "/data00/leonardo/RSA/analyses/rsa/rsa_viz/"
ROI_filename <- "ROI.nii.gz"


ROI_nii <- readNifti(paste0(bd_ROI, "/", ROI_filename))
# view(ROI_nii)

region_labels <- ROI_nii[ROI_nii > 0] %>% unique %>% sort

```


## Define subs set
```{r message=F}

# Vector of zeropadded sub_ids
subs_file <- "/data00/leonardo/RSA/sub_list.txt"
subs <- sprintf("%02d",readLines(subs_file) %>% as.numeric)

# ---------- ONLY TOP RATERS BELOW ------------
subs <- c("02","03","12","11","22","09","29","28","26","32","23","15","20","19")

# load emotion labels
emotion_labels <- read_csv(paste0(bd,"/RATINGS/emotion_ratings.csv")) %>% 
  filter(sub == "02") %>% select(emotion) %>% pull

```


## Aux functions
```{r}
# plot_heatmap(D %>% as.matrix)
plot_heatmap <- function(M, title=NULL) {
  heatmap(M, Rowv = NA, Colv = NA, symm=T, revC = T, main=title)
}

# The following is to plot also side colors for emotion
# NB:   # The input matrix MUST already have rownames for emotions
plot_heatmap_sides <- function(M, title=NULL) {
  
  emo_names <- rownames(M)
  
  # Define a blue to red color palette for the heatmap
  blue_to_red <- colorRampPalette(c("blue", "white", "red"))(n = 100)
  
  # Define the palette for emotions
  paletta_emotions <- c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3", "#FF7F00", "#FFFF33", "#FFFFFF")
  emo_labels <-       c("anger",   "pain",    "disgust", "sad",     "fear",    "happy",   "neutral")
  
  # Map each unique emotion to a color
  label_colors <- setNames(paletta_emotions, emo_labels)
  
  # Create vectors of colors for row and column side colors based on emo_labels_reord
  row_side_colors <- col_side_colors <- label_colors[emo_names]
  
  # Plot it
  heatmap(M, Rowv = NA, Colv = NA, symm=T, revC = T, col = blue_to_red,
          RowSideColors = row_side_colors, ColSideColors = col_side_colors, 
          main = title)
}


```


# Rats RDM

## Function to calculate model-specific mean_rats RDM matrix

**NB**: the reordering should be provided in the output in order to 
differentiate emotion, arousal and valence model

```{r, message=FALSE}

get_rats_RDM <- function(ratings_type, dist_method_rating) {

  bd_ratings = paste0(bd,"/RATINGS")
  ratings_path <- paste0(bd_ratings,"/",ratings_type,"_ratings.csv")
  rats <- read_csv(ratings_path)
  
  # mean rats
  rats_mean <- rats %>% 
    select(sub, emotion, video, starts_with("r_")) %>% 
    filter(sub %in% subs) %>% 
    group_by(video) %>% 
    summarise(
      emotion = max(emotion),
      across(starts_with("r_"), ~ round(mean(.x, na.rm = TRUE), 2))
    )
  
  # get the reordering vector for increasing arousal
  idx_sort <- switch(
    ratings_type,
    "emotion" = order(rats_mean$emotion),
    "arousal" = order(rats_mean$r_arousal),
    "valence" = order(rats_mean$r_valence)
  )    
  
  # calculate rats RDM
  rats_RDM <- DDOS(
    rats_mean %>% select(starts_with("r_")),
    method = dist_method_rating
  ) %>% as.matrix()
  
  rownames(rats_RDM) <- rats_mean$emotion
  colnames(rats_RDM) <- rats_mean$emotion
  
  return(
    list(idx_sort, rats_mean, rats_RDM)
  )

}
```


## Get the reordered rats RDM for all three models
The idx_sort_[model] will be used to reorder the fmri_RDM
```{r, message=FALSE}

# AROUSAL
out_arousal <- get_rats_RDM("arousal", dist_method_rating)
idx_sort_arousal <- out_arousal[[1]]
rats_mean_arousal <- out_arousal[[2]]
RDM_arousal <- out_arousal[[3]]


# EMOTION
out_emotion <- get_rats_RDM("emotion", dist_method_rating)
idx_sort_emotion <- out_emotion[[1]]
rats_mean_emotion <- out_emotion[[2]][idx_sort_emotion,]
RDM_emotion <- out_emotion[[3]]


# VALENCE
out_valence <- get_rats_RDM("valence", dist_method_rating)
idx_sort_valence <- out_valence[[1]]
rats_mean_valence <- out_valence[[2]]
RDM_valence <- out_valence[[3]]


# VALENCE
out_aroval <- get_rats_RDM("aroval", dist_method_rating)
idx_sort_aroval <- out_aroval[[1]]
rats_mean_aroval <- out_aroval[[2]]
RDM_aroval <- out_aroval[[3]]


plot_heatmap_sides(RDM_emotion[idx_sort_emotion,idx_sort_emotion], "emotion")
# plot_heatmap_sides(RDM_arousal[idx_sort_arousal,idx_sort_arousal], "arousal")
# plot_heatmap_sides(RDM_valence[idx_sort_valence,idx_sort_valence], "valence")
# plot_heatmap_sides(RDM_aroval[idx_sort_valence,idx_sort_valence], "aroval")


# all reordered by emotion
plot_heatmap_sides(RDM_arousal[idx_sort_emotion,idx_sort_emotion], "arousal")
plot_heatmap_sides(RDM_valence[idx_sort_emotion,idx_sort_emotion], "valence")
plot_heatmap_sides(RDM_aroval[idx_sort_emotion,idx_sort_emotion], "aroval")


```




```{r, eval=FALSE, include=FALSE}
# RDM_ratings dimred

library(umap)
library(patchwork)

umap_result <- umap::umap(RDM_valence, input="dist", n_components=5)

paletta_emotions <- c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3", "#FF7F00", "#FFFF33", "#FFFFFF")

# plot pc1,2 and pc2,3
do_UMAP_lowdim_plots <- function(umap_result, lowdim_x, lowdim_y) {
  
  umap_lowdim <- data.frame(
    pc_x = umap_result$layout[,lowdim_x], 
    pc_y = umap_result$layout[,lowdim_y], 
    emotion_labels
  )
  
  p <- ggplot(umap_lowdim, aes(x = pc_x, y = pc_y, color = emotion_labels)) +
    geom_point(size = 3, alpha = 0.7) +  # Open circles for points
    scale_color_manual(values = paletta_emotions) +
    scale_shape_manual(values = c(1, 3)) +  # Assign specific shapes
    labs(
      title = "UMAP Projection", 
      x = paste0("lowdim ",lowdim_x), 
      y = paste0("lowdim ",lowdim_y)
    ) +
    theme_minimal() +
    guides(color = guide_legend(override.aes = list(shape = 16)))  # Filled circles in the legend
  
  return(p)
}


p <- list(c(1,2),c(2,3)) %>% pmap(~ do_UMAP_lowdim_plots(umap_result, lowdim_x = .x, lowdim_y = .y))

p[[1]] + p[[2]]
```






# fmri RDM

## Get copes location
NB: The cope numbers in the `cope` column are NOT zeropadded since this is how they come out from FSL Feat

```{r,message=FALSE}

# the following is our contrast of interest and it's always the same
copes_type <- "one_ev_per_movie"

copes_location_csv <- paste0(bd,"/rsa/copes_location.csv")

df_path_copes <- if (file.exists(copes_location_csv) ) {
  df_path_copes <- read_csv(copes_location_csv)
} else {
  df_path_copes <- import_df_path_copes(bd, copes_type)
}

copes_numba <- df_path_copes$cope %>% unique
ncopes <- length(copes_numba)

```



## Calculate RDMs
```{r}

# 5 workers give the best performance
plan(multisession, workers = 5)

tic()
RDMs_fmri <- subs %>% future_map_dfr(~{
  
  paste0("Calculating RDMs for sub ",.x,"\n") %>% cat
  
  df_copes <- load_sub_copes(.x, copes_numba, df_path_copes)
  
  one_RDM_fmri <- calculate_fmri_RDMs(
    df_copes,
    ROI_nii,
    dist_method_fmri,
    RDM_type = "tril"
  )

  one_RDM_fmri %>% mutate(sub = .x) %>% relocate(sub)
  
})
toc()


plan(sequential)

```



## Calculate mean triu and plot full
```{r}

mean_triu <- RDMs_fmri %>% 
  group_by(sub) %>% 
  mutate(triu_cell = row_number()) %>% 
  ungroup() %>% 
  group_by(triu_cell) %>% 
  summarise(mean_triu_cell = mean(RDM_region_1)) %>% 
  select(mean_triu_cell) %>% pull

# mean_triu
hist(mean_triu)


# reconstruct full matrix and assign row/colnames
reconstruct_full <- function(triu, ncopes, emotion_labels) {
  RDM_full <- matrix(0, nrow = ncopes, ncol = ncopes)
  RDM_full[lower.tri(RDM_full, diag = F)] = triu
  RDM_full = RDM_full + t(RDM_full)
  diag(RDM_full) <- NA
  rownames(RDM_full) <- colnames(RDM_full) <- emotion_labels
  return(RDM_full)
}

RDM_full <- reconstruct_full(mean_triu, ncopes, emotion_labels)

# # plot only reordered by emotion for consistency with the other rsa flavours 
# plot_heatmap_sides(RDM_full[idx_sort_arousal, idx_sort_arousal], "arousal")
plot_heatmap_sides(RDM_full[idx_sort_emotion, idx_sort_emotion], "emotion")
# plot_heatmap_sides(RDM_full[idx_sort_valence, idx_sort_valence], "valence")

```


## UMAPs on RDM_full
```{r fig.fullwidth=TRUE}

p1 <- do_umap(RDM_full, "arousal", emotion_labels, rats_mean_arousal, rats_mean_valence)
p2 <- do_umap(RDM_full, "emotion", emotion_labels, rats_mean_arousal, rats_mean_valence)
p3 <- do_umap(RDM_full, "valence", emotion_labels, rats_mean_arousal, rats_mean_valence)

p1
p2
p3

```















































# Garbage collection

# Same plot but with percentiles

```{r, include=FALSE, eval=FALSE}
# Same plot but with percentiles (like in K2008 plots)

# estimate empirical cumulative distribution function
ecdf_func <- ecdf(mean_triu)
# convert each value to percentile
pctiles <- ecdf_func(mean_triu)

RDM_full <- reconstruct_full(pctiles, ncopes, emotion_labels)

plot_heatmap_sides(RDM_full[idx_sort_arousal, idx_sort_arousal], "arousal")
plot_heatmap_sides(RDM_full[idx_sort_emotion, idx_sort_emotion], "emotion")
plot_heatmap_sides(RDM_full[idx_sort_valence, idx_sort_valence], "valence")

```



## Rats and fmri RDM corr

```{r, include=FALSE, eval=FALSE}

vec_rats <- RDM_arousal[lower.tri(RDM_arousal, diag = FALSE)]

vec_fmri <- RDM_full[lower.tri(RDM_full, diag = FALSE)]

cor(vec_rats, vec_fmri)
plot(vec_rats, vec_fmri)

```



















