---
title: "do_RSA_v11"
author: "LC"
date: "2024-08-12"
output: html_document
---

# Update notes v11
- option to remove neutral
- updated load_sub_copes to select the path column
- switched from heatmap to ComplexHeatmap


# Load libraries and define parameters

NB: this Rmd can be used for all types of ratings: emotion, arousal, valence

Abbreviations:
- sub : subjects(s)
- rat : rating(s)

# Load libraries
```{r load_libraries, message=F}
library(tidyverse)
library(future)
library(furrr)
library(tictoc)
library(RNifti)
library(proxy) # distances
library(profvis)
library(DT)
library(formattable)
library(janitor)
library(ComplexHeatmap)
library(circlize)  # For color mapping functions
library(grid)  # For gpar()
# library(pheatmap)
# library(heatmaply)
library(papayaWidget)
library(ggstatsplot)


source("funs_V11_ROIs.R")
```



# Define paths and rsa_[flavour] specific option
```{r}

bd="/data00/leonardo/RSA/analyses"
bd_ratings = paste0(bd,"/RATINGS")
ratings_type <- c("emotion","arousal","valence","aroval")


rsa_flavour="rsa_ROI"

remove_neutrals = "YES"

# choose the type of betas to be used, e.g.
# - one_ev_per_movie 
# - one_ev_per_movie_minus_neutral
# - emotion_predictors
copes_type <- "one_ev_per_movie_minus_neutral"


# Choose one atlas, e.g.
# Schaefer100_2mm.nii.gz
# Schaefer200_2mm.nii.gz
# Yeo_7_liberal_2mm.nii.gz
atlas_filename <- "Yeo_17_liberal_2mm.nii.gz"
# Atlases repos:
# https://www.lead-dbs.org/helpsupport/knowledge-base/atlasesresources/cortical-atlas-parcellations-mni-space/
# https://github.com/neurodata/neuroparc
# https://www.fmrib.ox.ac.uk/datasets/brainmap+rsns/

# Choose the distance metric to use for fmri RDM
# The metric for ratings RDM *should be* euclidean, since arousal and valence
# ratings have only one value
# Supported methods: 'pearson', 'spearman', 'euclidean', 'cosine' or 'mahalanobis'
dist_method_rating <- "euclidean"
dist_method_fmri = "euclidean"
dist_method_rsa = "pearson"

# Vector of zeropadded sub_ids
subs_file <- "/data00/leonardo/RSA/sub_list.txt"
subs <- sprintf("%02d",readLines(subs_file) %>% as.numeric)

# ---------- ONLY TOP RATERS BELOW ------------
subs <- c("02","03","12","11","22","09","29","28","26","32","23","15","20","19")

```


# Create copes metadata (to identify and remove neutrals)
Before anything else, we need to identify the neutral copes and their associated 
ratings. 

To this aim we load the emotion ratings which contain also the correspondence
between movie, cope and rating. Then we select only one sub (the order is the
same for all subs)
```{r, message=FALSE}

metadata_copes <- read_csv(paste0(bd_ratings, "/emotion_ratings.csv")) %>% 
  filter(sub == subs[1]) %>% 
  select(high_low_code, movie_cope_number, emotion) %>% 
  rename(cope = movie_cope_number, label = high_low_code)
  
metadata_copes

```




# Create a df_path_copes with the location of the 56 cope niis from the one_movie_per_ev model
Extract the pathname of all copes using the `list.files()` function.
Also define a copes_numba vector with all the copes numbers.

NB: The cope numbers in the `cope` column are NOT zeropadded since this is how they come out from FSL Feat
```{r, message=FALSE}

# load df_path_copes and add metadata (label and emotion/neutral)
df_path_copes <- import_df_path_copes(bd, copes_type, rsa_flavour) %>% 
  inner_join(metadata_copes, by="cope")

# If remove_neutrals is "YES", remove neutral copes and arrange by sub/cope (just to be sure) 
if (remove_neutrals == "YES") {
  df_path_copes <- df_path_copes %>%  
  filter(emotion != "neutral") %>% 
  arrange(sub,cope)
} 

copes_numba <- df_path_copes$cope %>% unique

cat(
  "there are",length(copes_numba)," copes including ",
  nrow(df_path_copes %>% filter(sub==subs[1], emotion=="neutral"))," neutral copes \n"
)

# df_path_copes
# length(copes_numba)
```


# Read atlas and create a vector of atlas labels
```{r}

bd_atlases = paste0("/data00/leonardo/RSA/analyses/",rsa_flavour,"/atlases")

atlas_path <- paste0(bd_atlases,"/",atlas_filename)

atlas_nii <- readNifti(paste0(bd_atlases,"/",atlas_filename))
region_labels <- atlas_nii[atlas_nii > 0] %>% unique %>% sort

```



# Calculate RATINGS RDMs
```{r, message=FALSE}

# NB: ratings_type is now a *vector*, e.g.: 
# ratings_type <- c("emotion","arousal","valence","aroval")
# to test the fn below for one rating type : do_RDMs_rats("emotion")

do_RDMs_rats <- function(ratings_type, remove_neutrals) {

  ratings_path <- paste0(bd_ratings,"/",ratings_type,"_ratings.csv")
  rats <- read_csv(ratings_path)
  
  if (remove_neutrals == "YES") {
    rats <- rats %>% filter(emotion != "neutral")
  }
  
  rdm <- rats %>%
    filter(sub %in% subs) %>%
    select(sub, starts_with("r_")) %>%
    group_by(sub) %>%
    nest %>%
    mutate(!!paste0("RDM_",ratings_type) := data %>% map(~ DDOS_vec(.x, dist_method_rating))) %>%
    ungroup %>% 
    # remove the sub with the prospect of merging rdms of different ratings/models
    select(!data)
  
  return(rdm)
  
}


# calculate rdms for all ratings_type and return one column for each ratings_type
RDMs_rats <- map_dfc(ratings_type, ~ do_RDMs_rats(.x, remove_neutrals)) %>%
  # remove the duplicated sub_ columns (generated by do_RDMs_rats) 
  janitor::clean_names() %>% 
  mutate(sub = sub_1) %>% 
  select(sub, starts_with("rdm"))


# # example heatmap
# plot_tril(
#   RDMs_rats[1,]$rdm_emotion, model = "emotion",
#   fontsize=7, side_mm=100, reord = TRUE
# )


```



# Checkpoint - control same order in ratings and fmri copes
Check that the ordering in the `do_RDMs_rats` and in `df_path_copes` is the same.
```{r, message=FALSE, eval=FALSE}

df_path_copes4consistency <- df_path_copes %>% 
  filter(sub == sub[1]) %>% 
  select(label, cope, emotion)


rats4consistency <- read_csv(paste0(bd_ratings,"/emotion_ratings.csv")) %>% 
  filter(sub == subs[1]) %>% 
  select(high_low_code, movie_cope_number, emotion) %>% 
  rename(label = high_low_code, cope = movie_cope_number)


if(remove_neutrals == "YES") {
  rats4consistency <- rats4consistency %>% filter(emotion != "neutral")
} 

cbind(df_path_copes4consistency, rats4consistency) %>% 
  janitor::clean_names() %>% 
  relocate(colnames(.) %>% sort)


```




# Calculate FMRI RDMs
**IMPORTANT : ** Since we already removed neutral from df_path_copes we don't 
need to filter anything anymore here.
```{r}

# Fn to calculate the rdm_fmri for one sub
# Test for one sub: do_RDM_fmri(subs[1], copes_numba, df_path_copes)

do_RDM_fmri <- function(sub_id, copes_numba, df_path_copes) {
  
  # load all the copes nii.gz into a df
  df_copes <- load_sub_copes(sub_id, copes_numba, df_path_copes)
  
  # the code below does the calculation of the tril of RDM_fmri for all rois for one sub
  
  one_sub_RDM_fmri <- tibble(
    sub = sub_id,
    roi = region_labels
  ) %>%
  
    # extract idx_roi for atlas voxels in that roi
    mutate(idx_roi = roi %>% map(~ which(atlas_nii == .x)) ) %>%
    
    # extract the df_copes for the voxels in that region (idx_roi)
    mutate(df_copes_region = idx_roi %>% map(~ df_copes[.x,]) ) %>%
    
    # calculate tril of RDM_roi (output as numeric vector)
    mutate(rdm_fmri = df_copes_region %>% map(~ DDOS_vec(t(.x), dist_method_fmri)) ) %>% 
    
    # select only sub, roi (numba) and RDM_roi
    select(sub, roi, rdm_fmri) 
  
  return(one_sub_RDM_fmri)
}



# Calculate RDMs_fmri for all subs
plan(multisession, workers = 5)

RDMs_fmri <- subs %>% future_map_dfr(
  ~ {
    paste0("Calculating RDMs for sub ",.x,"\n") %>% cat
    do_RDM_fmri(.x, copes_numba, df_path_copes)
  }
)

plan(sequential)

# RDMs_fmri

```





# Calculate RSA
```{r}
# The dist_method_rsa is set above.
# can be pearson, spearman
# dist_method_rsa <- "spearman"

# Join fmri and rats RDMs and create a copy of the rat RDM for each sub/roi
RSA <- right_join(RDMs_fmri, RDMs_rats, by = "sub") %>%
  group_by(sub) %>%
  
  # Dynamically create rsa_ columns for every ratings type
  mutate(across(
    all_of(paste0("rdm_", ratings_type)),
    ~ map2_dbl(.x, rdm_fmri, ~ cor(.x, .y, method = dist_method_rsa)),
    .names = "rsa_{.col}"
  )) %>%
  mutate(across(starts_with("rsa_"), round, 3)) %>%
  ungroup()


# Get the mean RSA for each ratings_type
RSA_mean <- RSA %>%
  select(roi, starts_with("rsa_")) %>%
  group_by(roi) %>%
  reframe(
    across(
      starts_with("rsa_"),
      list(mean = ~ round(mean(.x, na.rm = TRUE), 2))
    )
  ) %>% 
  rename_with(
    ~ str_replace(.x, "rsa_rdm_", ""),
    starts_with("rsa_")
  )

RSA_mean %>% datatable()
```



# Calculate mean tril for ratings and for fmri (for each roi in the fmri case)
```{r}
# Takes a df and a "column" containing a nested vector in each row
# and returns the element-wise mean of that vector across rows
get_elementwise_mean <- function(mydf, nested_col, fun=mean) {
  m <- matrix(
    mydf[[nested_col]] %>% unlist,
    nrow = nrow(mydf),
    ncol = length(mydf[[nested_col]][[1]]),
    byrow = TRUE
  )
  
  m_summary <- apply(m, 2, fun)
  return(m_summary)
}


# mean RDMs_rats
mean_RDMs_rats <- RDMs_rats %>% select(starts_with("rdm_")) %>% colnames %>%  
  map_dfc(~{
    mean_tril <- get_elementwise_mean(RDMs_rats, .x, fun = mean)
    tibble(!!.x := mean_tril)
  })



# mean RDMs_fmri
plan(multisession, workers = 5)

mean_RDMs_fmri <- RDMs_fmri$roi %>% 
  future_map_dfr(~{
    mean_tril <- get_elementwise_mean(RDMs_fmri %>% filter(roi == .x), "rdm_fmri")
    tibble(roi = .x, mean_tril = mean_tril)
  }) %>% 
  group_by(roi) %>% 
  nest

plan(sequential)
```


## Plot mean RDMs_rats
```{r}

# Plot mean RDMs_rats
ratings_type %>% 
  map(
    ~ plot_tril(
      mean_RDMs_rats[[paste0("rdm_",.x)]], model = .x,
      fontsize=7, side_mm=80, reord = TRUE
    )
  )

```

## Plot mean RDMs_fmri for all ROIs
```{r, warning=FALSE}

mean_RDMs_fmri

seq(nrow(mean_RDMs_fmri)) %>% 
  map(
    ~ plot_tril(
      mean_RDMs_fmri[.x,]$data, model = paste0("fmri_roi_",.x),
      fontsize=7, side_mm=80, reord = TRUE
    )
  )

```



# Elements for the app

## RSA table
```{r}
RSA_mean %>% datatable()
```


## Boxplot for one roi
```{r}

roi_numba = 6

RSA %>% 
  select(sub, roi, starts_with("rsa_")) %>% 
  filter(roi == roi_numba) %>% 
  select(starts_with("rsa_")) %>% 
  pivot_longer(cols = starts_with("rsa_"), names_to = "model") %>%
  mutate(model = str_replace(model, "rsa_rdm_","")) %>% 
  ggwithinstats(
    x = model, y = value,
    type = "p",   # p, np, r, b
    results.subtitle = TRUE
  )
  
```


## View one ROI
```{r}

atlas_filename <- "Yeo_7_liberal_2mm"

render_papaya <- function(atlas_filename, roi_numba) {
  
  f <- function(nii_filename) {
    return(paste0(bd_atlases,"/", nii_filename,".nii.gz"))
  }
  
  nii_atlas <- RNifti::readNifti( f(atlas_filename) )
  # view(nii_atlas)
  
  nii_atlas_thr <- ifelse(nii_atlas == roi_numba, 1, 0)
  RNifti::writeNifti(nii_atlas_thr, template = nii_atlas, paste0(bd_atlases,"/tmp_ROI.nii.gz"))
  
  papaya(
    c(f("Dummy"), f("MNI152_T1_2mm_brain"), f("tmp_ROI")),
    options = list(
      papayaOptions(alpha = 1, lut = "Grayscale"),
      papayaOptions(alpha = 0.5, lut = "Grayscale"),
      papayaOptions(alpha = 0.5, lut = "Red Overlay", min = 0, max = 5)
    ),
    interpolation = FALSE,
    orthogonal = FALSE
  )
  
}

render_papaya(atlas_filename, roi_numba = 6)

```



## Plot MEAN tril for fmri and all models for one roi
```{r, message=FALSE, warning=FALSE}

global_reord = FALSE
global_fontsize = 7
global_side_mm = 100

roi_numba = 6

# fmri one roi
tril_fmri <- plot_tril(
  mean_RDMs_fmri[roi_numba,]$data, model = paste0("fmri_roi",roi_numba),
  fontsize = global_fontsize, side_mm = global_side_mm, reord = global_reord
)


# emotion
tril_emotion <- plot_tril(
  mean_RDMs_rats$rdm_emotion, model = "emotion",
  fontsize = global_fontsize, side_mm = global_side_mm, reord = global_reord
)

# arousal
tril_arousal <- plot_tril(
  mean_RDMs_rats$rdm_arousal, model = "arousal",
  fontsize = global_fontsize, side_mm = global_side_mm, reord = global_reord
)

tril_fmri + tril_emotion


```




# TOBEDONE -- Write the map with the values -- TOBEDONE
```{r, eval=FALSE}

results_to_map <- RSA %>% 
  mutate(region = as.integer(str_replace(region, "RDM_region_",""))) %>% 
  select(region, mean_similarity)

results_to_map

# results_path <- ("/data00/leonardo/RSA/analyses/rsa/rsa_results")
# filename_results_nii <- paste0("rsa_results/rsa_results_",ratings_type,"_",atlas_filename)


results_path <- "/data00/leonardo/RSA/analyses/rsa/rsa_results"

results_filename <- create_filename_results(
  atlas_filename, ratings_type, dist_method_rating, dist_method_fmri, dist_method_rsa
)


# Function to write the results to a nii.gz volume
# (purrr gives issues here)
write_results_to_nii <- function(results_to_map, filename_results_nii, atlas_nii) {
  
  results_nii = atlas_nii
  
  for (ROI_numba in region_labels) {
    cat(paste0(ROI_numba," "))
    row_region <- which(results_to_map$region == ROI_numba)
    value <- results_to_map$mean_similarity[row_region]
    
    idx <- which(results_nii == ROI_numba)
    results_nii[idx] = value
  }

  writeNifti(results_nii, filename_results_nii)
}


# Write the results nii.gz in the rsa_results dir
write_results_to_nii(
  results_to_map, 
  paste0(results_path, "/", results_filename), 
  atlas_nii
)

```








