---
title: "prepare_model"
author: "LC"
date: "2024-05-17"
output: html_document
---

# Order of the participants in the 4D
To use randomise, we need to merge all the subs RSA volumes into a 4D. 
This is a very delicate procedure, since the model will have to reflect the 
order of the volumes in the 4D file.

During the `03_prepare_4D_for_randomise.sh` a text file was created reflecting
the order used in the creation of the 4D file. Of course there is one for the
N14 and one for the N26 subs set. 

A copy of those files is in the present directory:
- N14_4D_order.csv
- N26_4D_order.csv

Since this is a repeated measurement, we need to add additional columns 
modelling the mean for each participant across ratings. We try to do this
programmatically here to avoid mistakes if this would be done manually.

```{r, message=FALSE}
library(tidyverse)

filename_4D_order <- "N14_4D_order"

# Vector of zeropadded sub_ids
subs_file <- "/data00/leonardo/RSA/sub_list.txt"
subs <- sprintf("%02d",readLines(subs_file) %>% as.numeric)

```


```{r, message=FALSE}

# read the file
df <- read_csv(paste0(filename_4D_order,".csv"), col_names = "filename") %>% 
  
  # extract sub and rating columns
  mutate(sub = str_extract(filename, paste0(subs, collapse = "|"))) %>% 
  mutate(rating = str_extract(filename, "arousal|emotion|valence")) %>% 
  
  # create columns for emotion, arousal, valence
  mutate(arousal = ifelse(rating == "arousal",1,0)) %>% 
  mutate(emotion = ifelse(rating == "emotion",1,0)) %>% 
  mutate(valence = ifelse(rating == "valence",1,0)) 
  
```


```{r}

# create columns to model the mean of each sub
df_means <- unique(df$sub) %>% map_dfc(~ ifelse(df$sub == .x, 1,0))
colnames(df_means) <- unique(df$sub)

```


```{r}
# column bind to get the final table
df_model <- cbind(df, df_means)
df_model

write_csv(df_model, paste0(filename_4D_order,"_model.csv"))


```






```{r}
# Define the subjects and conditions
subjects &lt;- c("02", "03", "09", "11", "12", "15", "19", "20", "22", "23", "26", "28", "29", "32")
conditions <- c("arousal", "emotion", "valence")


# Generate the filenames using expand.grid and paste
df <- expand.grid(sub = subs, condition = conditions)  %>% 
  mutate(filename = paste0("sub_", sub, "_RSA_", condition, ".nii.gz")) %>%
  select(filename)
```






