---
title: "RSA results"
subtitle: "Inferential stats with permutation testing"
author: "Leonardo Cerliani"
date: "20/5/2023"
output:
  html_document:
    self_contained: true
    code_folding: hide
    highlight: pygments
    theme: cerulean
    css: styles.css
---

Libs and bd_results
```{r, message=FALSE}
# install.packages("papayaWidget")
# devtools::install_github("muschellij2/papayaWidget")

library(papayaWidget)
library(RNifti)
library(tidyverse)

# papaya widget github : https://github.com/muschellij2/papayaWidget
# app code here: https://github.com/muschellij2/linked_viewer/blob/master/app.R

bd_results = "/data00/leonardo/RSA/analyses/rsa/rsa_results"

pthr = 0.95

```


Some defs/fns to make strings to pass to Papaya concise
```{r, message=FALSE}
# define the MNI and the Dummy (for transparency of the MNI)
MNI=paste0(bd_results,"/","MNI152_T1_2mm_brain.nii.gz")
Dummy <- paste0(bd_results,"/","Dummy.nii.gz")

stat_type = "_tfce_corrp_tstat1.nii.gz"

# this function takes the contrast name and the params above to
# build the pathname of the contrast of interest to be fed
# into papaya
#
# E.g: f("Arousal"); f("Arousal_vs_Emotion")
f <- function(contrast_name) {
  template <- paste0(bd,"/stats_CONTRAST",stat_type)
  contrast_path <- str_replace(template, "CONTRAST", contrast_name)
  return(contrast_path)
}
```


# N14
```{r, message=FALSE}
flavour = "N14_GM_clean_bilat"
bd=paste0(bd_results,"/",flavour)
```


## N14 Arousal, Emotion, Valence > 0
- Arousal : Red
- Emotion : Blue
- Valence : Green
```{r, message=FALSE}

papaya(
  c(Dummy, MNI, f("Arousal"), f("Emotion"), f("Valence")),
  options = list(
    papayaOptions(alpha = 1, lut = "Grayscale"),
    papayaOptions(alpha = 0.5, lut = "Grayscale"),
    papayaOptions(alpha = 1, lut = "Red Overlay", min = pthr, max = 1),
    papayaOptions(alpha = 1, lut = "Blue Overlay", min = pthr, max = 1),
    papayaOptions(alpha = 1, lut = "Green Overlay", min = pthr, max = 1)
  )
)

```

<br><br>

## N14 Arousal > Emotion/Valence
- Arousal > Emotion : Red
- Arousal > Valence : Blue
```{r}

papaya(
  c(Dummy, MNI, f("Arousal_vs_Emotion"), f("Arousal_vs_Valence")),
    options = list(
    papayaOptions(alpha = 1, lut = "Grayscale"),
    papayaOptions(alpha = 0.5, lut = "Grayscale"),
    papayaOptions(alpha = 1, lut = "Red Overlay", min = pthr, max = 1),
    papayaOptions(alpha = 1, lut = "Blue Overlay", min = pthr, max = 1)
  )
)

```

<br><br>

## N14 Emotion > Arousal/Valence
- Emotion > Arousal : Red
- Emotion > Valence : Blue
```{r}

papaya(
  c(Dummy, MNI, f("Emotion_vs_Arousal"), f("Emotion_vs_Valence")),
    options = list(
    papayaOptions(alpha = 1, lut = "Grayscale"),
    papayaOptions(alpha = 0.5, lut = "Grayscale"),
    papayaOptions(alpha = 1, lut = "Red Overlay", min = pthr, max = 1),
    papayaOptions(alpha = 1, lut = "Blue Overlay", min = pthr, max = 1)
  )
)

```



<br><br>

## N14 Valence > Emotion/Arousal
- Valence > Arousal : Red
- Valence > Emotion : Blue
```{r}

papaya(
  c(Dummy, MNI, f("Valence_vs_Arousal"), f("Valence_vs_Emotion")),
    options = list(
    papayaOptions(alpha = 1, lut = "Grayscale"),
    papayaOptions(alpha = 0.5, lut = "Grayscale"),
    papayaOptions(alpha = 1, lut = "Red Overlay", min = pthr, max = 1),
    papayaOptions(alpha = 1, lut = "Blue Overlay", min = pthr, max = 1)
  )
)

```














<br><br>

# N26
```{r, message=FALSE}
flavour = "N26_GM_clean_bilat"
bd=paste0(bd_results,"/",flavour)
```


## N26 Arousal, Emotion, Valence > 0
- Arousal : Red
- Emotion : Blue
- Valence : Green
```{r, message=FALSE}

papaya(
  c(Dummy, MNI, f("Arousal"), f("Emotion"), f("Valence")),
  options = list(
    papayaOptions(alpha = 1, lut = "Grayscale"),
    papayaOptions(alpha = 0.5, lut = "Grayscale"),
    papayaOptions(alpha = 1, lut = "Red Overlay", min = pthr, max = 1),
    papayaOptions(alpha = 1, lut = "Blue Overlay", min = pthr, max = 1),
    papayaOptions(alpha = 1, lut = "Green Overlay", min = pthr, max = 1)
  )
)

```

<br><br>

## N26 Arousal > Emotion/Valence
- Arousal > Emotion : Red
- Arousal > Valence : Blue
```{r}

papaya(
  c(Dummy, MNI, f("Arousal_vs_Emotion"), f("Arousal_vs_Valence")),
    options = list(
    papayaOptions(alpha = 1, lut = "Grayscale"),
    papayaOptions(alpha = 0.5, lut = "Grayscale"),
    papayaOptions(alpha = 1, lut = "Red Overlay", min = pthr, max = 1),
    papayaOptions(alpha = 1, lut = "Blue Overlay", min = pthr, max = 1)
  )
)

```

<br><br>

## N26 Emotion > Arousal/Valence
- Emotion > Arousal : Red
- Emotion > Valence : Blue
```{r}

papaya(
  c(Dummy, MNI, f("Emotion_vs_Arousal"), f("Emotion_vs_Valence")),
    options = list(
    papayaOptions(alpha = 1, lut = "Grayscale"),
    papayaOptions(alpha = 0.5, lut = "Grayscale"),
    papayaOptions(alpha = 1, lut = "Red Overlay", min = pthr, max = 1),
    papayaOptions(alpha = 1, lut = "Blue Overlay", min = pthr, max = 1)
  )
)

```



<br><br>

## N26 Valence > Emotion/Arousal
- Valence > Arousal : Red
- Valence > Emotion : Blue
```{r}

papaya(
  c(Dummy, MNI, f("Valence_vs_Arousal"), f("Valence_vs_Emotion")),
    options = list(
    papayaOptions(alpha = 1, lut = "Grayscale"),
    papayaOptions(alpha = 0.5, lut = "Grayscale"),
    papayaOptions(alpha = 1, lut = "Red Overlay", min = pthr, max = 1),
    papayaOptions(alpha = 1, lut = "Blue Overlay", min = pthr, max = 1)
  )
)

```






















