---
title: "Distances in R"
author: "LC"
date: "2024-03-10"
output: html_document
---

# Motivation
When dealing with distances, every language and library has its own peculiarities,
which can go unnoticed, leading to big mistakes. 

Let's see how distances can be calculated
in R and how the vectors and matrices should be passed to the distance calculating
functions

# Libraries
Correlation-based distances can be calculated with `cor`, while a few fundamental
distances can be calculated with `stats::dist`. 

The library [`proxy::dist`](https://cran.r-project.org/web/packages/proxy/proxy.pdf) extends the
`stats::dist` function with many kind of distance metrics. 

Another package which provides distance calculation is [`philentropy::distance`](https://cran.r-project.org/web/packages/philentropy/vignettes/Distances.html)

Here we will look at how to work with `proxy` and its relation to `cor`. We use the Pearson correlation coefficient, however the same usage is applied to all listed measures.

```{r, message=FALSE}
library(tidyverse)
library(proxy)
```

A comprehensive list of the distances and similarity measures implemented in the `proxy` package can be obtained with `summary(pr_DB)`, while `pr_DB$get_entries()` gives a description with references about the implemented measures
```{r}
summary(pr_DB)
# pr_DB$get_entries()
```



# Distance / similarity between single vectors

First of all, when a vector is defined, R does not store it as a row or column
vector - like e.g. in Matlab. This property becomes apparent only when we either 
`rbind` or `cbind` multiple vectors

```{r}
set.seed(124)

x <- rnorm(100)
y = -0.8*x + 0.5*rnorm(100)
# plot(x ~ y)

dim(x)

rbind(x,y) %>% dim

cbind(x,y) %>% dim

```


Given two vectors x and y we can then calculate their correlation using `cor` or
with `dist([],method="correlation")`. However while `cor` accepts single vectors,
`dist` requires the two vectors to be _concatenated by row in a matrix_ of 2-by-length(x or y)
size.

**Importantly**, although "correlation" is listed among the similarity measures, if
`dist` is used, it will actually return 1-correlation. To get the actual correlation
we can either use the `proxy::simil` function or just calculate `1 - dist(method="correlation")`

**Also note** that when using `dist`, in any case the _absolute value_ of 1-correlation is 
returned. So to avoid confusion, it is definitively better to remember to use `simil` in these
cases

```{r}
# NB!!! When using "correlation" in dist(), it is actually 1-r
cor(x,y, method = "pearson")
simil(rbind(x,y), method = "correlation") %>% as.vector()
1 - dist(rbind(x,y), method = "correlation") %>% as.vector()
```

Note that both `simil` and `dist` return 1-by-1 matrices, therefore we probably want
to convert them `as.vector` as done above.

We can also pass a matrix to `cor`, however in this case the vectors should be _column binded_. 
Also in this case it will return a matrix with ones on the diagonal (the correlation of each
vector with itself)
```{r}
cor(cbind(x,y), method = "pearson")
```


# Distance / Similarity matrices

The reason why all methods in `proxy` return matrices is that they are mainly thought
to calculate distance / similarity among many vectors.

As hinted above, in this case we need to _row-bind_ the vectors.

Again, note how the `dist` function returns the _absolute_ value of 1-correlation.

```{r}
set.seed(124)

z = 0.5*y + 0.4*x + rnorm(100)

1 - dist(rbind(x,y,z), method = "correlation")
simil(rbind(x,y,z), method = "correlation")
```

The `dist` and `simil` functions return not the full distance / similarity matrix.
Since the latter is symmetrical, they instead return the lower triangular (tril).

To get the full matrix, we can convert this `as.matrix`, which is the same that 
we would obtain by using the `cor` function.

Another difference is that in the full matrix returned by `simil`, the diagonal
is set to `NA`

```{r}

simil(rbind(x,y,z), method = "correlation") %>% as.matrix

cor(cbind(x,y,z))
```


# Summary

- The [`proxy::dist`](https://cran.r-project.org/web/packages/proxy/proxy.pdf) library contains many readily available distance metrics. It can be extended by user-defined distances

- The single vectors among which the similarity / distance one wishes to calculate should be 
_row binded_ (`rbind`). In other words, the provided matrix should have the standard arrangement with observations on the rows and variables on the columns

- The output is the lower triangular of the similarity / distance matrix. To convert it to full
matrix, pass the output to `as.matrix`



# Appendix - function to handle different measures
```{r}

# ------- DDOS - Do Distance Or Similarity -------
# The input matrix should be observations-by-variables, that is:
# - rows index observations
# - columns index variables
# This is the format required by dist(), while for cor() is the
# opposite, so when using the cor() function we pass t(X) 

DDOS <- function(X, method) {
  
  X[is.na(X)] = 0
  X[is.infinite(X)] = 0
  
  switch(method,
    pearson = {
      D <- cor(t(X), method = "pearson") %>% as.dist()
    },
    
    spearman = {
      D <- cor(t(X), method = "spearman") %>% as.dist()
    },
    
    euclidean = {
      D <- dist(X, method = "euclidean")
    },
    
    cosine = {
      D <- simil(X, method = "cosine")
    },
    
    mahalanobis = {
      D <- dist(X, method = "mahalanobis")
    },
    stop("Supported methods: 'pearson', 'spearman', 'euclidean', 'cosine' or 'mahalanobis'.")
  )
  return(D)
}

# # example usage:
# DDOS(Y, method = "cosine")


```


```{r}

DDOS(Y, "correlation")
cor(t(Y), method = "pearson") %>% as.dist

DDOS(Y, "spearman")
cor(t(Y), method = "spearman") %>% as.dist

DDOS(Y, "euclidean")
dist(Y, method = "euclidean")

DDOS(Y, "cosine")
simil(Y, method = "cosine")


```



```{r}

x = c(0,0,0)
y = c(1,1,1)
z = c(3,4,5)

X <- rbind(x,y,z)
X

s1 <- 1
s2 <- 2
s3 <- 3
S = rbind(s1,s2,s3)

dist(S, method = "euclidean") %>% as.matrix

set.seed(124)
y1 <- rnorm(100,1,2)
y2 <- 0.7*y1 + 0.3*rnorm(100,2,3)
y3 <- -0.7*y1 + 0.3*rnorm(100,4,5)
Y <- rbind(y1,y2,y3)

X %>% dist(method = "euclidean") %>% as.matrix() %>% as.dist()
Y %>% dist(method = "Mahalanobis") %>% as.matrix() %>% as.dist()

cov(t(Y)) %>% inv


Y %>% simil(method = "cosine")
Y %>% simil(method = "correlation")
t(Y) %>% cor(method = "pearson") %>% as.dist()
t(Y) %>% cor(method = "spearman") %>% as.dist()

```



```{r}
# pr_DB$get_entries()
pr_DB %>% summary
```



## Mahalanobis distance
```{r}

Y <- matrix(1:5, nrow=5)
Y <- matrix(rnorm(200), nrow=20)

# Step 1: Calculate the covariance matrix
(Sigma <- cov(Y))

# Step 2: Calculate the inverse of the covariance matrix
(Sigma_inv <- solve(Sigma))

# Step 3: Calculate the Mahalanobis distance matrix
n <- nrow(Y)
D <- matrix(NA, n, n) # Initialize a matrix to store the distances

for(i in 1:n) {
  for(j in i:n) { # Can start from i to avoid redundant calculations
    diff <- Y[i,] - Y[j,]
    D[i,j] <- sqrt(t(diff) %*% Sigma_inv %*% diff)
    D[j,i] <- D[i,j] # The distance matrix is symmetric
  }
}

D

```



```{r}
Y <- matrix(rnorm(200), nrow=20)

cor(t(Y)) %>% dim

```



