---
title: "RSA v3"
output: html_notebook
---

```{r load_libraries, message=F}
library(tidyverse)
library(future)
library(furrr)
library(tictoc)
library(RNifti)
library(proxy) # distances
library(profvis)

bd="/data00/leonardo/RSA/analyses"

ratings_type <- "emotion"

# Vector of zeropadded sub_ids
subs_file <- "/data00/leonardo/RSA/sub_list.txt"
subs <- sprintf("%02d",readLines(subs_file) %>% as.numeric)


```


# Aux functions
```{r}
# plot_heatmap(D %>% as.matrix)
plot_heatmap <- function(M) {
  heatmap(M, Rowv = NA, Colv = NA, symm=T, revC = T)
}
```


# Create a df_path_copes with the location of the 56 cope niis from the one_movie_per_ev model
Extract the pathname of all copes using the `list.files()` function.
Also define a copes_numba vector with all the copes numbers.

NB: The cope numbers in the `cope` column are NOT zeropadded since this is how they come out from FSL Feat
```{r}

bd_copes = paste0(bd, "/one_ev_per_movie/results/2nd_level")

df_path_copes <- list.files(bd_copes, recursive = T) %>% tibble()
names(df_path_copes) <- "fname"

df_path_copes <- df_path_copes %>%
  filter(str_detect(fname, "gfeat/cope[0-9]+.feat/stats/cope")) %>%
  tidyr::separate(
    fname, c("sub","cope","tmp1","tmp2"),
    sep = "/", fill = "right", remove = F
  ) %>% 
  select(!starts_with("tmp")) %>% 
  mutate(sub = str_extract(sub, "[0-9]+")) %>% 
  mutate(cope = (str_extract(cope, "[0-9]+") %>% as.numeric) ) %>% 
  mutate(path = paste0(bd_copes,"/",fname)) %>% 
  select(!fname) %>% 
  arrange(sub,cope)

df_path_copes

copes_numba <- df_path_copes$cope %>% unique

write_csv(df_path_copes, paste0(bd,"/rsa/copes_location.csv"))


```


# Read atlas and create a vector of atlas labels
```{r}

bd_atlases = "/data00/leonardo/RSA/analyses/rsa/atlases"

atlas_filename <-  "juelich_2mm.nii.gz"
atlas_path <- paste0(bd_atlases,"/",atlas_filename)

atlas_nii <- readNifti(paste0(bd_atlases,"/",atlas_filename))

region_labels <- atlas_nii[atlas_nii > 0] %>% unique %>% sort


```



# Calculate RATINGS RDMs
We do this first since the RDM for the ratings will be the same for whatever
atlas will be used. Also, it's much faster than the RDM_fmri and can be done
for all the sub_ids at once.

The final RDMs_rats has 40040 nrows : 26 sub_ids * 1540, where the latter
derives from ((56^2)-56)/2 = 1540, i.e. tril from the D matrix of 56 movies

```{r, message=FALSE}

bd_ratings = paste0(bd,"/RATINGS")
ratings_path <- paste0(bd_ratings,"/",ratings_type,"_ratings.csv")
rats <- read_csv(ratings_path)

RDMs_rats <- subs %>% map_dfc(~ {
  
  D_rats <- rats %>% 
  filter(sub == .x) %>% 
  select(starts_with("r_")) %>% 
  dist(method = "euclidean")
  
  D_rats_tril <- D_rats[!is.na(D_rats)]

  tibble(!!.x := D_rats_tril)
}) %>% 
  pivot_longer(cols = everything(), names_to = "sub") %>% 
  arrange(sub)


# # alternative with nested columns
# # following https://www.youtube.com/watch?v=rz3_FDVt9eg  26:57
# do_extract_tril <- function(df_rats) {
#   D_rats <- df_rats %>% dist(method = "euclidean")
#   D_rats_tril <- D_rats[!is.na(D_rats)]
# }
# 
# ff <- rats %>%
#   select(sub, starts_with("r_")) %>% 
#   group_by(sub) %>% 
#   nest %>% 
#   mutate(RDM_rats = data %>% map(do_extract_tril))
# 
# ff[ff$sub==sub_id,]$RDM_rats
# 
# # 1540 * 26 sub_ids
# ff %>% unnest(RDM_rats)

RDMs_rats
```




# Calculate FMRI RDMs

## Define the functions
1. df_copes[one_sub] <- load_sub_copes(sub_id, copes_numba)
2. RDM_fmri[one_sub] <- calculate_fmri_RDMs(df_copes, atlas_path)

```{r load_copes}

# Load copes for one sub_id
# (reasonably fast: reads 56 nifti at 2mm resolution in ~ 2.5 sec)
sub_id = "02"


# Function : load_sub_copes(sub_id, copes_numba)
# Loads the nii for all 56 copes for one sub, and returns a df of (91x109x91) rows
# and 56 (length(copes_numba)) columns, each one named with the cope numba
load_sub_copes <- function(sub_id, copes_numba) {
  copes_numba %>% map_dfc(~ {
    nii <- df_path_copes %>% 
      filter(sub == sub_id, cope == .x) %>% pull %>%  
      readNifti %>% as.vector
    tibble(!!paste0("cope_", .x) := nii)
  }) 
}

tic()
df_copes <- load_sub_copes(sub_id, copes_numba)
toc()

# Read atlas and calculate fmri RDMs : the final df has size n_tril_D-by-n_regions, where 
# - n_tril_D : number of elements in the distance matrix = ((56^2)-56)/2 = 1540
# - n_regions : number of distinctly labelled regions in the atlas 
calculate_fmri_RDMs <- function(df_copes, atlas_path) {
  
  # Calculate the RDM for each atlas region, 
  # i.e. distance matrix of the betas of each movie across all voxels in that ROI.
  # Returns a df of length(triu(D))-by-n_atlas_regions
  region_labels %>% map_dfc(~ {
    idx <- which(atlas_nii == .x)
    df_copes_region <- df_copes[idx, ]
    
    # The following line will be replaced by the calculate_distance() fn
    D_fmri <- dist( t(df_copes_region), method = "euclidean" )
    
    D_fmri <- D_fmri[!is.na(D_fmri)]
    tibble(!!paste0("RDM_region_", .x) := D_fmri)
  })
}

tic()
one_RDM_fmri <- calculate_fmri_RDMs(df_copes, atlas_path)
toc()
```


## Furrr it up
First, the copes are loaded into a df of 91x109x91 (MNI 2mm) rows and 56 columns.
Then the RDM is estimated for each atlas region as the tril of the similarity. 
matrix across the 56 subs
This is carried out in parallel for all 26 subs using `furrr::future_map_dfr`
```{r}
# 5 workers give the best performance (~ 15 sec)
plan(multisession, workers = 5)

tic()
RDMs_fmri <- subs %>% future_map_dfr(~{
  
  paste0("Calculating RDMs for sub ",.x,"\n") %>% cat
  
  df_copes <- load_sub_copes(sub_id, copes_numba)
  one_RDM_fmri <- calculate_fmri_RDMs(df_copes, atlas_path)
  
  one_RDM_fmri %>% mutate(sub = .x) %>% relocate(sub)
  
})
toc()

plan(sequential)
```


# RSA 
In the approach below purrr, tidy data and R in general shine at their best.

The final aim is to calculate the similarity between each sub/region fmri RDM
and each sub rat RDM. Therefore the fmri rdms is grouped by sub/region and the
rat rdms is grouped only by sub.

First, nesting both fmri and rat rdms allows to keep the dfs tidy. 
Soon after nesting, we can *right*-join the fmri to the rat, to create a copy
of the rat rdm for each sub/region of the atlas.

At this point, calculating the similarity is just a matter of using a purrr:map2
(or pmap as I prefer) to the fmri and rat columns

The simplicity yielded by the tidy data approach and the purrr functions is 
unparalleled.

```{r}


# nest the RDMs for both rat and fmri. In the second case, they are nested by sub AND region
rdms_fmri_nested <- RDMs_fmri %>% 
  pivot_longer(cols = starts_with("RDM"), names_to = "region") %>% 
  group_by(sub, region) %>% 
  nest %>% 
  rename(fmri = data)

rdms_rats_nested <- RDMs_rats %>% 
  group_by(sub) %>% 
  nest %>% 
  rename(rat = data)

# we put a copy of the rating RDM for each atlas region's RDM
joint_fmri_rat_nested <- right_join(rdms_fmri_nested,rdms_rats_nested, by = "sub")
# # the following is just to check that the join went as expected
# ff %>% mutate(sum_rat = rat %>% map_dbl(sum)  )

joint_fmri_rat_nested

# Calculate all similarity
# This is effectively when we are calculating the RSA
all_similarities <- joint_fmri_rat_nested %>% 
  mutate(similarity = list(fmri,rat) %>% pmap_dbl(~ cor(.x,.y)) ) %>% 
  ungroup

# Produce estimates of RSA acros subjects
all_similarities %>%
  select(region, similarity) %>% 
  group_by(region) %>% 
  reframe(
    mean_similarity = round(mean(similarity),2),
    sd_similarity = round(sd(similarity),2)
  ) %>% 
  arrange(desc(mean_similarity))


```














# Try proxy for distances

```{r, eval=F}
library(proxy)
# pr_DB$get_entries() 
# summary(pr_DB)


# df_region <- df[idx,]
# D <- dist( t(df_region) )
# dim(as.matrix(D))
# # plot_heatmap(D %>% as.matrix)
# D[!is.na(D)]

df_region <- df[idx,]

D <- dist( t(df_region) )
D <- dist(t(df_region), method = "euclidean")

dim(as.matrix(D))
plot_heatmap(D %>% as.matrix)
# D[!is.na(D)]

heatmap(D %>% as.matrix, symm = T)

types = c("angular","correlation","cosine","Euclidean") 
par(mfrow=c(2,2))
for (i in types) {
  D = proxy::dist(t(df_region), method = i)
  image(D %>% as.matrix, axes = F, asp=1)
  title(i)
}

```


# Explore proxy distances package
```{r, eval=F}

library(proxy)
# pr_DB$get_entries() 
# summary(pr_DB)

set.seed(124)

M = matrix(runif(100*10), ncol = 100)

method <- "correlation"

Dstats <- dist(M, method = method)
Dproxy <- proxy::dist(M, method = method)

sum(Dstats-Dproxy)


par(mfrow = c(1,2))
image(Dstats %>% as.matrix()); title("pippo")
image(Dproxy %>% as.matrix()); title("topolino")



```


















