---
title: "do randomise emotion vs arousal"
author: "LC"
date: "2024-05-02"
output: html_document
---


```{r, message=FALSE}
library(tidyverse)
library(glue)

rsa_flavour <- "insula_HO_GM_EER"
mask_name <- "insula_HO_GM"

mask_file <- dir("/data00/leonardo/RSA/analyses/rsa/masks/", pattern = mask_name, full.names = T)

bd <- "/data00/leonardo/RSA/analyses/rsa/rsa_results"

# rd = results_dir
rd <- paste0(bd,"/",rsa_flavour)
```


# Create the model for a two-samples paired t-test
```{r}
df <- tibble(path = dir(rd, pattern = "sub"))

df <- df %>% 
  separate_wider_delim(path, "_", names = c("prefix","sub","rsa","model")) %>% 
  select(sub,model) %>% 
  mutate(sub = factor(sub)) %>% 
  mutate(emotion = ifelse(str_detect(model,"emotion"),1,0 )) %>% 
  mutate(arousal = ifelse(str_detect(model,"arousal"),1,0 )) %>% 
  arrange(desc(model))

df <- df %>% 
  mutate(emo_minus_arousal = emotion - arousal) %>% 
  select(sub, model, emo_minus_arousal)

# Create dummy variables using model.matrix
dummy_vars <- model.matrix(~ sub - 1, data = df)  # The '-1' removes the intercept

# Convert to a tibble and bind back to the original data
df_dummies <- as_tibble(dummy_vars)
df_with_dummies <- bind_cols(df, df_dummies)

df_with_dummies
```



# Extract only the design.txt and write it to disk
```{r}

df_with_dummies %>%
  select(-sub, -model) %>%
  write_delim(
    paste0(rd,"/design.txt"),
    delim = " ",
    col_names = FALSE
  )

```


# Create the contrast contrasts.txt
```{r}

df_contrast <- tribble(
  ~contrast_name,  ~x, 
  "emotion_bigger_than_arousal", 1,
  "arousal_bigger_than_emotion", -1,
)


N_sub <- df$sub %>% unique %>% length

dummy_contrasts <- matrix(0, nrow(df_contrast), ncol = N_sub ) %>% as_tibble()

df_contrast_with_dummies <- cbind(df_contrast, dummy_contrasts)
df_contrast_with_dummies


# Write the matrix to a file
df_contrast_with_dummies %>% 
  select(!c("contrast_name")) %>% 
  write_delim(
    paste0(rd,"/contrasts.txt"),
    delim = " ",
    col_names = FALSE
  )

```


# Prepare the list of niis to be put in the 4D

**BE VERY CAREFUL!** The order should match that in the
`design.mat` and `design.con`. 

That's why we use the order in the `df_with_dummies`, which was
used to create `design.mat`

Also, we create sysenv vars so that we can use them in the subsequent bash
cell to run randomise
```{r}

nii_list <- df_with_dummies %>% 
  select(sub,model) %>% 
  mutate(filename = glue('sub_{sub}_RSA_{model}')) %>% 
  select(filename) %>% pull %>% paste(collapse = " ")

Sys.setenv(NII_LIST = nii_list)
Sys.setenv(RESULTS_DIR=rd)
Sys.setenv(MASK_FILE=mask_file)
```




```{bash}
FSLDIR=/data00/leonardo/warez/fsl

PATH=${FSLDIR}/share/fsl/bin:${PATH}
export FSLDIR PATH
. ${FSLDIR}/etc/fslconf/fsl.sh

echo cd to ${RESULTS_DIR} 
cd ${RESULTS_DIR}

echo creating the 4D file for randomise
fslmerge -t emotion_arousal_4D ${NII_LIST}

Text2Vest design.txt design.mat
Text2Vest contrasts.txt contrasts.con

time randomise \
  -i emotion_arousal_4D.nii.gz \
  -o stats \
  -d design.mat \
  -t contrasts.con \
  -m ${MASK_FILE} \
  -T -n 5000 \
  --quiet

```





























