---
title: "do_RSA_v10_Searchlight"
author: "LC"
date: "2024-05-03"
output: html_document
---

# Load libraries
```{r load_libraries, message=F}
library(tidyverse)
library(future)
library(furrr)
library(tictoc)
library(RNifti)
library(proxy) # distances
library(profvis)
library(DT)
library(formattable)

source("funs_V10_Searchlight.R")
```



# Set the following paramters manually
```{r}

# ratings_type can be emotion, arousal, valence
ratings_type <- "emotion"

# must be one of the masks in rsa/masks
gm_file <- "GM_clean"

# Define searchlight radius in mm
r_mm <- 4

# if set to TOP_RATERS, it will use only the subs below
subs_selection = "ALL_SUBS"
top_raters_subs <- c("02","03","12","11","22","09","29","28","26","32","23","15","20","19")

```


# Other parameters
Do not touch unless you know what you are doing
```{r}
bd="/data00/leonardo/RSA/analyses"

# the following is our contrast of interest and it's always the same
copes_type <- "one_ev_per_movie"

# full path of the gm mask onto which the searchlight will be used
gm_path <- paste0(bd,"/rsa/masks/", gm_file, ".nii.gz")


# Choose the distance metric to use for fmri RDM
# The metric for ratings RDM *should be* euclidean, since arousal and valence
# ratings have only one value
# Supported methods: 'pearson', 'spearman', 'euclidean', 'cosine' or 'mahalanobis'
dist_method_rating <- "euclidean"
dist_method_fmri = "euclidean"
dist_method_rsa = "pearson"


# for RDM_type use:
# - tril for the lower triangular of the D matrix
# - svd for the first 3 pc from svd of the D matrix
# - in case of svd, set the number of components to return with ncomp_svd
RDM_type = "tril"  # svd or tril. Must be the same for both rating and fmri
ncomp_svd = 3  


# Vector of zeropadded sub_ids
subs_file <- "/data00/leonardo/RSA/sub_list.txt"
subs <- sprintf("%02d",readLines(subs_file) %>% as.numeric)

if (subs_selection == "TOP_RATERS") {
  subs <- top_raters_subs
}

```


# Create the directory where the results will be stored
This will allow us also to store a text file with all the info about the analysis

```{r}
results_flavour <- generate_results_flavour(
  gm_file, dist_method_rating, dist_method_fmri, dist_method_rsa, subs
)

rsa_results_path <- paste0(bd, "/rsa/rsa_results")
results_dir <- paste0(rsa_results_path, "/", results_flavour)

# Create the results if it does not exist
if (!dir.exists(results_dir)) {
  # Create the directory if it does not exist
  dir.create(results_dir, recursive = TRUE)
  print(paste0("Created ", results_dir))
} else {
  print(paste0(results_dir, " already exists"))
}
```


Store important information about the parameters for the analysis in a text file

```{r}

# redirect the cat msgs to the logfile
logfile <- paste0(results_dir, "/", ratings_type, "_log.txt")
sink(logfile, split = TRUE)


cat("\n")
cat("Using", ratings_type, "rating\n\n")
cat("Calculating RSA for every voxel in the", gm_file, "mask\n") 
cat("using a searchlight of", r_mm, "mm\n\n")
cat(paste0("Ratings RDM to be calculated with : ", dist_method_rating, "\n"))
cat(paste0("fMRI RDM to be calculated with : ", dist_method_fmri, "\n\n"))

# If subs_selection is TOP_RATERS only the subs with highest movie/emotion
# congruency - defined above - will be included
if (subs_selection == "TOP_RATERS") {
  subs <- top_raters_subs
  cat("Only the", length(subs), "top raters")
} else {
  cat("All", length(subs), "subs")
}

cat("\n\n")

# stop sinking the cat msgs into the logfile
sink()

```






# Aux functions
```{r}
# plot_heatmap(D %>% as.matrix)
plot_heatmap <- function(M) {
  heatmap(M, Rowv = NA, Colv = NA, symm=T, revC = T)
}
```


# Create a df_path_copes with the location of the 56 cope niis from the one_movie_per_ev model
Extract the pathname of all copes using the `list.files()` function.
Also define a copes_numba vector with all the copes numbers.

NB: The cope numbers in the `cope` column are NOT zeropadded since this is how they come out from FSL Feat
```{r, message=FALSE}

copes_location_csv <- paste0(bd,"/rsa/copes_location.csv")

df_path_copes <- if (file.exists(copes_location_csv) ) {
  df_path_copes <- read_csv(copes_location_csv)
} else {
  df_path_copes <- import_df_path_copes(bd, copes_type)
}

copes_numba <- df_path_copes$cope %>% unique

# df_path_copes

```





# Calculate RATINGS RDMs
We do this first since the RDM for the ratings will be the same for whatever
atlas will be used. Also, it's much faster than the RDM_fmri and can be done
for all the sub_ids at once.


- **tril** : The final RDMs_rats has 40040 nrows : 26 sub_ids * 1540, where the latter
derives from ((56^2)-56)/2 = 1540, i.e. tril from the D matrix of 56 movies
- **svd** : The final RDMs_rats has 4368 nrows : 26 sub_ids * 168, where the latter
derives from 56*3, i.e. the first three components from svd of the D matrix
```{r, message=FALSE}

bd_ratings = paste0(bd,"/RATINGS")
ratings_path <- paste0(bd_ratings,"/",ratings_type,"_ratings.csv")
rats <- read_csv(ratings_path)

# cols are subs
# rows are elements of the tril
RDMs_rats_wider <- subs %>% map_dfc(~ {
  do_RDM_ratings_one_sub(.x, dist_method_rating, rats, RDM_type =  RDM_type)
})

# pivot the df above so that there is only one col for subs
# and one for the values of the tril for each sub
RDMs_rats <- RDMs_rats_wider %>%
  pivot_longer(cols = everything(), names_to = "sub") %>%
  arrange(sub)

RDMs_rats
```


# Calculate fMRI RDMs

## Create SEARCHLIGHT regions

This code cell yields two outputs:

- `idx` : the index of each voxel in the mask, i.e. an integer from 1 to 91x109x91

- `idx_within_radius` : a list of list: each inner list contains the index of the
voxels which are within the defined radius, i.e. within the searchlight having that
voxel as a 'center'. 

Note that not all voxels have the same numba of other voxels
within their searchlight. The histogram below shows the distribution
of numba of voxels in the searchlight for each voxel in the mask.


```{r}

# 1. Read the gm mask and get voxel size, get idx of voxels in gm and their
#    xyz coordinates
gm <- readNifti(gm_path)
voxel_size <- pixdim(gm) %>% mean   # voxel size of gm
idx <- which(gm != 0)   # index of nonzero voxels
coords <- which(gm != 0, arr.ind = TRUE)   # get coordinates


# 2. Define a searchlight function returning the volume (92x109x92) idx of the 
#    voxels within the searchlight
searchlight <- function(center_coord, r_mm, coords, idx, voxel_size) {

  # Calculate distance of the given voxel from every other voxel, considering voxel size
  distances <- sqrt(rowSums((t(t(coords) - center_coord) * voxel_size)^2))
  
  # # TEST ONLY: return the coords of the voxels in the searchlight
  # return(coords[distances <= r_mm, , drop = FALSE])

  # return the idx of the voxels in the searchlight
  return(idx[distances <= r_mm])
}


# # Testing the searchlight fn with one coordinate 
# # See also the nb : earchlight_development.Rmd
# mycoord <- coords[1,]
# idx_within_radius <- searchlight(mycoord, r_mm=6, coords, idx, voxel_size)

# # Convert indices to coordinates (just to check)
# mycoord
# coords_within_radius <- coords[match(idx_within_radius, idx), , drop = FALSE]
# coords_within_radius


# 3. Get the idx of the vxls within radius for each voxel in gm (~ 20 sec)
plan(multisession, workers = 15)

tic()

idx_within_radius <- 1:length(idx) %>% future_map(
  ~ searchlight(coords[.x,], r_mm, coords, idx, voxel_size)
)

toc()

plan(sequential)

#
idx_within_radius %>% map_dbl(length) %>% hist


```



## Fn to carry out the RMDs Searchlight calculation for one sub

```{r, message=FALSE}

# In order to test for one sub, I need to load the corresponding df_copes.
# In the next cells the df_copes of each sub are dynamically loaded
# when subs %>% future_map_dfr
sub_id = "02"
df_copes <- load_sub_copes(sub_id, copes_numba, df_path_copes)


calculate_fmri_RDMs_Searchlight <- function(df_copes, idx, idx_within_radius) {
  
  one_rdm <- idx_within_radius %>% map_dfc(~ {
    idx_spot <- unlist(.x)
    df_copes_region <- df_copes[idx_spot, ]
    
    # center (i.e. demean) variables
    # (treat also the case of single voxels)
    if(nrow(df_copes_region) > 1) {
      df_copes_region_demean <- df_copes_region %>% 
        mutate(across(everything(), ~ .x - mean(.x)))  
    } else {
      df_copes_region_demean <- df_copes_region
    }
    
    # calculate distance / similarity
    D_fmri <- DDOS( t(df_copes_region_demean), method = dist_method_fmri )
    D_fmri[is.na(D_fmri)] <- 0 
    
    # flat onto a vector
    D_feature_vector <- D_fmri[!is.na(D_fmri)]
  })
  
  colnames(one_rdm) = idx
  
  return(one_rdm)
}

# Again the following is just for testing on one sub
# ~ 90 seconds, no need to use furrr
N <- 10 #length(idx)
one_RDM_fmri <- calculate_fmri_RDMs_Searchlight(df_copes, idx[1:N], idx_within_radius[1:N])

one_RDM_fmri

```



## Furrr the RDMs Searchlight calculation across subs
```{r, message=FALSE}

# for testing on a subset of idx
N = length(idx)

# 5 workers give the best performance
plan(multisession, workers = 10)

tic()
RDMs_fmri <- subs %>% future_map_dfr(~{
  
  paste0("Calculating RDMs Searchlight for sub ",.x,"\n") %>% cat
  
  df_copes <- load_sub_copes(.x, copes_numba, df_path_copes)
  
  one_RDM_fmri <- calculate_fmri_RDMs_Searchlight(
    df_copes, 
    idx[1:N], 
    idx_within_radius[1:N]
  )
  
  one_RDM_fmri %>% mutate(sub = .x) %>% relocate(sub)
  
})
toc()

plan(sequential)
```



# RSA Searchlight

## Calculate RSA and mean RSA across subs inside a tibble
```{r}

rdms_fmri_nested <- RDMs_fmri %>% 
  pivot_longer(cols = !starts_with("sub"), names_to = "region") %>% 
  group_by(sub,region) %>% 
  nest() %>% 
  rename(fmri = data)


rdms_rats_nested <- RDMs_rats %>% 
  group_by(sub) %>% 
  nest %>% 
  rename(rat = data)


# we put a copy of the rating RDM for each atlas region's RDM
joint_fmri_rat_nested <- right_join(rdms_fmri_nested,rdms_rats_nested, by = "sub")
# # the following is just to check that the join went as expected
# ff %>% mutate(sum_rat = rat %>% map_dbl(sum)  )

joint_fmri_rat_nested

# Calculate all RSA for each sub and each region/voxel
# and also round similarity and fix region data type
RSA <- joint_fmri_rat_nested %>% 
  mutate(
    similarity = list(fmri,rat) %>% pmap_dbl(~ cor(.x,.y, method = dist_method_rsa))
  ) %>%
  ungroup %>% 
  mutate(
    region = as.integer(region),
    similarity = round(similarity, 3)
  )

# Estimate avg_RSA
avg_RSA <- RSA %>%
  select(region, similarity) %>% 
  group_by(region) %>% 
  reframe(
    mean_similarity = round(mean(similarity),2),
    sd_similarity = round(sd(similarity),2)
  ) %>% 
  arrange(desc(mean_similarity))

avg_RSA %>% datatable()
```


## Write the RSA results to niis

Two types of results are written in nii images, for each voxel in the mask:
- the RSA for each sub
- the average RSA across subs

### Define `results_flavor` and create `results_dir` 

The following chunk is all to create an informative name for the results dir for this 
particular mask, rating type and distance methods, as well as to create the 
directory if it does not exist yet

**NB: The following section was moved to the very top in order to be able to** 
**write a text file with the parameters of the analysis**

```{r}

# # NB: [results_flavour].nii.gz will also be the name of the avg RSA across subs
# results_flavour <- generate_results_flavour(
#   gm_file, dist_method_rating, dist_method_fmri, dist_method_rsa, subs
# )
# 
# rsa_results_path <- paste0(bd, "/rsa/rsa_results")
# results_dir <- paste0(rsa_results_path, "/", results_flavour)
# 
# # Create the results if it does not exist
# if (!dir.exists(results_dir)) {
#   # Create the directory if it does not exist
#   dir.create(results_dir, recursive = TRUE)
#   print(paste0("Created ", results_dir))
# } else {
#   print(paste0(results_dir, " already exists"))
# }

```


### Write the RSA nii for each sub - these will be used by `randomise`

```{r}

# Using the function write_RSA_results_to_nii. Documentation in the funs_[Vx]_Searchlight.R

RSA %>%
  select(sub, region, similarity) %>% 
  group_split(sub) %>% 
  walk(~ write_RSA_results_to_nii(.x, ratings_type, results_dir, gm))

```




### Write the avg_RSA nii - for viz

```{r}

write_avg_RSA_to_nii <- function(avg_RSA, results_flavour, 
                                 ratings_type, rsa_results_path, gm) {
  
  results_nii <- gm
  
  idx_voxel <- avg_RSA$region
  results_nii[idx_voxel] = avg_RSA$mean_similarity
  
  nii2write_path <- paste0(rsa_results_path,"/", results_flavour, "_", ratings_type,".nii.gz")
  cat(paste0(nii2write_path,"\n"))
  
  writeNifti(results_nii, nii2write_path)
}


write_avg_RSA_to_nii(avg_RSA, results_flavour, ratings_type, rsa_results_path, gm)

```









