---
title: "Ratings separability"
author: "LC"
date: "2024-03-09"
output: html_document
---

Here we experiment with some dimred methods to assess the consistency of the ratings.

NB: umap requires reticulate. We need to install a version of reticulate which does not require `RcppTOML` (v1.4)
`remotes::install_version("reticulate", version = "1.4")`


# Load the ratings for all subs
```{r, message=FALSE}
library(tidyverse)
library(proxy)
library(tictoc)
library(patchwork)
library(ggstatsplot)

# pr_DB$get_entries() 
# summary(pr_DB)

bd="/data00/leonardo/RSA/analyses"

ratings_type <- "emotion"
bd_ratings = paste0(bd,"/RATINGS")
ratings_path <- paste0(bd_ratings,"/",ratings_type,"_ratings.csv")
rats <- read_csv(ratings_path)

# arrange by emotion to have a reference
rats <- rats %>% arrange(emotion)

subs_file <- "/data00/leonardo/RSA/sub_list.txt"
subs <- sprintf("%02d",readLines(subs_file) %>% as.numeric)

# plot_heatmap(D %>% as.matrix)
plot_heatmap <- function(M) {
  heatmap(M, Rowv = NA, Colv = NA, symm=T, revC = T)
}

# palette for different emotions
paletta <- c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3", "#FF7F00", "#FFFF33", "#A65628")

```


# Subject exclusion
```{r}

# # If you want to select subs, do it here
# subs_to_exclude <- c(7,34,24,10,16,31,30,27,05,21,14) %>% as.character()
# subs <- subs[!subs %in% subs_to_exclude]
# rats <- rats %>% 
#   filter(sub %in% subs)
# 
# subs

```




# List the available metrics
```{r}
summary(pr_DB)
```


# Prepare the D matrix
```{r}

labels <- rats$emotion
high_low <- str_extract(rats$high_low_code, "high|low")
length(labels)

dist_method <- "euclidean"

tic()
D <- rats %>% 
  filter(sub %in% subs) %>%
  arrange(emotion) %>% 
  select(starts_with("r_")) %>% 
  dist(method = dist_method) %>% 
  as.matrix
toc()

rownames(D) <- colnames(D) <- labels

dim(D)

# # replacing the NAs which come out mostly from neutral movies, which have
# # mostly scores at 0 across all emotion rating
#
# # check with the folliwing:
# rats %>%
#   mutate(r_sum = rowSums(select(., starts_with("r_")), na.rm = TRUE)) %>% 
#   select(video, r_sum) %>% 
#   filter(r_sum == 0)

D[is.na(D)] <- 0
# sum(is.na(D)) / length(D)
# sum(is.infinite(D)) / length(D)

plot_heatmap(D %>% as.matrix)
title(dist_method)

```


# SVD dimred
```{r}

# Perform SVD
svd_result <- svd(D)

# Extract the first two eigenvectors (U matrix in SVD)
U <- svd_result$u[,1:3]
S <- svd_result$d[1:3] %>% diag
US <- U %*% S

# plot pc1,2 and pc2,3
do_svd_lowdim_plots <- function(US, lowdim_x, lowdim_y) {
  
  svd_lowdim <- data.frame(pc_x = US[,lowdim_x], pc_y = US[,lowdim_y], labels, high_low)
  
  p <- ggplot(svd_lowdim, aes(x = pc_x, y = pc_y, color = labels, shape = high_low)) +
    geom_point(size = 3, alpha = 0.7) +  # Open circles for points
    scale_color_manual(values = paletta) +
    scale_shape_manual(values = c(1, 3)) +  # Assign specific shapes
    labs(
      title = "SVD Projection", 
      x = paste0("lowdim ",lowdim_x), 
      y = paste0("lowdim ",lowdim_y)
    ) +
    theme_minimal() +
    guides(color = guide_legend(override.aes = list(shape = 16)))  # Filled circles in the legend
  
  return(p)
}

p <- list(c(1,2),c(2,3)) %>% pmap(~ do_svd_lowdim_plots(US, lowdim_x = .x, lowdim_y = .y))

# library(patchwork)
p[[1]] + p[[2]]

```




# UMAP dimred
```{r, warning=FALSE}

library(umap)
# tunings here : https://cran.r-project.org/web/packages/umap/vignettes/umap.html

# # Perform UMAP
umap_result <- umap::umap(D, input="dist", n_components=5)

# plot pc1,2 and pc2,3
do_UMAP_lowdim_plots <- function(umap_result, lowdim_x, lowdim_y) {
  
  umap_lowdim <- data.frame(
    pc_x = umap_result$layout[,lowdim_x], 
    pc_y = umap_result$layout[,lowdim_y], 
    labels, high_low
  )
  
  p <- ggplot(umap_lowdim, aes(x = pc_x, y = pc_y, color = labels, shape = high_low)) +
    geom_point(size = 3, alpha = 0.7) +  # Open circles for points
    scale_color_manual(values = paletta) +
    scale_shape_manual(values = c(1, 3)) +  # Assign specific shapes
    labs(
      title = "UMAP Projection", 
      x = paste0("lowdim ",lowdim_x), 
      y = paste0("lowdim ",lowdim_y)
    ) +
    theme_minimal() +
    guides(color = guide_legend(override.aes = list(shape = 16)))  # Filled circles in the legend
  
  return(p)
}


p <- list(c(1,2),c(2,3)) %>% pmap(~ do_UMAP_lowdim_plots(umap_result, lowdim_x = .x, lowdim_y = .y))

# library(patchwork)
p[[1]] + p[[2]]

```

# Separability in arousal and valence
These are unidimensional, so we can just resort to a barplot

## Arousal
```{r, message=FALSE, warning=FALSE}

ratings_type <- "arousal"
bd_ratings = paste0(bd,"/RATINGS")
ratings_path <- paste0(bd_ratings,"/",ratings_type,"_ratings.csv")
rats <- read_csv(ratings_path)

rats_arousal <- rats %>% 
  select(sub, emotion, r_arousal) %>% 
  # filter(sub == "02") %>% 
  arrange(emotion) %>%
  group_by(sub,emotion) %>% 
  reframe(
    median_within_sub = median(r_arousal, na.rm = TRUE)
  ) %>%
  ungroup %>%
  group_by(emotion) %>% 
  mutate(median_across_subs = median(median_within_sub)) %>% 
  arrange(median_across_subs)

increasing_arousal <- rats_arousal %>%
  select(emotion, median_across_subs) %>% 
  unique() %>% 
  select(emotion) %>% pull
  
  

ggbetweenstats(
  data = rats_arousal,
  x = emotion,
  y = median_within_sub,
  # results.subtitle = FALSE,
  pairwise.display = "none",
  ggplot.component = list(scale_x_discrete(limits=increasing_arousal))
) + labs(title = "Arousal ratings")


```


## Valence
```{r, message=FALSE, warning=FALSE}

ratings_type <- "valence"
bd_ratings = paste0(bd,"/RATINGS")
ratings_path <- paste0(bd_ratings,"/",ratings_type,"_ratings.csv")
rats <- read_csv(ratings_path)

rats_valence <- rats %>% 
  select(sub, emotion, r_valence) %>% 
  # filter(sub == "02") %>% 
  arrange(emotion) %>%
  group_by(sub,emotion) %>% 
  reframe(
    median_within_sub = median(r_valence, na.rm = TRUE)
  ) %>%
  ungroup %>%
  group_by(emotion) %>% 
  mutate(median_across_subs = median(median_within_sub)) %>% 
  arrange(median_across_subs)

increasing_valence <- rats_valence %>%
  select(emotion, median_across_subs) %>% 
  unique() %>% 
  select(emotion) %>% pull
  
  

ggbetweenstats(
  data = rats_valence,
  x = emotion,
  y = median_within_sub,
  # results.subtitle = FALSE,
  pairwise.display = "none",
  ggplot.component = list(scale_x_discrete(limits=increasing_valence))
) + labs(title = "Valence ratings")


```


