---
title: "RDM MDS Explorer"
author: "LC"
date: "2024-03-16"
output: html_document
---

```{r load_libraries, message=F}
library(tidyverse)
library(future)
library(furrr)
library(tictoc)
library(RNifti)
library(proxy) # distances
library(profvis)
library(DT)
library(formattable)

source("funs_V8.R")

bd="/data00/leonardo/RSA/analyses"

# load ratings
rats <- read_csv("/data00/leonardo/RSA/analyses/RATINGS/emotion_ratings.csv")

# the following is our contrast of interest and it's always the same
copes_type <- "one_ev_per_movie"

# Vector of zeropadded sub_ids
subs_file <- "/data00/leonardo/RSA/sub_list.txt"
subs <- sprintf("%02d",readLines(subs_file) %>% as.numeric)

# ---------- ONLY TOP RATERS BELOW ------------
subs <- c("02","03","12","11","22","09","29","28","26","32","23","15","20","19")

```


# Use the rats to get - from one sub only - the label for emotion and high/low code
```{r}
labels <- rats %>% 
  filter(sub == "02") %>% 
  select(movie_cope_number,emotion, high_low_code) %>%
  mutate(high_low_code = str_extract(high_low_code,"high|low"))

labels
  
```


# Read atlas and create a vector of atlas labels
```{r}

atlas_filename <- "insula_HO_RH_LH.nii.gz"

bd_atlases = "/data00/leonardo/RSA/analyses/rsa/atlases"

atlas_path <- paste0(bd_atlases,"/",atlas_filename)

atlas_nii <- readNifti(paste0(bd_atlases,"/",atlas_filename))
region_labels <- atlas_nii[atlas_nii > 0] %>% unique %>% sort

```


# Get the path_copes
```{r}

df_path_copes <- import_df_path_copes(bd, copes_type)
copes_numba <- df_path_copes$cope %>% unique

```


Calculate fmri RDMs for a given atlas across all subs
```{r}

dist_method_fmri = "pearson"
RDM_type = "tril"  # svd or tril. Must be the same for both rating and fmri


plan(multisession, workers = 5)

tic()
RDMs_fmri <- subs %>% future_map_dfr(~{
  
  paste0("Calculating RDMs for sub ",.x,"\n") %>% cat
  
  df_copes <- load_sub_copes(.x, copes_numba, df_path_copes)
  
  one_RDM_fmri <- calculate_fmri_RDMs(
    df_copes, 
    atlas_nii, 
    dist_method_fmri, 
    RDM_type = RDM_type
  )
  
  one_RDM_fmri %>% mutate(sub = .x) %>% relocate(sub)
  
})
toc()

plan(sequential)

```


Calculate avg RDM for each region in the atlas across subs
```{r}

# subs_subset <- "02"
subs_subset <- subs

avg_RDMs_fmri <- RDMs_fmri %>% 
  
  # Choose a specific (list of) subs
  filter(sub %in% subs_subset) %>% 
  
  # Convert specific columns to long format, specifying which start with "RDM_region"
  pivot_longer(
    cols = starts_with("RDM_region"), names_to = "RDM_region_N", values_to = "value"
  ) %>% 
  
  # Group the data by 'sub' and the newly created 'RDM_region_N'
  group_by(sub, RDM_region_N) %>% 
  
  # Create a nested data frame for the 'value' column
  reframe(nested_data = list(value)) %>% 
  
  # Convert back to wide format with each 'RDM_region_N' as separate columns
  pivot_wider(names_from = RDM_region_N, values_from = nested_data) %>% 
  ungroup %>% 
  
  # Calculate the mean for each 'RDM_region' column across all 'sub' groups
  reframe(across(starts_with("RDM_region"), ~ {
    mean = list(pmap(RDM_region_1, ~ sum(...) / n()))
  }, .names = "mean_{.col}"))

avg_RDMs_fmri

```

# Select one region and recreate the original distance matrix
```{r}
insula_RH <- avg_RDMs_fmri$mean_RDM_region_1 %>% unlist
insula_LH <- avg_RDMs_fmri$mean_RDM_region_2 %>% unlist

region <- insula_RH

D_size <- labels$movie_cope_number %>% length()

D <- matrix(0, nrow = D_size, ncol = D_size)

D[lower.tri(D, diag = F)] = region

D = D + t(D)
diag(D) <- NA


# get the reordered labels
emo_labels <- labels$emotion
idx_reorded <- order(emo_labels)
emo_labels_reord <- emo_labels[idx_reorded]



plot_heatmap <- function(D_reord, emo_labels_reord) {
  # Define a blue to red color palette for the heatmap
  blue_to_red <- colorRampPalette(c("blue", "white", "red"))(n = 100)
  
  # Define the palette for emotions
  paletta_emotions <- c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3", "#FF7F00", "#FFFF33", "#A65628")
  unique_labels <- unique(emo_labels_reord)
  
  # Map each unique emotion to a color
  label_colors <- setNames(paletta_emotions, unique_labels)
  
  # Create vectors of colors for row and column side colors based on emo_labels_reord
  row_side_colors <- label_colors[emo_labels_reord]
  col_side_colors <- row_side_colors
  
  # Apply the row and column names
  rownames(D_reord) <- colnames(D_reord) <- emo_labels_reord
  
  # Use the defined color palette in the heatmap with RowSideColors and ColSideColors
  heatmap(D_reord, Rowv = NA, Colv = NA, symm = T, revC = T, col = blue_to_red, 
          RowSideColors = row_side_colors, ColSideColors = col_side_colors)
}


# Reorder cols and rows to group movies/copes with the same emotion
D_reord <- D[idx_reorded, idx_reorded]
rownames(D_reord) <- colnames(D_reord) <- emo_labels_reord

plot_heatmap(D_reord, emo_labels_reord)

```

## For arousal
```{r}
# get the reordered labels
arousal_labels <- labels$high_low_code
idx_arousal_reorded <- order(arousal_labels)
arousal_labels_reord <- arousal_labels[idx_arousal_reorded]

plot_heatmap_arousal <- function(D_reord, arousal_labels_reord) {
  # Define a blue to red color palette for the heatmap
  blue_to_red <- colorRampPalette(c("blue", "white", "red"))(n = 100)
  
  # Define the palette for emotions
  paletta_arousal <- c("#E41A1C", "#377EB8")
  unique_labels <- unique(arousal_labels_reord)
  
  # Map each unique emotion to a color
  label_colors <- setNames(paletta_arousal, unique_labels)
  
  # Create vectors of colors for row and column side colors based on emo_labels_reord
  row_side_colors <- label_colors[arousal_labels_reord]
  col_side_colors <- row_side_colors
  
  # Apply the row and column names
  rownames(D_reord) <- colnames(D_reord) <- arousal_labels_reord
  
  # Use the defined color palette in the heatmap with RowSideColors and ColSideColors
  heatmap(D_reord, Rowv = NA, Colv = NA, symm = T, revC = T, col = blue_to_red, 
          RowSideColors = row_side_colors, ColSideColors = col_side_colors)
}


# Reorder cols and rows to group movies/copes with the same emotion
D_reord <- D[idx_arousal_reorded, idx_arousal_reorded]
rownames(D_reord) <- colnames(D_reord) <- arousal_labels_reord

plot_heatmap(D_reord, arousal_labels_reord)


M <- 1 - D_reord
diag(M) = 0

mds_result <- data.frame(cmdscale(M, k = 3)) # k=2 for 2-dimensional MDS

mds_result$label <- arousal_labels_reord  

mds_result %>% 
  ggplot(aes(x = X1, y = X2, color = label)) +
  geom_point(size = 4) +
  theme_minimal() +
  labs(title = "MDS")

```




# MDS
```{r}
# Prepare the M matrix for MDS:
# take 1 - D_reord
# put 0 on the diag
M <- 1 - D_reord
diag(M) = 0

mds_result <- data.frame(cmdscale(M, k = 3)) # k=2 for 2-dimensional MDS

mds_result$label <- emo_labels_reord  

mds_result %>% 
  ggplot(aes(x = X1, y = X2, color = label)) +
  geom_point(size = 4) +
  theme_minimal() +
  labs(title = "MDS")

```


# UMAP
```{r}

library(umap)

umap_result <- umap::umap(M, input="dist", n_components=2)

lowdim_x = 1
lowdim_y = 2

umap_lowdim <- data.frame(
  X1 = umap_result$layout[,lowdim_x], 
  X2 = umap_result$layout[,lowdim_y], 
  label = emo_labels_reord
)


umap_lowdim %>% 
  ggplot(aes(x = X1, y = X2, color = label)) +
  geom_point(size = 4) +
  theme_minimal() +
  labs(title = "UMAP")

```





```{r, eval=F, include=F}
# alternative heatmap

# library(fields)
# library(viridis)
# 
# plot_heatmap <- function(M) {
#   # Assuming M is your matrix with NA on the diagonal
#   # Choose a colormap, for example, using the 'viridis' colormap from RColorBrewer
#   # If you want to use other colormaps, you can specify them here. RColorBrewer and viridisLite are good sources.
#   colormap <- viridis::viridis(100) # You can replace 'viridis' with any other desired colormap
#   
#   # Use image.plot from the fields package for plotting, with the specified colormap
#   fields::image.plot(M, col = colormap, asp=1, main = "Heatmap with Custom Colormap")
# }
# 
# plot_heatmap(D)

```






# let's redo it for one sub manually
```{r}
sub_id="02"
df_copes <- load_sub_copes(sub_id, copes_numba, df_path_copes) %>% as.matrix

atlas_nii <- readNifti(paste0(bd_atlases,"/",atlas_filename))
region_labels <- atlas_nii[atlas_nii > 0] %>% unique %>% sort

idx <- which(atlas_nii == 1)

df_copes <- df_copes[idx,]

library(proxy)
D <- cor(df_copes, method = "pearson")

dim(D)


# get the reordered labels
emo_labels <- labels$emotion
idx_reorded <- order(emo_labels)
emo_labels_reord <- emo_labels[idx_reorded]
rownames(D_reord) <- colnames(D_reord) <- emo_labels_reord

D_reord <- D[idx_reorded, idx_reorded]

plot_heatmap(D_reord)


```
































```{r, eval=FALSE, include=FALSE}
# https://stackoverflow.com/questions/54607053/dplyr-element-wise-summarise-in-nested-lists
df <- data.frame(X1 = c(1,2,2))
df$X2 <- list(list(1, 2), list(0, 1), list(1,0))
```



