---
title: "Create regions for searchlight"
author: "LC"
date: "2024-04-30"
output: html_document
---

Notes on searchlight RSA in R [here](https://jacbel.github.io/virtem_code/run-rsa-searchlight.html)

Searchlight is also in [Brainiak](https://brainiak.org/)


```{r, message=FALSE}
library(tidyverse)
library(RNifti)
library(proxy)
library(future)
library(furrr)
library(tictoc)

# Define searchlight radius in mm
r_mm <- 6

# read mask
gm <- readNifti("masks/GM_clean.nii.gz")

# voxel size of gm
voxel_size <- pixdim(gm) %>% mean

# index of nonzero voxels
idx <- which(gm != 0)

# get coordinates
coords <- which(gm != 0, arr.ind = TRUE)

```



# Define a searchlight function


## DEV version
NB: the one below is for development: three outputs are returned (distances, coordinates and indices) to carry out the checks needed to know that the function works correctly. 

The function used in production will be much simpler and just return the indices 

```{r}

searchlight_DEV <- function(center_coord, r_mm, coords, idx, voxel_size) {
  
  cat(paste0("Search radius in mm : ", r_mm))
  
  # Calculate distance of the given voxel from every other voxel, considering voxel size
  # Voxel dimensions are applied to account for the real distances in mm
  distances <- sqrt(rowSums((t(t(coords) - center_coord) * voxel_size)^2))
  
  # return the coords of the voxels in the searchlight
  sl_coords <- coords[distances <= r_mm, , drop = FALSE]

  # return only the distances <- r_mm (for test/check)
  sl_distances <- distances[distances <= r_mm]
  
  # return the idx of the voxels in the searchlight
  sl_idx <- idx[distances <= r_mm]
  
  return(list(
    sl_coords = sl_coords, 
    sl_distances = sl_distances, 
    sl_idx = sl_idx
  ))
}


# Testing the searchlight fn with one coordinate
mycoord <- coords[1,]
out <- searchlight_DEV(mycoord, r_mm=4, coords, idx, voxel_size)

# Test the max distance
out$sl_distances %>% hist

# Test the correspondence of the coordinates
cbind(
  out$sl_coords,
  coords[match(out$sl_idx, idx), , drop = FALSE]
)

```



## PROD version
```{r}
searchlight <- function(center_coord, r_mm, coords, idx, voxel_size) {

  # Calculate distance of the given voxel from every other voxel, considering voxel size
  distances <- sqrt(rowSums((t(t(coords) - center_coord) * voxel_size)^2))
  
  # # TEST ONLY: return the coords of the voxels in the searchlight
  # return(coords[distances <= r_mm, , drop = FALSE])

  # return the idx of the voxels in the searchlight
  return(idx[distances <= r_mm])
}
```






## Test parallelization
Fastest with 10-20 workers 
```{r}

plan(multisession, workers = 15)

tic()
searchlight_voxels <- 1:length(idx) %>% future_map(
  ~ searchlight(coords[.x,], r_mm, coords, idx, voxel_size)
)
toc()

plan(sequential)

searchlight_voxels %>% map_dbl(length) %>% hist

```





