---
title: "Test svd-RDM idea"
author: "LC"
date: "2024-03-11"
output: html_document
---

```{r load_libraries, message=F}
library(tidyverse)
library(future)
library(furrr)
library(tictoc)
library(RNifti)
library(proxy) # distances
library(profvis)
library(DT)
library(formattable)

source("funs_V7.R")

bd="/data00/leonardo/RSA/analyses"

# can be emotion, arousal, valence
ratings_type <- "emotion"

# the following is out contrast of interest and it's always the same
copes_type <- "one_ev_per_movie"

atlas_filename <-  "HO_cort_2mm.nii.gz"
# choices:
# HO_cort_2mm.nii.gz
# HO_sub_2mm.nii.gz
# juelich_2mm.nii.gz


# Choose the distance metric to use for fmri RDM
# The metric for ratings RDM *should be* euclidean, since arousal and valence
# ratings have only one value
# Supported methods: 'pearson', 'spearman', 'euclidean', 'cosine' or 'mahalanobis'
dist_method_rating <- "euclidean"
dist_method_fmri = "euclidean"

cat(paste0("Ratings RDM to be calculated with : ", dist_method_rating, "\n"))
cat(paste0("fMRI RDM to be calculated with : ", dist_method_fmri, "\n"))


RDM_type = "svd"  # svd or tril. Must be the same for both rating and fmri
ncomp_svd = 3

# Vector of zeropadded sub_ids
subs_file <- "/data00/leonardo/RSA/sub_list.txt"
subs <- sprintf("%02d",readLines(subs_file) %>% as.numeric)
subs <- subs[1]

```


```{r, message=FALSE}
copes_location_csv <- paste0(bd,"/rsa/copes_location.csv")

df_path_copes <- if (file.exists(copes_location_csv) ) {
  df_path_copes <- read_csv(copes_location_csv)
} else {
  df_path_copes <- import_df_path_copes(bd, copes_type)
}

copes_numba <- df_path_copes$cope %>% unique
```


# Read atlas and create a vector of atlas labels
```{r}

bd_atlases = "/data00/leonardo/RSA/analyses/rsa/atlases"

atlas_path <- paste0(bd_atlases,"/",atlas_filename)

atlas_nii <- readNifti(paste0(bd_atlases,"/",atlas_filename))
region_labels <- atlas_nii[atlas_nii > 0] %>% unique %>% sort

```


# Calculate RATINGS RDMs
We do this first since the RDM for the ratings will be the same for whatever
atlas will be used. Also, it's much faster than the RDM_fmri and can be done
for all the sub_ids at once.


- **tril** : The final RDMs_rats has 40040 nrows : 26 sub_ids * 1540, where the latter
derives from ((56^2)-56)/2 = 1540, i.e. tril from the D matrix of 56 movies
- **svd** : The final RDMs_rats has 4368 nrows : 26 sub_ids * 168, where the latter
derives from 56*3, i.e. the first three components from svd of the D matrix
```{r, message=FALSE}

bd_ratings = paste0(bd,"/RATINGS")
ratings_path <- paste0(bd_ratings,"/",ratings_type,"_ratings.csv")
rats <- read_csv(ratings_path)

# dist_method_rating <- "euclidean"
# RDM_type = "tril"  # svd or tril, see below

# for RDM_type use:
# - tril for the lower triangular of the D matrix
# - svd for the first 3 pc from svd of the D matrix
RDMs_rats_wider <- subs[1] %>% map_dfc(~ {
  do_RDM_ratings_one_sub(.x, dist_method_rating, rats, RDM_type =  RDM_type)
})

RDMs_rats <- RDMs_rats_wider %>% 
  pivot_longer(cols = everything(), names_to = "sub") %>% 
  arrange(sub)

RDMs_rats

rats_pc1 <- RDMs_rats$value[1:56]
rats_pc2 <- RDMs_rats$value[57:112]
rats_pc3 <- RDMs_rats$value[113:168]

paletta <- c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3", "#FF7F00", "#FFFF33", "#A65628")

rats_pcs <- rats %>% 
  filter(sub == subs[1]) %>% 
  select(emotion) %>% 
  mutate(pc1 = rats_pc1, pc2 = rats_pc2, pc3 = rats_pc3) 

rats_pcs %>%
  ggplot(aes(x = pc2, y = pc3, color = emotion)) +
  geom_point(shape = 16, size = 4) +  # 1 = empty, 16 = filled
    scale_color_manual(values = paletta) +
    theme_minimal()

all_rats_pcs <- c(rats_pc1, rats_pc2, rats_pc3) %>% scale
```



# Calculate FMRI RDMs

## Define the functions
1. df_copes[one_sub] <- load_sub_copes(sub_id, copes_numba)
2. RDM_fmri[one_sub] <- calculate_fmri_RDMs(df_copes, atlas_path)

```{r load_copes, eval=FALSE}

sub_id <- subs[1]

# load the copes
df_copes <- load_sub_copes(sub_id, copes_numba, df_path_copes)

# Read atlas and calculate fmri RDMs for each atlas region
dist_method_fmri = "euclidean"
cat(RDM_type)
one_RDM_fmri <- calculate_fmri_RDMs(df_copes, atlas_nii, dist_method_fmri, RDM_type = RDM_type)

one_region = "RDM_region_1"

one_RDM_fmri_one_region <- one_RDM_fmri %>% select(any_of(one_region))

fmri_pc1 <- one_RDM_fmri_one_region[[one_region]][1:56]
fmri_pc2 <- one_RDM_fmri_one_region[[one_region]][57:112]
fmri_pc3 <- one_RDM_fmri_one_region[[one_region]][113:168]

paletta <- c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3", "#FF7F00", "#FFFF33", "#A65628")

# We start with the rats just to have a vector with the emotion name.
# The copes number are *already* aligned (sorted) according to the emotion names
# in the ratings
fmri_pcs <- rats %>% 
  filter(sub == sub_id) %>% 
  select(emotion) %>% 
  mutate(pc1 = fmri_pc1, pc2 = fmri_pc2, pc3 = fmri_pc3) 

fmri_pcs %>%
  ggplot(aes(x = pc1, y = pc2, color = emotion)) +
  geom_point(shape = 16, size = 4) +  # 1 = empty, 16 = filled
    scale_color_manual(values = paletta) +
    theme_minimal()

all_fmri_pcs <- c(fmri_pc1, fmri_pc2, fmri_pc3) %>% scale

```

# try the RSA with svd for one region
```{r}

cor(all_rats_pcs, all_fmri_pcs, method = "spearman")

plot(all_rats_pcs, all_fmri_pcs)


```
































