---
title: "RSA results"
subtitle: "Inferential stats with permutation testing"
author: "Leonardo Cerliani"
date: "20/5/2023"
output:
  html_document:
    self_contained: true
    code_folding: hide
    highlight: pygments
    theme: cerulean
    css: styles.css
---

# Contrasts :

- tstat1 : RSA sig > zero

- tstat2 : RSA sig < zero

- tstat3 : covariate sig > zero

- tstat4 : covariate sig < zero


```{r, message=FALSE, echo=FALSE}
library(tidyverse)

# choose contrast here
stat_type = "_tfce_corrp_tstat1.nii.gz"

different <- ifelse(str_detect(stat_type, "tstat1|tstat3"), ">", "<")

# choose p thresh here
pthr = 0.95

```




```{r, message=FALSE}
# Libs and bd_results

# install.packages("papayaWidget")
# devtools::install_github("muschellij2/papayaWidget")

library(papayaWidget)
library(RNifti)


# papaya widget github : https://github.com/muschellij2/papayaWidget
# app code here: https://github.com/muschellij2/linked_viewer/blob/master/app.R

bd_results = "/data00/leonardo/RSA/analyses/rsa_emotion_high_low_predictors/rsa_results_INITIAL_3_MODELS"

```


```{r, message=FALSE}
# Some defs/fns to make strings to pass to Papaya concise

# define the MNI and the Dummy (for transparency of the MNI)
MNI=paste0(bd_results,"/","MNI152_T1_2mm_brain.nii.gz")
Dummy <- paste0(bd_results,"/","Dummy.nii.gz")

# this function takes the contrast name and the params above to
# build the pathname of the contrast of interest to be fed
# into papaya
#
# E.g: f("Arousal"); f("Arousal_vs_Emotion")
f <- function(contrast_name) {
  template <- paste0(bd,"/stats_CONTRAST",stat_type)
  contrast_path <- str_replace(template, "CONTRAST", contrast_name)
  return(contrast_path)
}
```


# N14`r stat_type`
```{r, message=FALSE}
flavour = "N14_GM_clean_bilat"
bd=paste0(bd_results,"/",flavour)
```

## N14 Arousal, Emotion, Valence MEAN
- Arousal : Red  
- Emotion : Blue
- Valence : Green
```{r}

# simial to the f above, but much simpler, only for the median
fm <- function(contrast_name) {
  return(paste0(bd,"_",contrast_name,".nii.gz"))
}

papaya(
  c(Dummy, MNI, fm("arousal"), fm("emotion"), fm("valence")),
  options = list(
    papayaOptions(alpha = 1, lut = "Grayscale"),
    papayaOptions(alpha = 0.5, lut = "Grayscale"),
    papayaOptions(alpha = 1, lut = "Red Overlay", min = 0.25, max = 0.35),
    papayaOptions(alpha = 1, lut = "Blue Overlay", min = -0.1, max = -0.2),
    papayaOptions(alpha = 1, lut = "Green Overlay", min = 0.15, max = 0.2)
  ),
  interpolation = FALSE
)

```


<br><br>


## N14 Arousal, Emotion, Valence sig `r different` zero
- Arousal : Red
- Emotion : Blue
- Valence : Green
```{r, message=FALSE}

papaya(
  c(Dummy, MNI, f("Arousal"), f("Emotion"), f("Valence")),
  options = list(
    papayaOptions(alpha = 1, lut = "Grayscale"),
    papayaOptions(alpha = 0.5, lut = "Grayscale"),
    papayaOptions(alpha = 1, lut = "Red Overlay", min = pthr, max = 1),
    papayaOptions(alpha = 1, lut = "Blue Overlay", min = pthr, max = 1),
    papayaOptions(alpha = 1, lut = "Green Overlay", min = pthr, max = 1)
  ),
  interpolation = FALSE
)

```

<br><br>

## N14 Arousal sig `r different` Emotion/Valence
- Arousal `r different` Emotion : Red
- Arousal `r different` Valence : Blue
```{r}

papaya(
  c(Dummy, MNI, f("Arousal_vs_Emotion"), f("Arousal_vs_Valence")),
    options = list(
    papayaOptions(alpha = 1, lut = "Grayscale"),
    papayaOptions(alpha = 0.5, lut = "Grayscale"),
    papayaOptions(alpha = 1, lut = "Red Overlay", min = pthr, max = 1),
    papayaOptions(alpha = 1, lut = "Blue Overlay", min = pthr, max = 1)
  ),
  interpolation = FALSE
)

```

<br><br>

## N14 Emotion sig `r different` Arousal/Valence
- Emotion `r different` Arousal : Red
- Emotion `r different` Valence : Blue
```{r}

papaya(
  c(Dummy, MNI, f("Emotion_vs_Arousal"), f("Emotion_vs_Valence")),
    options = list(
    papayaOptions(alpha = 1, lut = "Grayscale"),
    papayaOptions(alpha = 0.5, lut = "Grayscale"),
    papayaOptions(alpha = 1, lut = "Red Overlay", min = pthr, max = 1),
    papayaOptions(alpha = 1, lut = "Blue Overlay", min = pthr, max = 1)
  ),
  interpolation = FALSE
)

```



<br><br>

## N14 Valence sig `r different` Emotion/Arousal
- Valence `r different` Arousal : Red
- Valence `r different` Emotion : Blue
```{r}

papaya(
  c(Dummy, MNI, f("Valence_vs_Arousal"), f("Valence_vs_Emotion")),
    options = list(
    papayaOptions(alpha = 1, lut = "Grayscale"),
    papayaOptions(alpha = 0.5, lut = "Grayscale"),
    papayaOptions(alpha = 1, lut = "Red Overlay", min = pthr, max = 1),
    papayaOptions(alpha = 1, lut = "Blue Overlay", min = pthr, max = 1)
  ),
  interpolation = FALSE
)

```














<br><br>

# N26`r stat_type`
```{r, message=FALSE}
flavour = "N26_GM_clean_bilat"
bd=paste0(bd_results,"/",flavour)
```


## N26 Arousal, Emotion, Valence MEAN
- Arousal : Red
- Emotion : Blue
- Valence : Green
```{r}

# simial to the f above, but much simpler, only for the median
fm <- function(contrast_name) {
  return(paste0(bd,"_",contrast_name,".nii.gz"))
}

papaya(
  c(Dummy, MNI, fm("arousal"), fm("emotion"), fm("valence")),
  options = list(
    papayaOptions(alpha = 1, lut = "Grayscale"),
    papayaOptions(alpha = 0.5, lut = "Grayscale"),
    papayaOptions(alpha = 1, lut = "Red Overlay", min = 0.25, max = 0.35),
    papayaOptions(alpha = 1, lut = "Blue Overlay", min = -0.1, max = -0.2),
    papayaOptions(alpha = 1, lut = "Green Overlay", min = 0.15, max = 0.2)
  ),
  interpolation = FALSE
)

```


<br><br>

## N26 Arousal, Emotion, Valence sig `r different` from zero
- Arousal : Red
- Emotion : Blue
- Valence : Green
```{r, message=FALSE}

papaya(
  c(Dummy, MNI, f("Arousal"), f("Emotion"), f("Valence")),
  options = list(
    papayaOptions(alpha = 1, lut = "Grayscale"),
    papayaOptions(alpha = 0.5, lut = "Grayscale"),
    papayaOptions(alpha = 1, lut = "Red Overlay", min = pthr, max = 1),
    papayaOptions(alpha = 1, lut = "Blue Overlay", min = pthr, max = 1),
    papayaOptions(alpha = 1, lut = "Green Overlay", min = pthr, max = 1)
  ),
  interpolation = FALSE
)

```

<br><br>

## N26 Arousal sig `r different` Emotion/Valence
- Arousal `r different` Emotion : Red
- Arousal `r different` Valence : Blue
```{r}

papaya(
  c(Dummy, MNI, f("Arousal_vs_Emotion"), f("Arousal_vs_Valence")),
    options = list(
    papayaOptions(alpha = 1, lut = "Grayscale"),
    papayaOptions(alpha = 0.5, lut = "Grayscale"),
    papayaOptions(alpha = 1, lut = "Red Overlay", min = pthr, max = 1),
    papayaOptions(alpha = 1, lut = "Blue Overlay", min = pthr, max = 1)
  ),
  interpolation = FALSE
)

```

<br><br>

## N26 Emotion sig `r different` Arousal/Valence
- Emotion `r different` Arousal : Red
- Emotion `r different` Valence : Blue
```{r}

papaya(
  c(Dummy, MNI, f("Emotion_vs_Arousal"), f("Emotion_vs_Valence")),
    options = list(
    papayaOptions(alpha = 1, lut = "Grayscale"),
    papayaOptions(alpha = 0.5, lut = "Grayscale"),
    papayaOptions(alpha = 1, lut = "Red Overlay", min = pthr, max = 1),
    papayaOptions(alpha = 1, lut = "Blue Overlay", min = pthr, max = 1)
  ),
  interpolation = FALSE
)

```



<br><br>

## N26 Valence sig `r different` Emotion/Arousal
- Valence `r different` Arousal : Red
- Valence `r different` Emotion : Blue
```{r}

papaya(
  c(Dummy, MNI, f("Valence_vs_Arousal"), f("Valence_vs_Emotion")),
    options = list(
    papayaOptions(alpha = 1, lut = "Grayscale"),
    papayaOptions(alpha = 0.5, lut = "Grayscale"),
    papayaOptions(alpha = 1, lut = "Red Overlay", min = pthr, max = 1),
    papayaOptions(alpha = 1, lut = "Blue Overlay", min = pthr, max = 1)
  ),
  interpolation = FALSE
)

```






















