---
title: "Papaya Coordinates Headache"
author: "LC"
date: "2024-05-25"
output: html_document
---

# Motivation
The [Papaya Widget by Jon Muschelli](https://johnmuschelli.com/papayaWidget/) 
is a wonderful thing which allows storing image browsers inside a self-contained
HMTL generated from an Rmarkdown document.

Unfortunately, the voxel coordinates displayed in Papaya are totally confusing.
So if you want e.g. to build a searchlight mask around a specific coordinate, 
and take as reference the one displayed by the papaya widget, you will find
your mask lying in a totally different location.

After one full day of suffering, below you can find the way to go from
Papaya coordinates to actual coordinates (e.g. those displayed in fslview).
Do not expect a full explanation. Some of the things working below are just 
the result of trial and error.


```{r, message=FALSE}
library(tidyverse)
library(RNifti)
library(papayaWidget)

```




If you want to write values in specific coordinates, you can first retrieve
the linear indices of the volume. 

```{r}
MNI <- readNifti("MNI152_T1_2mm_brain.nii.gz")

# Transform the coordinates displayed in papaya (p_coord) 
# first into real coordinates (coord) and then into linear indices
coord_2_ind <- function(p_coord, MNI) {

  p_coord = p_coord + 1
  center_MNI <- ceiling((dim(MNI) / 2) - 1) + 1
  
  # no idea what happened in the widget, but this does the conversion
  coord <- 2*center_MNI - p_coord
  
  # coordinates to linear indices
  dims <- dim(MNI)
  ind <- coord[1] + (coord[2] - 1) * dims[1] + (coord[3] - 1) * dims[1] * dims[2]
  return(ind)
}

# E.g. with the volume center (NOT origin)
coord <- c(45,54,45)
ind <- coord_2_ind(coord, MNI)
MNI[ind]
```


If you want to do something around a specific coordinate - e.g. building a
searchlight - it might be better to first convert the coordinate of the center
and then work with the coordinates from the volume indices (those are correct)

```{r}

# Papaya to real coordinates
# The p_coord is the voxel coordinate displayed in Papaya
pap_2_real_coord <- function(p_coord, MNI) {
  p_coord = p_coord + 1
  center_MNI <- ceiling((dim(MNI) / 2) - 1) + 1
  return(2*center_MNI - p_coord)
}


# Searchlight function : returns the linear indices of the voxels in the searchlight
#
# NB: the coordinate of the ROI center of gravity (COG)
# must already be the actual one, NOT the one displayed by Papaya!
searchlight <- function(ROI_COG, r_mm, MNI) {
  
  voxel_size <- pixdim(MNI) %>% mean 
  ind <- which(MNI != 0)
  coords <- which(MNI != 0, arr.ind = TRUE)

  # Calculate distance of the given voxel from every other voxel, considering voxel size
  distances <- sqrt(rowSums((t(t(coords) - ROI_COG) * voxel_size)^2))

  # return the idx of the voxels in the searchlight
  return(ind[distances <= r_mm])
}


```




E.g. creating a lightsearch ball around the coordinate that Papaya
displays as 28,15,45 - which is actually 63,91,46 as in the ROI_COG below

```{r}

p_coord <- c(28,18,45)
cat("Papaya coords : ",p_coord,"\n")

ROI_COG <- pap_2_real_coord(p_coord,MNI)
cat("Real coords : ",ROI_COG,"\n")


ind_ball <- searchlight(ROI_COG, 4, MNI)


# Write the nii
nii <- MNI * 0
nii[ind_ball] = 1
writeNifti(nii, "ROI.nii.gz")

papaya(
  c("MNI152_T1_2mm_brain.nii.gz","ROI.nii.gz"),
  options = list(
    papayaOptions(alpha = 1, lut = "Grayscale"),
    papayaOptions(alpha = 1, lut = "Red Overlay", min = 0, max = 2)
  ), interpolation = FALSE
)

```























