---
title: "RSA ROI analysis"
subtitle: "rating and fmri RDMs for a given roi.nii.gz"
author: "Leonardo Cerliani"
date: "23/5/2023"
output:
  html_document:
    self_contained: true
    code_folding: hide
    highlight: pygments
    theme: cerulean
    css: styles.css
---


# Initialization

## Load libraries
```{r message=F}
library(tidyverse)
library(future)
library(furrr)
library(tictoc)
library(RNifti)
library(proxy) # distances
library(profvis)
library(DT)
library(formattable)
library(gplots) # heatmap.2
library(umap)
library(patchwork)


source("funs_V8_ROIs.R")

bd="/data00/leonardo/RSA/analyses"
bd_ratings = paste0(bd,"/RATINGS")

# Manually modify according to the rsa flavour
# rsa                     :   one_ev_per_movie
# rsa_emotion_predictors  :   emotion_predictors
# rsa_emotion_high_low_predictors   :   emotion_high_low_predictors
rsa_flavour <- "rsa_emotion_predictors_minus_neutral_NN"
copes_type <- "emotion_predictors_minus_neutral"


# Choose the distance metric to use for fmri RDM
# Supported methods: 'pearson', 'spearman', 'euclidean', 'cosine' or 'mahalanobis'
dist_method_rating <- "euclidean"
dist_method_fmri = "euclidean"
dist_method_rsa = "pearson"


bd_ROI <- paste0("/data00/leonardo/RSA/analyses/",rsa_flavour,"/rsa_viz/")
ROI_filename <- "ROI.nii.gz"


```


## Read in the ROI(s) for which you want to calculate RDMs
NB: there can be multiple ROIs in one nii

```{r}

ROI_nii <- readNifti(paste0(bd_ROI, "/", ROI_filename))
# view(ROI_nii)

# just in case there is more than one label (with ROI.nii.gz it is not the case)
region_labels <- ROI_nii[ROI_nii > 0] %>% unique %>% sort

```


## Define subs set
```{r message=F}

# Vector of zeropadded sub_ids
subs_file <- "/data00/leonardo/RSA/sub_list.txt"
subs <- sprintf("%02d",readLines(subs_file) %>% as.numeric)

# ---------- ONLY TOP RATERS BELOW ------------
subs <- c("02","03","12","11","22","09","29","28","26","32","23","15","20","19")

# load emotion labels
emotion_labels <- read_csv(paste0(bd,"/RATINGS/emotion_ratings.csv")) %>% 
  filter(sub == "02") %>% select(emotion) %>% pull

```


## Heatmap plot functions
```{r}
# plot_heatmap(D %>% as.matrix)
plot_heatmap <- function(M, title=NULL) {
  heatmap(M, Rowv = NA, Colv = NA, symm=T, revC = T, main=title)
}

# The following is to plot also side colors for emotion
# NB:   # The input matrix MUST already have rownames for emotions
plot_heatmap_sides <- function(M, title=NULL) {
  
  emo_names <- rownames(M)
  
  # Define a blue to red color palette for the heatmap
  blue_to_red <- colorRampPalette(c("blue", "white", "red"))(n = 100)
  
  # Define the palette for emotions
  paletta_emotions <- c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3", "#FF7F00", "#FFFF33", "#FFFFFF")
  emo_labels <-       c("anger",   "pain",    "disgust", "sad",     "fear",    "happy",   "neutral")
  
  # Map each unique emotion to a color
  label_colors <- setNames(paletta_emotions, emo_labels)
  
  # Create vectors of colors for row and column side colors based on emo_labels_reord
  row_side_colors <- col_side_colors <- label_colors[emo_names]
  
  # Plot the heatmap
  heatmap.2(M, Rowv = NA, Colv = NA, symm=T, revC = F, col = blue_to_red,
            RowSideColors = row_side_colors, ColSideColors = col_side_colors, 
            main = title, trace='none', cellnote=round(M, 2), notecol="black", notecex=1.3, 
            dendrogram='none', margins=c(10,10), key = TRUE)
}


```


# Rats RDM

## Function to calculate model-specific mean_rats RDM matrix (tril of the median RDM)
```{r, message=FALSE}

get_rats_median_RDM <- function(ratings_type, dist_method_rating) {

  bd_ratings = paste0(bd,"/RATINGS")
  ratings_path <- paste0(bd_ratings,"/",ratings_type,"_ratings.csv")
  rats <- read_csv(ratings_path)
  
  
  rats <- rats %>% 
    filter(sub %in% subs) %>%
    select(sub, emotion, starts_with("r_")) %>% 
    # IMPORTANT : AVG ACROSS EMOTION (8 movies for each emotion) FOR EMOTION PREDICTORS MODEL
    group_by(sub, emotion) %>% 
    reframe(across(starts_with("r_"), median, na.rm = TRUE)) %>% 
    ungroup %>% 
    arrange(sub, emotion) 
  
  
  # get rats_median
  rats_median <- rats %>% 
    group_by(emotion) %>% 
    summarise(
      emotion = max(emotion),
      across(starts_with("r_"), ~ round(median(.x, na.rm = TRUE), 2))
    ) 
  
  # calculate median RDM
  # NB: this is the median of the sub-specific RDM, 
  # *not* the RDM of the median rats (for concordance with median fMRI RDM)
  RDMs_rats_wider <- subs %>% map_dfc(~ {
    do_RDM_ratings_one_sub(.x, dist_method_rating, rats, RDM_type =  'tril')
  })
  
  RDMs_rats <- RDMs_rats_wider %>% 
    pivot_longer(cols = everything(), names_to = "sub") %>% 
    arrange(sub)
  
  # average across subs - for this we need to group across subs
  # and add a tril_element
  df_median_RDMs_rats <- RDMs_rats %>%
    group_by(sub) %>% 
    mutate(tril_element = row_number()) %>%
    ungroup %>% 
    group_by(tril_element) %>% 
    reframe(median_value = median(value, na.rm = TRUE))
  
  # calculate the number of elements on each side of the matrix (careful here)
  dim_matrix <- rats$emotion %>% unique %>% length
  
  m = matrix(0, dim_matrix, dim_matrix)
  m[lower.tri(m)] <- df_median_RDMs_rats$median_value
  m <- m + t(m)
  diag(m) <- NA

  rownames(m) <- rats_median$emotion
  colnames(m) <- rats_median$emotion
  
  median_RDMs_rats <- m
  
    
  # get the reordering vector for different emotion 
  #  OR increasing arousal OR increasing valence
  idx_sort <- switch(
    ratings_type,
    "emotion" = order(rats_median$emotion),
    "arousal" = order(rats_median$r_arousal),
    "valence" = order(rats_median$r_valence)
  )
  
  # just to check
  rats_median[idx_sort,]
  
  return(
    list(idx_sort, rats_median, median_RDMs_rats)
  )
  
}


```


## Calculate and Plot the reordered median rats RDM for all three models
The idx_sort_[model] will be used to reorder the fmri_RDM
```{r, message=FALSE}
# AROUSAL
out_arousal <- get_rats_median_RDM("arousal", dist_method_rating)
idx_sort_arousal <- out_arousal[[1]]
rats_median_arousal <- out_arousal[[2]]
RDM_arousal <- out_arousal[[3]]


# EMOTION
out_emotion <- get_rats_median_RDM("emotion", dist_method_rating)
idx_sort_emotion <- out_emotion[[1]]
rats_median_emotion <- out_emotion[[2]]
RDM_emotion <- out_emotion[[3]]


# VALENCE
out_valence <- get_rats_median_RDM("valence", dist_method_rating)
idx_sort_valence <- out_valence[[1]]
rats_median_valence <- out_valence[[2]]
RDM_valence <- out_valence[[3]]


# AROVAL
out_aroval <- get_rats_median_RDM("aroval", dist_method_rating)
idx_sort_aroval <- out_aroval[[1]]
rats_median_aroval <- out_aroval[[2]]
RDM_aroval <- out_aroval[[3]]


# plot_heatmap_sides(RDM_emotion[idx_sort_emotion,idx_sort_emotion], "emotion") 
# plot_heatmap_sides(RDM_arousal[idx_sort_arousal,idx_sort_arousal], "arousal")
# plot_heatmap_sides(RDM_valence[idx_sort_valence,idx_sort_valence], "valence")

plot_heatmap_sides(RDM_emotion, "emotion") 
plot_heatmap_sides(RDM_arousal, "arousal")
plot_heatmap_sides(RDM_valence, "valence")
plot_heatmap_sides(RDM_aroval, "aroval")
```


## Alternative method : RDM of the median tril
```{r, message=FALSE}
# Just to check, here are the median RDM calculated on the median tril

m_arousal <- rats_median_arousal %>% 
  select(starts_with("r_")) %>% 
  as.matrix %>% 
  DDOS(method = "euclidean") %>% as.matrix

rownames(m_arousal) <- colnames(m_arousal) <- rownames(RDM_arousal)
diag(m_arousal) <- NA
plot_heatmap_sides(m_arousal, "arousal")


m_emotion <- rats_median_emotion %>% 
  select(starts_with("r_")) %>% 
  as.matrix %>% 
  DDOS(method = "euclidean") %>% as.matrix

rownames(m_emotion) <- colnames(m_emotion) <- rownames(RDM_arousal)
diag(m_emotion) <- NA
plot_heatmap_sides(m_emotion, "emotion")


```



## UMAP plot
```{r, eval=TRUE, include=TRUE, message=FALSE}
# RDM_ratings dimred



umap_result <- umap::umap(RDM_emotion, input="dist", n_neighbors = 3, n_components=2)

paletta_emotions <- c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3", "#FF7F00", "#FFFF33", "#FFFFFF")

# plot pc1,2 and pc2,3
do_UMAP_lowdim_plots <- function(umap_result, lowdim_x, lowdim_y) {
  
  umap_lowdim <- data.frame(
    pc_x = umap_result$layout[,lowdim_x], 
    pc_y = umap_result$layout[,lowdim_y], 
    emotion_labels
  )
  
  p <- ggplot(umap_lowdim, aes(x = pc_x, y = pc_y, color = emotion_labels)) +
    geom_point(size = 3, alpha = 0.7) +  # Open circles for points
    scale_color_manual(values = paletta_emotions) +
    scale_shape_manual(values = c(1, 3)) +  # Assign specific shapes
    labs(
      title = "UMAP Projection", 
      x = paste0("lowdim ",lowdim_x), 
      y = paste0("lowdim ",lowdim_y)
    ) +
    theme_minimal() +
    guides(color = guide_legend(override.aes = list(shape = 16)))  # Filled circles in the legend
  
  return(p)
}


p <- do_UMAP_lowdim_plots(umap_result, lowdim_x = 1, lowdim_y = 2)
p

```






# fmri RDM

## Get copes location
NB: The cope numbers in the `cope` column are NOT zeropadded since this is how they come out from FSL Feat

```{r,message=FALSE}

copes_location_csv <- paste0(bd,"/",rsa_flavour,"/copes_location.csv")

df_path_copes <- if (file.exists(copes_location_csv) ) {
  df_path_copes <- read_csv(copes_location_csv)
} else {
  df_path_copes <- import_df_path_copes(bd, copes_type)
}

copes_numba <- df_path_copes$cope %>% unique
ncopes <- length(copes_numba)

```



## Calculate RDMs
```{r}

# 5 workers give the best performance
plan(multisession, workers = 5)

tic()
RDMs_fmri <- subs %>% future_map_dfr(~{
  
  paste0("Calculating RDMs for sub ",.x,"\n") %>% cat
  
  df_copes <- load_sub_copes(.x, copes_numba, df_path_copes)
  
  one_RDM_fmri <- calculate_fmri_RDMs(
    df_copes,
    ROI_nii,
    dist_method_fmri,
    RDM_type = "tril"
  )

  one_RDM_fmri %>% mutate(sub = .x) %>% relocate(sub)
  
})
toc()


plan(sequential)

```



## Calculate median triu and plot full
```{r}

mean_triu <- RDMs_fmri %>% 
  group_by(sub) %>% 
  mutate(triu_cell = row_number()) %>% 
  ungroup() %>% 
  group_by(triu_cell) %>% 
  summarise(mean_triu_cell = median(RDM_region_1)) %>% 
  select(mean_triu_cell) %>% pull

# mean_triu
hist(mean_triu)


# reconstruct full matrix and assign row/colnames
reconstruct_full <- function(triu, ncopes, emotion_labels) {
  RDM_full <- matrix(0, nrow = ncopes, ncol = ncopes)
  RDM_full[lower.tri(RDM_full, diag = F)] = triu
  RDM_full = RDM_full + t(RDM_full)
  diag(RDM_full) <- NA
  rownames(RDM_full) <- colnames(RDM_full) <- rownames(RDM_emotion)  # careful here!!
  return(RDM_full)
}

RDM_full <- reconstruct_full(mean_triu, ncopes, emotion_labels)

# plot_heatmap_sides(RDM_full[idx_sort_arousal, idx_sort_arousal], "arousal")
# plot_heatmap_sides(RDM_full[idx_sort_emotion, idx_sort_emotion], "emotion")
# plot_heatmap_sides(RDM_full[idx_sort_valence, idx_sort_valence], "valence")

plot_heatmap_sides(RDM_full, "fmri")

```


## Scatterplot rats fmri
```{r}

triu_fmri <- RDM_full[lower.tri(RDM_full)]

triu_arousal <- RDM_arousal[lower.tri(RDM_arousal)]
triu_emotion <- RDM_emotion[lower.tri(RDM_emotion)]

cor(triu_fmri, triu_arousal, method = "spearman")
plot(triu_fmri, triu_arousal, main = "fmri vs arousal")

cor(triu_fmri, triu_emotion, method = "spearman")
plot(triu_fmri, triu_emotion, main = "fmri vs emotion")


```





## Test for distances
```{r, include=FALSE, eval=FALSE}
emo_labels <- rats_median_emotion$emotion

# Arousal
m_arousal <- rats_median_arousal %>% 
  select(starts_with("r_")) %>% 
  as.matrix 

rownames(m_arousal) <- emo_labels
m_arousal

D_arousal <- dist(m_arousal) %>% as.matrix
diag(D_arousal) <- NA
rownames(D_arousal) <- colnames(D_arousal) <- emo_labels
plot_heatmap_sides(D_arousal, "arousal")


# Emotion
m_emotion <- rats_median_emotion %>% 
  select(starts_with("r_")) %>% 
  as.matrix 

rownames(m_emotion) <- emo_labels
m_emotion


D_emotion <- dist(m_emotion) %>% as.matrix
diag(D_emotion) <- NA
rownames(D_emotion) <- colnames(D_emotion) <- emo_labels
plot_heatmap_sides(D_emotion, "emotion")
```




## Difference between unidimensional arousal and multidimensional emotion rating

```{r, include=T, eval=T}
v1 = 1
v5 = 5
v6 = 6

all <- rbind(v1,v5,v6)

D <- dist(all) %>% as.matrix %>% round()
diag(D) = NA

heatmap.2(D, Rowv = NA, Colv = NA, symm=T, revC = F, dendrogram = 'none',
          cellnote = D, notecol = 'black',notecex = 1.0, trace = 'none', key = F)

```


```{r, include=T, eval=T}
v1 = c(1,0,0)
v5 = c(0,5,0)
v6 = c(0,0,6)

all <- rbind(v1,v5,v6)

D <- dist(all) %>% as.matrix %>% round()
diag(D) = NA


heatmap.2(D, Rowv = NA, Colv = NA, symm=T, revC = F, dendrogram = 'none',
          cellnote = D, notecol = 'black',notecex = 1.0, trace = 'none', key = F)
```































