---
title: "sketches"
author: "LC"
date: "2025-10-16"
output: html_document
---

NB: this script uses the `do_RSA_tables_pdf_FUNCTIONS.R`

The `Phylogenetic_tree_Yeo7_ROI[5/4].csv` used by `create_roi_filter()` come from
the Phylogenetic ROI tree created in `ROIS_REPO`


```{r message=FALSE}
library(gt)
library(tidyverse)
library(BayesFactor)
library(webshot2)  # for saving as PDF
library(stringr)
library(tictoc)


source("do_RSA_tables_FUNCTIONS.R")

# Set root and output folder
root <- "/data00/leonardo/RSA/analyses/RSA_ROI_APP/results_RSA_ROI"
tables_dir <- file.path(root, "Tables")

# Create Output Tables folder if it doesn't exist
if (!dir.exists(tables_dir)) dir.create(tables_dir)


ROI_Yeo7_number = 4  # OFC=5  vATTN=4


if (ROI_Yeo7_number == 4) {
  atlas_list <- c("Yeo7", "Yeo17", "HO_cort", "anatomy_toolbox", "AT_insula")
} else {
  atlas_list <- c("Yeo7", "Yeo17", "HO_cort", "anatomy_toolbox")
}

```


# V1 - One atlas at the time table 
NB: saving is disabled since it fails and this code is just to check a basic version of the output
```{r message=FALSE}

# atlas_name <- "Yeo7"
# atlas_name <- "Yeo17"
# atlas_name <- "HO_cort"
# atlas_name <- "anatomy_toolbox"
atlas_name <- "AT_insula"


# ------- end of user-defined parameters

# Make a list of all the RData results for one atlas
RData_files <- make_rdata_list(atlas_name)

# Run all the ttest and ttestBF and report the results
results <- RData_files %>% 
  map_dfr(~ rsa_stats_table(.x$file, .x$atlas, .x$label)) %>% 
  arrange(roi)

# Return the list of the regions to examine for that atlas_name and choice of ROI_Yeo7_number
atlas_roi_filter <- if (atlas_name != "AT_insula") {
  create_roi_filter(atlas_name, ROI_Yeo7_number)
} else {
  NULL  # or `character(0)` if you prefer an empty vector
}

# Produces the gt_table for the desired ROIs for that atlas
gt_table <- make_rsa_gt_table_colored(results, roi_filter = atlas_roi_filter)
gt_table

# # Saves it into a pdf file
# gtsave(gt_table, paste0(atlas_name, "_rsa_table.pdf"))


```



# V2 - All atlases at once for each of Yeo7 ROI 4 or 5
There is a main header for each atlas, however for atlases with many regions
it is very difficult to read the header for ROIs down the line.
```{r, message=FALSE}

tic("Full RSA table generation")

# # Just for testing - atlas list defined in the first chunk
# atlas_list <- c("Yeo7", "Yeo17")


Yeo7_ROI <- paste0("Yeo", ROI_Yeo7_number)

# Create the output directory if it doesn't exist
if (!dir.exists("./Tables")) dir.create("./Tables")

# make all tables
all_gt_tables <- map(atlas_list, function(atlas) {
  RData_files <- make_rdata_list(atlas)
  results <- RData_files %>%
    map_dfr(~ rsa_stats_table(.x$file, .x$atlas, .x$label)) %>%
    arrange(roi)
  
  if (atlas != "AT_insula") {
    roi_filter <- create_roi_filter(atlas, ROI_Yeo7_number)
  } else {
    roi_filter <- NULL  # or character(0) if you prefer
  }
  
  make_rsa_gt_table_colored(results, roi_filter = roi_filter)
})


# save each table to a temporary HTML inside ./Tables
tmp_files <- map2(all_gt_tables, atlas_list, function(tab, name) {
  tmp <- file.path("./Tables", paste0(name, "_tmp.html"))
  gtsave(tab, tmp, inline_css = TRUE)
  tmp
})

# merge all HTMLs into one
html_all <- map_chr(tmp_files, readr::read_file) %>%
  paste(collapse = "<div style='page-break-after:always;'></div>")

# Write combined HTML into ./Tables
final_html <- file.path("./Tables", paste0("combined_RSA_tables_", Yeo7_ROI, ".html"))
writeLines(html_all, final_html)

# Convert HTML to PDF inside ./Tables
final_pdf <- file.path("./Tables", paste0("combined_RSA_tables_", Yeo7_ROI, ".pdf"))
webshot(final_html, final_pdf, vwidth = 1200)

# Remove temporary files
tmp_files <- unlist(tmp_files)
tmp_files <- tmp_files[file.exists(tmp_files)]
file.remove(tmp_files)

toc()

```



# V3 - Single ROI tables and then merge them into HTML or pdf, to have headers for each ROI
Now there is no main title, instead all the ROIs are treated as single tables
and then merged into a single HTML / PDF file. In this way there is a header
for each ROI
```{r, message=FALSE}

# # Just for testing - atlas list defined in the first chunk
# atlas_list <- c("Yeo7", "Yeo17")

# make all tables
all_gt_tables <- map(atlas_list, function(atlas) {
  RData_files <- make_rdata_list(atlas)
  results <- RData_files %>%
    map_dfr(~ rsa_stats_table(.x$file, .x$atlas, .x$label)) %>%
    arrange(roi)
  
  if (atlas != "AT_insula") {
    roi_filter <- create_roi_filter(atlas, ROI_Yeo7_number)
  } else {
    roi_filter <- NULL  # or character(0) if you prefer
  }
  
  # returns a list of gt tables per ROI
  make_rsa_gt_table_colored_single_ROI(results, roi_filter = roi_filter)
})

# Flatten the list (all tables across all atlases)
all_gt_tables <- unlist(all_gt_tables, recursive = FALSE)

# Save each table to a temporary HTML inside ./Tables
tmp_files <- map2(all_gt_tables, seq_along(all_gt_tables), function(tab, idx) {
  tmp <- file.path("./Tables", paste0("tmp_table_", idx, ".html"))
  gtsave(tab, tmp, inline_css = TRUE)
  tmp
})

# Convert tmp_files list to character vector and keep only existing files
tmp_files <- unlist(tmp_files)
tmp_files <- tmp_files[file.exists(tmp_files)]

# Merge all HTMLs into one
html_all <- map_chr(tmp_files, readr::read_file) %>%
  paste(collapse = "<div style='page-break-after:always;'></div>")

# Save combined HTML
Yeo7_ROI <- paste0("Yeo", ROI_Yeo7_number)

final_html <- file.path("./Tables", paste0("combined_RSA_tables_all_headers_", Yeo7_ROI, ".html"))
writeLines(html_all, final_html)

# Convert to PDF
final_pdf <- file.path("./Tables", paste0("combined_RSA_tables_all_headers_", Yeo7_ROI, ".pdf"))
webshot(final_html, final_pdf, vwidth = 1200)

# Remove temporary files safely
file.remove(tmp_files)
```




# V4 - Reactable table in a standalone HTML
Stand-alone HTML in Reactable so that one can filter for atlas, ROI number and 
RSA flavour.

NB: the reactable HTML comes with a folder, but it's not necessary to use the
ordering and filtering widgets
```{r, message=FALSE}
library(tidyverse)
library(stringr)
library(reactable)
library(htmlwidgets)
library(htmltools)  # for tags

# # Just for testing - atlas list defined in the first chunk
# atlas_list <- c("Yeo7", "Yeo17")
# atlas_list <- c("AT_insula")

# Combine all results into a plain data frame
df <- map_dfr(atlas_list, function(atlas) {
  RData_files <- make_rdata_list(atlas)
  
  results <- RData_files %>%
    map_dfr(~ rsa_stats_table(.x$file, .x$atlas, .x$label)) %>%
    arrange(roi)
  
  if (atlas != "AT_insula") {
    roi_filter <- create_roi_filter(atlas, ROI_Yeo7_number)
    results <- results %>% filter(roi %in% roi_filter)
  }
  
  results %>% mutate(atlas = atlas, roi_group = paste0("ROI ", roi))
}) %>%
  select(atlas, roi_group, label, everything(), -roi) %>%
  mutate(across(everything(), as.character))  # ensure all columns are simple text

# Create reactable
react_tbl <- reactable(
  df,
  filterable = TRUE,         # adds a textbox filter to each column
  bordered = FALSE,
  striped = TRUE,
  highlight = TRUE,
  compact = TRUE,
  defaultPageSize = 20,
  showPageSizeOptions = TRUE,
  style = list(fontFamily = "Arial Narrow")  # set font
)

react_tbl

# Save as standalone HTML inside ./Tables
output_file <- file.path(tables_dir, paste0("rsa_reactable_Yeo", ROI_Yeo7_number, ".html"))
saveWidget(
  react_tbl,
  output_file,
  selfcontained = TRUE
)

```










# ---------------------------- IMPORTANT -------------------------------------

**THE CHUNKS BELOW NEED NOT TO BE RUN.**

They refer to the development stage. Now all these functions are stored 
in `do_RSA_tables_pdf_FUNCTIONS.R`. The pdf is generated already at the end
of the previous chunk.


# Return the list of the regions to examine for that atlas_name and choice of ROI_Yeo7_number
```{r, message=FALSE}
create_roi_filter <- function(atlas_name, ROI_Yeo7_number) {
 ROIs_phylo_tree <- read_csv(paste0("Phylogenetic_tree_Yeo7_ROI",ROI_Yeo7_number,".csv"))
 roi_filter <- ROIs_phylo_tree[[atlas_name]] %>% unique
 return(roi_filter)
}

roi_filter <- create_roi_filter(atlas_name, ROI_Yeo7_number)

```



# Build the list that will be fed into the function to compute all the stats
```{r}

make_rdata_list <- function(atlas_name) {
  # Define all possible filename patterns
  files <- c(
    paste0(atlas_name, "_N26__EER.RData"),
    paste0(atlas_name, "_N26_filtered_EER.RData"),
    paste0(atlas_name, "_N26_res_EER.RData"),
    paste0(atlas_name, "_N26_res_filtered_EER.RData")
  )
  
  # Define corresponding labels
  labels <- c("original", "filtered", "res", "res_filtered")
  
  # Build list of lists
  rdata_list <- map2(files, labels, ~ list(file = .x, atlas = atlas_name, label = .y))
  
  rdata_list
}

RData_files <- make_rdata_list(atlas_name)




```



# Function to compute the table of ttest + BF01
```{r}

rsa_stats_table <- function(file_path, atlas, label) {
  
  # Internal helper to load the RData
  load_RData <- function(file_path) {
    local({
      load(file_path)  # assumes object inside is named RSA
      RSA %>% dplyr::select(sub, roi, dplyr::starts_with('rsa'))
    })
  }
  
  # Load the data
  df <- load_RData(file_path)
  
  # Detect RSA columns
  rsa_cols <- grep("^rsa_rdm", names(df), value = TRUE)
  roi_list <- unique(df$roi)
  
  # Iterate over ROIs
  results <- map_dfr(roi_list, function(r) {
    df_roi <- df %>% filter(roi == r)
    
    # One-sample t-tests
    t_tests <- map(rsa_cols, function(col) {
      t_res <- t.test(df_roi[[col]])
      bf_res <- ttestBF(df_roi[[col]], mu = 0)
      bf01 <- 1 / as.vector(bf_res)  # correct BF01
      
      paste0("T=", round(t_res$statistic, 2),
             ", p=", round(t_res$p.value, 3),
             ", BF01=", round(bf01, 2))
    })
    names(t_tests) <- paste0(gsub("rsa_rdm_", "", rsa_cols), "_vs_0")
    
    # Paired comparisons
    combs <- combn(rsa_cols, 2, simplify = FALSE)
    paired_tests <- map(combs, function(pair) {
      t_res <- t.test(df_roi[[pair[1]]], df_roi[[pair[2]]], paired = TRUE)
      bf_res <- ttestBF(df_roi[[pair[1]]], df_roi[[pair[2]]], paired = TRUE)
      bf01 <- 1 / as.vector(bf_res)
      
      paste0("T=", round(t_res$statistic, 2),
             ", p=", round(t_res$p.value, 3),
             ", BF01=", round(bf01, 2))
    })
    names(paired_tests) <- map_chr(combs, ~ paste0(gsub("rsa_rdm_", "", .x[1]),
                                                   "_vs_",
                                                   gsub("rsa_rdm_", "", .x[2])))
    
    tibble(
      atlas = atlas,   # NEW: first column
      roi = r,
      label = label,
      !!!t_tests,
      !!!paired_tests
    )
  })
  
  results
}
```


# Run the function on all RDatas
```{r}

results <- map_dfr(RData_files, ~ rsa_stats_table(.x$file, .x$atlas, .x$label)) %>% 
  arrange(roi)

# results

```


# Build the gt_table (colors)
P < 0.05 are in green, BF01> 3 are in darkblue

```{r}

make_rsa_gt_table_colored <- function(results, roi_filter = NULL) {
  
  # Optionally filter ROIs
  if (!is.null(roi_filter)) {
    results <- results %>% filter(roi %in% roi_filter)
  }
  
  # Ensure atlas, roi, label are first columns
  results <- results %>% select(atlas, roi, label, everything()) %>% 
    mutate(roi_group = paste0("ROI ", roi))
  
  # Build initial gt table, grouped by ROI
  gt_table <- results %>%
    gt(groupname_col = "roi_group") %>%
    tab_header(
      title = "RSA Summary Table",
      subtitle = "T-tests and Bayes Factors per ROI"
    ) %>%
    cols_label(
      atlas = "ROI/Atlas",
      label = "RSA flavour",
      emotion_vs_0 = "Emotion ≠ 0",
      aroval_vs_0 = "AroVal ≠ 0",
      arousal_vs_0 = "Arousal ≠ 0",
      emotion_vs_arousal = "Emotion ≠ Arousal",
      emotion_vs_aroval = "Emotion ≠ AroVal",
      arousal_vs_aroval = "AroVal ≠ Arousal"
    ) %>%
    tab_style(
      style = cell_text(weight = "bold"),
      locations = cells_column_labels(everything())
    ) %>%
    tab_style(
      style = cell_text(weight = "bold"),
      locations = cells_row_groups()
    ) %>%
    tab_style(
      style = cell_text(weight = "bold", size = px(16)),
      locations = cells_title(groups = "title")
    ) %>%
    tab_style(
      style = cell_text(weight = "bold", size = px(14)),
      locations = cells_title(groups = "subtitle")
    ) %>%
    tab_options(
      table.font.size = px(12),
      heading.title.font.size = px(16),
      heading.subtitle.font.size = px(14)
    )
  
  # Apply colored text based on p < 0.05 (green) and BF01 > 3 (blue)
  for (col_name in names(results)[-(1:3)]) {  # skip atlas, roi, label
    # green text for p < 0.05
    rows_green <- which(as.numeric(str_extract(results[[col_name]], "(?<=p=)[0-9\\.e-]+")) < 0.05)
    if (length(rows_green) > 0) {
      gt_table <- gt_table %>%
        tab_style(
          style = cell_text(color = "darkgreen", weight = "bold"),
          locations = cells_body(columns = all_of(col_name), rows = rows_green)
        )
    }
    
    # blue text for BF01 > 3
    rows_blue <- which(as.numeric(str_extract(results[[col_name]], "(?<=BF01=)[0-9\\.e-]+")) > 3)
    if (length(rows_blue) > 0) {
      gt_table <- gt_table %>%
        tab_style(
          style = cell_text(color = "darkblue", weight = "bold"),
          locations = cells_body(columns = all_of(col_name), rows = rows_blue)
        )
    }
  }
  
  gt_table
}


# Preview
gt_table <- make_rsa_gt_table_colored(results)
# gt_table <- make_rsa_gt_table_colored(results, roi_filter = 5)

gt_table

```



```{r}

gtsave(gt_table, paste0(atlas_name, "_rsa_table.pdf"))

```








































# Table with bold only (OLE)
```{r, eval=FALSE}
# Build the gt_table
P < 0.05 and BF01> 3 are in bold
```{r}

make_rsa_gt_table <- function(results, roi_filter = NULL) {
  
  # Optionally filter ROIs
  if (!is.null(roi_filter)) {
    results <- results %>% filter(roi %in% roi_filter)
  }
  
  # Ensure atlas, roi, label are first columns
  results <- results %>% select(atlas, roi, label, everything())
  
  # Compute bold flags (p < 0.05 or BF01 > 3)
  bold_flags <- results %>%
    select(-atlas, -roi, -label) %>%
    map_df(~{
      p_val <- as.numeric(str_extract(.x, "(?<=p=)[0-9\\.e-]+"))
      bf01_val <- as.numeric(str_extract(.x, "(?<=BF01=)[0-9\\.e-]+"))
      (p_val < 0.05 | bf01_val > 3)
    })
  
  # Attach the atlas, roi, label columns
  bold_flags <- bind_cols(results %>% select(atlas, roi, label), bold_flags)
  
  # Create gt table
  gt_table <- results %>%
    gt(groupname_col = "roi") %>%   # keep ROI as visible column
    tab_header(
      title = "RSA Summary Table",
      subtitle = "T-tests and Bayes Factors per ROI"
    ) %>%
    cols_label(
      atlas = "ROI/Atlas",
      roi = "ROI",
      label = "RSA flavour",
      emotion_vs_0 = "Emotion ≠ 0",
      aroval_vs_0 = "AroVal ≠ 0",
      arousal_vs_0 = "Arousal ≠ 0",
      emotion_vs_arousal = "Emotion ≠ Arousal",
      emotion_vs_aroval = "Emotion ≠ AroVal",
      arousal_vs_aroval = "AroVal ≠ Arousal"
    ) %>%
    # Bold column headers
    tab_style(
      style = cell_text(weight = "bold"),
      locations = cells_column_labels(everything())
    ) %>%
    # Bold title
    tab_style(
      style = cell_text(weight = "bold", size = px(16)),
      locations = cells_title(groups = "title")
    ) %>%
    # Bold subtitle
    tab_style(
      style = cell_text(weight = "bold", size = px(14)),
      locations = cells_title(groups = "subtitle")
    ) %>%
    tab_options(
      table.font.size = px(12),
      heading.title.font.size = px(16),
      heading.subtitle.font.size = px(14)
    )
  
  # Apply bold formatting to flagged cells
  for (col_name in names(results)[-(1:3)]) {  # skip atlas, roi, label
    rows_to_bold <- which(bold_flags[[col_name]])
    if (length(rows_to_bold) > 0) {
      gt_table <- gt_table %>%
        tab_style(
          style = cell_text(weight = "bold"),
          locations = cells_body(
            columns = all_of(col_name),
            rows = rows_to_bold
          )
        )
    }
  }
  
  return(gt_table)
}


# Preview
# gt_table <- make_rsa_gt_table(results)
gt_table <- make_rsa_gt_table(results, roi_filter = 5)

gt_table
```
```





