---
title: "sketches"
author: "LC"
date: "2025-09-30"
output: html_document
---

NB: this script uses the `do_RSA_tables_pdf_FUNCTIONS.R`

The `Phylogenetic_tree_Yeo7_ROI[5/7].csv` used by `create_roi_filter()` come from
the Phylogenetic ROI tree created in `ROIS_REPO`


```{r message=FALSE}
library(tidyverse)
library(BayesFactor)
library(gt)
library(webshot2)  # for saving as PDF
library(stringr)
```



```{r message=FALSE}
source("do_RSA_tables_pdf_FUNCTIONS.R")

root="/data00/leonardo/RSA/analyses/RSA_ROI_APP/results_RSA_ROI"

atlas_name <- "Yeo7"
# atlas_name <- "Yeo17"
# atlas_name <- "HO_cort"
# atlas_name <- "anatomy_toolbox"

ROI_Yeo7_number = 5  # 5 or 7

# ------- end of user-defined parameters


# Make a list of all the RData results for one atlas
RData_files <- make_rdata_list(atlas_name)

# Run all the ttest and ttestBF and report the results
results <- RData_files %>% 
  map_dfr(~ rsa_stats_table(.x$file, .x$atlas, .x$label)) %>% 
  arrange(roi)

# Return the list of the regions to examine for that atlas_name and choice of ROI_Yeo7_number
atlas_roi_filter <- create_roi_filter(atlas_name, ROI_Yeo7_number)

# Produces the gt_table for the desired ROIs for that atlas
gt_table <- make_rsa_gt_table_colored(results, roi_filter = atlas_roi_filter)
gt_table

# # Saves it into a pdf file
# gtsave(gt_table, paste0(atlas_name, "_rsa_table.pdf"))

```

# ---------------------------- IMPORTANT -------------------------------------

**THE CHUNKS BELOW NEED NOT TO BE RUN.**

They refer to the development stage. Now all these functions are stored 
in `do_RSA_tables_pdf_FUNCTIONS.R`. The pdf is generated already at the end
of the previous chunk.


# Return the list of the regions to examine for that atlas_name and choice of ROI_Yeo7_number
```{r, message=FALSE}
create_roi_filter <- function(atlas_name, ROI_Yeo7_number) {
 ROIs_phylo_tree <- read_csv(paste0("Phylogenetic_tree_Yeo7_ROI",ROI_Yeo7_number,".csv"))
 roi_filter <- ROIs_phylo_tree[[atlas_name]] %>% unique
 return(roi_filter)
}

roi_filter <- create_roi_filter(atlas_name, ROI_Yeo7_number)

```



# Build the list that will be fed into the function to compute all the stats
```{r}

make_rdata_list <- function(atlas_name) {
  # Define all possible filename patterns
  files <- c(
    paste0(atlas_name, "_N26__EER.RData"),
    paste0(atlas_name, "_N26_filtered_EER.RData"),
    paste0(atlas_name, "_N26_res_EER.RData"),
    paste0(atlas_name, "_N26_res_filtered_EER.RData")
  )
  
  # Define corresponding labels
  labels <- c("original", "filtered", "res", "res_filtered")
  
  # Build list of lists
  rdata_list <- map2(files, labels, ~ list(file = .x, atlas = atlas_name, label = .y))
  
  rdata_list
}

RData_files <- make_rdata_list(atlas_name)




```



# Function to compute the table of ttest + BF01
```{r}

rsa_stats_table <- function(file_path, atlas, label) {
  
  # Internal helper to load the RData
  load_RData <- function(file_path) {
    local({
      load(file_path)  # assumes object inside is named RSA
      RSA %>% dplyr::select(sub, roi, dplyr::starts_with('rsa'))
    })
  }
  
  # Load the data
  df <- load_RData(file_path)
  
  # Detect RSA columns
  rsa_cols <- grep("^rsa_rdm", names(df), value = TRUE)
  roi_list <- unique(df$roi)
  
  # Iterate over ROIs
  results <- map_dfr(roi_list, function(r) {
    df_roi <- df %>% filter(roi == r)
    
    # One-sample t-tests
    t_tests <- map(rsa_cols, function(col) {
      t_res <- t.test(df_roi[[col]])
      bf_res <- ttestBF(df_roi[[col]], mu = 0)
      bf01 <- 1 / as.vector(bf_res)  # correct BF01
      
      paste0("T=", round(t_res$statistic, 2),
             ", p=", round(t_res$p.value, 3),
             ", BF01=", round(bf01, 2))
    })
    names(t_tests) <- paste0(gsub("rsa_rdm_", "", rsa_cols), "_vs_0")
    
    # Paired comparisons
    combs <- combn(rsa_cols, 2, simplify = FALSE)
    paired_tests <- map(combs, function(pair) {
      t_res <- t.test(df_roi[[pair[1]]], df_roi[[pair[2]]], paired = TRUE)
      bf_res <- ttestBF(df_roi[[pair[1]]], df_roi[[pair[2]]], paired = TRUE)
      bf01 <- 1 / as.vector(bf_res)
      
      paste0("T=", round(t_res$statistic, 2),
             ", p=", round(t_res$p.value, 3),
             ", BF01=", round(bf01, 2))
    })
    names(paired_tests) <- map_chr(combs, ~ paste0(gsub("rsa_rdm_", "", .x[1]),
                                                   "_vs_",
                                                   gsub("rsa_rdm_", "", .x[2])))
    
    tibble(
      atlas = atlas,   # NEW: first column
      roi = r,
      label = label,
      !!!t_tests,
      !!!paired_tests
    )
  })
  
  results
}
```


# Run the function on all RDatas
```{r}

results <- map_dfr(RData_files, ~ rsa_stats_table(.x$file, .x$atlas, .x$label)) %>% 
  arrange(roi)

# results

```


# Build the gt_table (colors)
P < 0.05 are in green, BF01> 3 are in darkblue

```{r}

make_rsa_gt_table_colored <- function(results, roi_filter = NULL) {
  
  # Optionally filter ROIs
  if (!is.null(roi_filter)) {
    results <- results %>% filter(roi %in% roi_filter)
  }
  
  # Ensure atlas, roi, label are first columns
  results <- results %>% select(atlas, roi, label, everything()) %>% 
    mutate(roi_group = paste0("ROI ", roi))
  
  # Build initial gt table, grouped by ROI
  gt_table <- results %>%
    gt(groupname_col = "roi_group") %>%
    tab_header(
      title = "RSA Summary Table",
      subtitle = "T-tests and Bayes Factors per ROI"
    ) %>%
    cols_label(
      atlas = "ROI/Atlas",
      label = "RSA flavour",
      emotion_vs_0 = "Emotion ≠ 0",
      aroval_vs_0 = "AroVal ≠ 0",
      arousal_vs_0 = "Arousal ≠ 0",
      emotion_vs_arousal = "Emotion ≠ Arousal",
      emotion_vs_aroval = "Emotion ≠ AroVal",
      arousal_vs_aroval = "AroVal ≠ Arousal"
    ) %>%
    tab_style(
      style = cell_text(weight = "bold"),
      locations = cells_column_labels(everything())
    ) %>%
    tab_style(
      style = cell_text(weight = "bold"),
      locations = cells_row_groups()
    ) %>%
    tab_style(
      style = cell_text(weight = "bold", size = px(16)),
      locations = cells_title(groups = "title")
    ) %>%
    tab_style(
      style = cell_text(weight = "bold", size = px(14)),
      locations = cells_title(groups = "subtitle")
    ) %>%
    tab_options(
      table.font.size = px(12),
      heading.title.font.size = px(16),
      heading.subtitle.font.size = px(14)
    )
  
  # Apply colored text based on p < 0.05 (green) and BF01 > 3 (blue)
  for (col_name in names(results)[-(1:3)]) {  # skip atlas, roi, label
    # green text for p < 0.05
    rows_green <- which(as.numeric(str_extract(results[[col_name]], "(?<=p=)[0-9\\.e-]+")) < 0.05)
    if (length(rows_green) > 0) {
      gt_table <- gt_table %>%
        tab_style(
          style = cell_text(color = "darkgreen", weight = "bold"),
          locations = cells_body(columns = all_of(col_name), rows = rows_green)
        )
    }
    
    # blue text for BF01 > 3
    rows_blue <- which(as.numeric(str_extract(results[[col_name]], "(?<=BF01=)[0-9\\.e-]+")) > 3)
    if (length(rows_blue) > 0) {
      gt_table <- gt_table %>%
        tab_style(
          style = cell_text(color = "darkblue", weight = "bold"),
          locations = cells_body(columns = all_of(col_name), rows = rows_blue)
        )
    }
  }
  
  gt_table
}


# Preview
gt_table <- make_rsa_gt_table_colored(results)
# gt_table <- make_rsa_gt_table_colored(results, roi_filter = 5)

gt_table

```



```{r}

gtsave(gt_table, paste0(atlas_name, "_rsa_table.pdf"))

```








































# Table with bold only (OLE)
```{r, eval=FALSE}
# Build the gt_table
P < 0.05 and BF01> 3 are in bold
```{r}

make_rsa_gt_table <- function(results, roi_filter = NULL) {
  
  # Optionally filter ROIs
  if (!is.null(roi_filter)) {
    results <- results %>% filter(roi %in% roi_filter)
  }
  
  # Ensure atlas, roi, label are first columns
  results <- results %>% select(atlas, roi, label, everything())
  
  # Compute bold flags (p < 0.05 or BF01 > 3)
  bold_flags <- results %>%
    select(-atlas, -roi, -label) %>%
    map_df(~{
      p_val <- as.numeric(str_extract(.x, "(?<=p=)[0-9\\.e-]+"))
      bf01_val <- as.numeric(str_extract(.x, "(?<=BF01=)[0-9\\.e-]+"))
      (p_val < 0.05 | bf01_val > 3)
    })
  
  # Attach the atlas, roi, label columns
  bold_flags <- bind_cols(results %>% select(atlas, roi, label), bold_flags)
  
  # Create gt table
  gt_table <- results %>%
    gt(groupname_col = "roi") %>%   # keep ROI as visible column
    tab_header(
      title = "RSA Summary Table",
      subtitle = "T-tests and Bayes Factors per ROI"
    ) %>%
    cols_label(
      atlas = "ROI/Atlas",
      roi = "ROI",
      label = "RSA flavour",
      emotion_vs_0 = "Emotion ≠ 0",
      aroval_vs_0 = "AroVal ≠ 0",
      arousal_vs_0 = "Arousal ≠ 0",
      emotion_vs_arousal = "Emotion ≠ Arousal",
      emotion_vs_aroval = "Emotion ≠ AroVal",
      arousal_vs_aroval = "AroVal ≠ Arousal"
    ) %>%
    # Bold column headers
    tab_style(
      style = cell_text(weight = "bold"),
      locations = cells_column_labels(everything())
    ) %>%
    # Bold title
    tab_style(
      style = cell_text(weight = "bold", size = px(16)),
      locations = cells_title(groups = "title")
    ) %>%
    # Bold subtitle
    tab_style(
      style = cell_text(weight = "bold", size = px(14)),
      locations = cells_title(groups = "subtitle")
    ) %>%
    tab_options(
      table.font.size = px(12),
      heading.title.font.size = px(16),
      heading.subtitle.font.size = px(14)
    )
  
  # Apply bold formatting to flagged cells
  for (col_name in names(results)[-(1:3)]) {  # skip atlas, roi, label
    rows_to_bold <- which(bold_flags[[col_name]])
    if (length(rows_to_bold) > 0) {
      gt_table <- gt_table %>%
        tab_style(
          style = cell_text(weight = "bold"),
          locations = cells_body(
            columns = all_of(col_name),
            rows = rows_to_bold
          )
        )
    }
  }
  
  return(gt_table)
}


# Preview
# gt_table <- make_rsa_gt_table(results)
gt_table <- make_rsa_gt_table(results, roi_filter = 5)

gt_table
```
```





