---
title: "Fix ratings filename/format and copy them to the raw_data directory"
output: html_notebook
---

The previous people who worked on this data left a messy situation: some files were saved as xslx, some with csv, and the names were always different, including sometimes the name of the parent directory.

To fix this, rather than writing a very complex script, I had to fix some filenames manually. After that, I parse the files and put them in the raw_data dir

- if the file is in excel format, I load it into a tmp dataframe and write it as csv to the destination

- if the file is already in csv, I copy it to the destination


NB: all the following is preceded by the manual creation of the `ratings` folder for each sub
in the raw data dir

```bash
cd /data00/leonardo/RSA/raw_data
for i in `ls`; do mkdir ${i}/ratings; done
```

```{r, message=FALSE}
library(tidyverse)

subs_file <- "/data00/leonardo/RSA/sub_list.txt"
subs <- sprintf("%02d",readLines(subs_file) %>% as.numeric)

keys <- read_csv("participant_key.csv") %>% 
  mutate(cas = sprintf("%02d", cas), ritu = as.character(ritu)) %>% 
  mutate(Rating_EXP1 = NA, Rating_EXP2 = NA, Arousal_EXP3 = NA, Valence_EXP4 = NA)

keys
```


```{r}

ritu_root <- "/data01/Ritu/7T_EmotionInsula/Data_collection/"
raw_data_root <- "/data00/leonardo/RSA/raw_data"

cp_rename <- function(sub_id, keys) {
  
  keys_one_sub <- keys %>% 
      filter(cas == sub_id) %>% 
      mutate(sub_id = sub_id) %>% 
      relocate(sub_id, cas, ritu)
  
  ritu_key <- keys_one_sub$ritu
  cas_key <- keys_one_sub$cas
  
  for (EXP in c("Rating_EXP1", "Rating_EXP2", "Arousal_EXP3", "Valence_EXP4")) {
    
    orig <- paste0(ritu_root, "/S", ritu_key, "/Posttasks/S",ritu_key,"_",EXP)
    
    dest <- paste0(raw_data_root,"/sub-", cas_key, "/ratings/sub-",cas_key,"_",EXP)
    
    # Excel file : write as csv
    if (paste0(orig,'.xlsx') %>% file.exists) {
      keys_one_sub[[EXP]] = 'XLSX'
      tmp <- read_excel(paste0(orig,'.xlsx'))
      write_csv(tmp, paste0(dest,'.csv'))
      print(paste0(dest,'.csv'))
    
    # csv file : cp it to destination
    } else if (paste0(orig,'.csv') %>% file.exists) {
      keys_one_sub[[EXP]] = 'CSV'
      file.copy(from = paste0(orig,'.csv'), to = paste0(dest,'.csv'))
      print(paste0(dest,'.csv'))
    
    # file does not exist either aas xlsx or csv
    } else {
      keys_one_sub[[EXP]] = NA
      print(orig)
    }
    
    
  }
  cat("\n")
  return(keys_one_sub)
}


# # Test for one sub
# sub_id = subs[1]
# cp_rename(sub_id, keys)

subs %>% map_dfr(~ cp_rename(.x, keys))


```


