---
title: "Prepare EMOTION ratings"
output: html_notebook
---

```{r, message=FALSE}
library(tidyverse)
library(janitor)

subs_file <- "/data00/leonardo/RSA/sub_list.txt"
subs <- sprintf("%02d",readLines(subs_file) %>% as.numeric)
movie_encodings_table <- read_csv("/data00/leonardo/RSA/encodings/movie_encodings.csv")

```


# Preparing the `emotion_ratings.csv` file
The final csv will be saved in `/data00/leonardo/RSA/analyses/RATINGS/emotion_ratings.csv`

This file will be used in the fmri as well as in the subsequent RSA. Note that while the participants carried out 8 runs inside the scanner, they only rated each movie once.

## Congruence testing
**IMPORTANT** : after preparing the csv file for the fmri analysis, we also conduct a simple analysis to verify sensitivity of the behavioural task. In brief, the highest rating given to any movie is hypothesized to refer to the emotion actually displayed in the movie.  

Scroll down to `Examine movie/rating congruence` to see the outcome of this analysis


## Code preparation 
NB: all of the code in the following cell is just for preparation. The actual conversion will be carried out with a function encompassing all of this code and carried out using `purrr` on all subs

```{r, message=FALSE, eval=FALSE}
sub_id <- subs[1]

orig_root = paste0("/data00/leonardo/RSA/raw_data/sub-",sub_id,"/ratings/sub-",sub_id,"_Rating_")

# Sorry for not being DRY : there are only two files and in this way it's much clearer what is going on

# Rating EXP 1
df_EXP_1_raw_allrows <- read_csv(paste0(orig_root,"EXP1.csv")) %>% janitor::clean_names()

# select only rows below the one with "Start Real Experiment"
row_exp_start <- which(df_EXP_1_raw_allrows$display == "Start Real Experiment") + 1
df_EXP_1_raw <- df_EXP_1_raw_allrows[row_exp_start:nrow(df_EXP_1_raw_allrows),]

df_EXP_1 <- df_EXP_1_raw %>% 
  filter(
    # video == "AK_A_04.mp4", 
    zone_name == "Ratingintensity",
    zone_type == "response_slider_endValue"
  ) %>% 
  mutate(rate_this_emo = str_extract(rating, "Fear|Pain|Anger|Happy|Sad|Disgust")) %>% 
  select(video, response, video, rate_this_emo) %>% 
  mutate(response = as.numeric(response))
  # pivot_wider(names_from = rate_this_emo, values_from = response)


# Rating EXP 2
df_EXP_2_raw_allrows <- read_csv(paste0(orig_root,"EXP2.csv")) %>% janitor::clean_names()

# select only rows below the one with "Start Real Experiment"
row_exp_start <- which(df_EXP_2_raw_allrows$display == "Start Real Experiment") + 1
df_EXP_2_raw <- df_EXP_2_raw_allrows[row_exp_start:nrow(df_EXP_2_raw_allrows),]

df_EXP_2 <- df_EXP_2_raw %>% 
  filter(
    # video == "AK_A_04.mp4", 
    zone_name == "Ratingintensity",
    zone_type == "response_slider_endValue"
  ) %>% 
  mutate(rate_this_emo = str_extract(rating, "Fear|Pain|Anger|Happy|Sad|Disgust")) %>% 
  select(video, response, video, rate_this_emo) %>% 
  mutate(response = as.numeric(response))
  # pivot_wider(names_from = rate_this_emo, values_from = response)


# Merge the results from the two experimental runs and pivot it wider to have columns
# for the ratings for each emotion
df <- rbind(df_EXP_1, df_EXP_2)
df_wide <- df %>% pivot_wider(names_from = rate_this_emo, values_from = response)

# add columns for:
# - original_code : extracted from the movie filename
# - high_low_code : the intensity of the emotion displayed in that movie
# - emotion : the emotion displayed in that movie
# - actor : the actor performing in that movie 
df_wide_codes <- df_wide %>% 
  janitor::clean_names() %>% 
  mutate(original_code = str_remove(video, "\\..*$")) %>% 
  inner_join(movie_encodings_table, by="original_code") %>% 
  rename_with(~ paste0("r_", .), c("happy","disgust","sad","fear","anger","pain") ) %>% 
  arrange(high_low_code)

df_wide_codes

```


## Write the function to be passed to purrr over subs

```{r, message=F}
prepare_csv_ratings <- function(sub_id) {
  
  paste0("Processing subj ",sub_id,"\n") %>% cat
  
  orig_root = paste0("/data00/leonardo/RSA/raw_data/sub-",sub_id,"/ratings/sub-",sub_id,"_Rating_")

  # Sorry for not being DRY : there are only two files and in this way it's much clearer what is going on
  
  # Rating EXP 1
  df_EXP_1_raw_allrows <- read_csv(paste0(orig_root,"EXP1.csv")) %>% janitor::clean_names()
  
  # select only rows below the one with "Start Real Experiment"
  row_exp_start <- which(df_EXP_1_raw_allrows$display == "Start Real Experiment") + 1
  df_EXP_1_raw <- df_EXP_1_raw_allrows[row_exp_start:nrow(df_EXP_1_raw_allrows),]
  
  df_EXP_1 <- df_EXP_1_raw %>% 
    filter(
      # video == "AK_A_04.mp4", 
      zone_name == "Ratingintensity",
      zone_type == "response_slider_endValue"
    ) %>% 
    mutate(rate_this_emo = str_extract(rating, "Fear|Pain|Anger|Happy|Sad|Disgust")) %>% 
    select(video, response, video, rate_this_emo) %>% 
    mutate(response = as.numeric(response))
    # pivot_wider(names_from = rate_this_emo, values_from = response)
  
  
  # Rating EXP 2
  df_EXP_2_raw_allrows <- read_csv(paste0(orig_root,"EXP2.csv")) %>% janitor::clean_names()
  
  # select only rows below the one with "Start Real Experiment"
  row_exp_start <- which(df_EXP_2_raw_allrows$display == "Start Real Experiment") + 1
  df_EXP_2_raw <- df_EXP_2_raw_allrows[row_exp_start:nrow(df_EXP_2_raw_allrows),]
  
  df_EXP_2 <- df_EXP_2_raw %>% 
    filter(
      # video == "AK_A_04.mp4", 
      zone_name == "Ratingintensity",
      zone_type == "response_slider_endValue"
    ) %>% 
    mutate(rate_this_emo = str_extract(rating, "Fear|Pain|Anger|Happy|Sad|Disgust")) %>% 
    select(video, response, video, rate_this_emo) %>% 
    mutate(response = as.numeric(response))
    # pivot_wider(names_from = rate_this_emo, values_from = response)
  
  
  # Merge the results from the two experimental runs and pivot it wider to have columns
  # for the ratings for each emotion
  df <- rbind(df_EXP_1, df_EXP_2)
  df_wide <- df %>% pivot_wider(names_from = rate_this_emo, values_from = response)
  
  # add columns for:
  # - original_code : extracted from the movie filename
  # - high_low_code : the intensity of the emotion displayed in that movie
  # - emotion : the emotion displayed in that movie
  # - actor : the actor performing in that movie 
  # - sub_id
  df_wide_codes <- df_wide %>% 
    janitor::clean_names() %>% 
    mutate(original_code = str_remove(video, "\\..*$")) %>% 
    inner_join(movie_encodings_table, by="original_code") %>% 
    rename_with(~ paste0("r_", .), c("happy","disgust","sad","fear","anger","pain") ) %>% 
    arrange(high_low_code) %>% 
    mutate(sub = sub_id) %>% 
    relocate(sub)
}
```


## Purrr over subs, save csv, display the table
```{r, message=F}
ff <- subs %>% map_dfr(~ prepare_csv_ratings(.x))

write_csv(ff, "/data00/leonardo/RSA/analyses/RATINGS/emotion_ratings.csv")

ff
```


# Examine movie/rating congruence
Each movie displays a specific emotion, however the participant is requested to rate each movie for all the 6 emotions. 

It is expected that across emotion rating, the highest will be for the emotion which is displayed in the movie. However it is better to verify this.

```{r}

df_congruence <- ff %>%
  select(sub,original_code, emotion, starts_with("r_")) %>%
  pivot_longer(starts_with("r_"), names_to = "rating") %>%
  group_by(sub, original_code, emotion) %>%
  filter(value == max(value)) %>%
  filter(value != 0) %>%
  mutate(rating = str_remove(rating,"r_")) %>%
  mutate(consistent = ifelse(emotion==rating,1,0)) %>%
  ungroup %>% 
  group_by(sub) %>%
  summarise(
    congruence = round(sum(consistent)/n(),2)
  ) %>% 
  arrange(congruence)

df_congruence

df_congruence %>%
  ggplot(aes(x = reorder(sub, congruence), y = congruence)) +
  geom_bar(stat = "identity", color = "grey", fill = "lightblue") +
  theme_minimal() +
  labs(title = "Congruence between displayed emotion and max rated emotion", x = "sub")

```

## Inspect participants with low movie/rating congruence
The ratings for each emotion for each movie are shown. Here we do not remove 0 ratings to show the rating given for each emotion.

```{r, eval=F}
ff %>%
  filter(sub=="34") %>%
  select(sub,original_code, emotion, starts_with("r_")) %>%
  pivot_longer(starts_with("r_"), names_to = "rating") %>%
  group_by(sub, original_code, emotion) %>%
  # filter(value != 0) %>%
  arrange(desc(value)) %>% 
  group_split()
```


## Look also at the congruence with the second highest rated emotion
Maybe the participants with _low_ congruence - when looking at the max rated emotion - did rate as second highest the actually congruent emotion? Not really.


```{r}
df_second_highest_congruence <- ff %>%
  select(sub, original_code, emotion, starts_with("r_")) %>%
  pivot_longer(starts_with("r_"), names_to = "rating", values_to = "value") %>%
  group_by(sub, original_code, emotion) %>%
  arrange(desc(value)) %>%
  slice(2) %>%
  filter(value != 0) %>%
  mutate(rating = str_remove(rating, "r_")) %>%
  mutate(consistent = ifelse(emotion == rating, 1, 0)) %>%
  ungroup() %>% 
  group_by(sub) %>%
  summarise(congruence = round(sum(consistent) / n(), 2)) %>% 
  arrange(congruence)

df_second_highest_congruence %>% 
  rename(second_congruence = congruence) %>% 
  select(sub, second_congruence) %>% 
  inner_join(df_congruence, by = "sub") %>% 
  relocate(sub, congruence, second_congruence) %>% 
  arrange(desc(congruence), desc(second_congruence)) %>% 
  ggplot(aes(x=congruence, y=second_congruence, label=sub)) + # Add label aesthetic for sub
  geom_point() + 
  geom_text(vjust=-0.5, hjust=0.5, check_overlap = TRUE) + # Add text labels for each point
  theme_minimal() + 
  labs(
    title = "Movie/rating congruence in the highest vs. second highest rated emotion",
    x = "congruence with highest rated emotion",
    y = "congruence with second highest rated emotion"
  )

```





