---
title: "do_RSA_v7"
author: "LC"
date: "2024-03-10"
output: html_document
---

NB: this Rmd can be used for all types of ratings: emotion, arousal, valence

```{r load_libraries, message=F}
library(tidyverse)
library(future)
library(furrr)
library(tictoc)
library(RNifti)
library(proxy) # distances
library(profvis)
library(DT)
library(formattable)

source("funs_emotion_predictors_V8.R")

bd="/data00/leonardo/RSA/analyses"

# can be emotion, arousal, valence
ratings_type <- "emotion"

# the following is out contrast of interest and it's always the same
copes_type <- "emotion_predictors"

# Choose one atlas among:
# HO_cort_2mm.nii.gz
# HO_sub_2mm.nii.gz
# juelich_2mm.nii.gz
# Schaefer100_2mm.nii.gz
# Schaefer200_2mm.nii.gz
# Schaefer400_2mm.nii.gz
# insula_HO_clusters_[N].nii.gz
atlas_filename <- "insula_HO_clusters_600.nii.gz"


# Choose the distance metric to use for fmri RDM
# The metric for ratings RDM *should be* euclidean, since arousal and valence
# ratings have only one value
# Supported methods: 'pearson', 'spearman', 'euclidean', 'cosine' or 'mahalanobis'
dist_method_rating <- "euclidean"
dist_method_fmri = "euclidean"
dist_method_rsa = "pearson"


cat(paste0("Ratings RDM to be calculated with : ", dist_method_rating, "\n"))
cat(paste0("fMRI RDM to be calculated with : ", dist_method_fmri, "\n"))


RDM_type = "tril"  # svd or tril. Must be the same for both rating and fmri
ncomp_svd = 3

# Vector of zeropadded sub_ids
subs_file <- "/data00/leonardo/RSA/sub_list.txt"
subs <- sprintf("%02d",readLines(subs_file) %>% as.numeric)

# ---------- ONLY TOP RATERS BELOW ------------
# subs <- c("02","03","12","11","22","09","29","28","26","32","23","15","20","19")

```


# Aux functions
```{r}
# plot_heatmap(D %>% as.matrix)
plot_heatmap <- function(M) {
  heatmap(M, Rowv = NA, Colv = NA, symm=T, revC = T)
}
```


# Create a df_path_copes with the location of the cope niis from the one_movie_per_ev model
Extract the pathname of all copes using the `list.files()` function.
Also define a copes_numba vector with all the copes numbers.

NB: The cope numbers in the `cope` column are NOT zeropadded since this is how they come out from FSL Feat
```{r, message=FALSE}

bd_copes = paste0(bd, "/",copes_type,"/results/2nd_level")

df_path_copes <- list.files(bd_copes, recursive = T) %>% tibble()
names(df_path_copes) <- "fname"

df_path_copes <- df_path_copes %>%
  filter(str_detect(fname, "gfeat/cope[0-9]+.feat/stats/cope")) %>%
  tidyr::separate(
    fname, c("sub","cope","tmp1","tmp2"),
    sep = "/", fill = "right", remove = F
  ) %>% 
  select(!starts_with("tmp")) %>% 
  mutate(sub = str_extract(sub, "[0-9]+")) %>% 
  mutate(cope = (str_extract(cope, "[0-9]+") %>% as.numeric) ) %>% 
  mutate(path = paste0(bd_copes,"/",fname)) %>% 
  select(!fname) %>% 
  arrange(sub,cope)

copes_numba <- df_path_copes$cope %>% unique


```


# Read atlas and create a vector of atlas labels
```{r}

bd_atlases = "/data00/leonardo/RSA/analyses/rsa/atlases"

atlas_path <- paste0(bd_atlases,"/",atlas_filename)

atlas_nii <- readNifti(paste0(bd_atlases,"/", atlas_filename))
region_labels <- atlas_nii[atlas_nii > 0] %>% unique %>% sort

```



# Calculate RATINGS RDMs - IMPORTANT DIFFERENCE FOR EMOTION_PREDICTORS MODEL

**THIS IS THE SAME CODE AS IN do_RSA_V[N].Rmd HOWEVER SINCE HERE WE HAVE DEFINED
ONE PREDICTOR FOR *ALL* MOVIES WITH THE SAME EMOTION, WE NEED TO CALCULATE
THE RATING FOR EACH EMOTION AS THE AVG (MEDIAN) RATING FOR EACH EMOTION MOVIE
FOR EACH SUB**

We do this first since the RDM for the ratings will be the same for whatever
atlas will be used. Also, it's much faster than the RDM_fmri and can be done
for all the sub_ids at once.


- **tril** : The final RDMs_rats has 546 nrows : 26 sub_ids * 21, where the latter
derives from ((7^2)-7)/7 = 21, i.e. tril from the D matrix of 56 movies
- **svd** : The final RDMs_rats has 4368 nrows : 26 sub_ids * 168, where the latter
derives from 56*3, i.e. the first three components from svd of the D matrix
```{r, message=FALSE}

bd_ratings = paste0(bd,"/RATINGS")
ratings_path <- paste0(bd_ratings,"/",ratings_type,"_ratings.csv")
rats <- read_csv(ratings_path)

# ------- IMPORTANT : AVG ACROSS MOVIES FOR EMOTION PREDICTORS MODEL -----
rats <- rats %>% 
  select(sub, emotion, starts_with("r_")) %>%
  group_by(sub, emotion) %>% 
  reframe(across(starts_with("r_"), median, na.rm = TRUE)) %>% 
  ungroup %>% 
  arrange(sub, emotion)
# ------- IMPORTANT : AVG ACROSS MOVIES FOR EMOTION PREDICTORS MODEL -----

rats

# for RDM_type use:
# - tril for the lower triangular of the D matrix
# - svd for the first 3 pc from svd of the D matrix
RDMs_rats_wider <- subs %>% map_dfc(~ {
  do_RDM_ratings_one_sub(.x, dist_method_rating, rats, RDM_type =  RDM_type)
})

RDMs_rats <- RDMs_rats_wider %>% 
  pivot_longer(cols = everything(), names_to = "sub") %>% 
  arrange(sub)

RDMs_rats
```



# Calculate FMRI RDMs

## Define the functions
1. df_copes[one_sub] <- load_sub_copes(sub_id, copes_numba)
2. RDM_fmri[one_sub] <- calculate_fmri_RDMs(df_copes, atlas_path)

```{r load_copes, eval=FALSE}

# # Load copes for one sub_id
# # (reasonably fast: reads 56 nifti at 2mm resolution in ~ 2.5 sec)
# sub_id = "02"
# 
# df_copes <- load_sub_copes(sub_id, copes_numba, df_path_copes)
# 
# # Read atlas and calculate fmri RDMs for each atlas region
# dist_method_fmri = "euclidean"
# one_RDM_fmri <- calculate_fmri_RDMs(
#   df_copes, atlas_nii, dist_method_fmri, RDM_type = RDM_type
# )

```


## Furrr the calculation of RDMs
First, the copes are loaded into a df of 91x109x91 (MNI 2mm) rows and 56 columns.
Then the RDM is estimated for each atlas region as the tril of the similarity. 
matrix across the 56 subs
This is carried out in parallel for all 26 subs using `furrr::future_map_dfr`
```{r}
# 5 workers give the best performance (~ 15 sec)
plan(multisession, workers = 5)

tic()
RDMs_fmri <- subs %>% future_map_dfr(~{
  
  paste0("Calculating RDMs for sub ",.x,"\n") %>% cat
  
  df_copes <- load_sub_copes(.x, copes_numba, df_path_copes)
  
  one_RDM_fmri <- calculate_fmri_RDMs(
    df_copes, 
    atlas_nii, 
    dist_method_fmri, 
    RDM_type = RDM_type
  )
  
  one_RDM_fmri %>% mutate(sub = .x) %>% relocate(sub)
  
})
toc()

plan(sequential)
```




# RSA 
In the approach below purrr, tidy data and R in general shine at their best.

The final aim is to calculate the similarity between each sub/region fmri RDM
and each sub rat RDM. Therefore the fmri rdms is grouped by sub/region and the
rat rdms is grouped only by sub.

First, nesting both fmri and rat rdms allows to keep the dfs tidy. 
Soon after nesting, we can *right*-join the fmri to the rat, to create a copy
of the rat rdm for each sub/region of the atlas.

At this point, calculating the similarity is just a matter of using a purrr:map2
(or pmap as I prefer) to the fmri and rat columns

The simplicity yielded by the tidy data approach and the purrr functions is 
unparalleled.

```{r}

# nest the RDMs for both rat and fmri. In the second case, they are nested by sub AND region
rdms_fmri_nested <- RDMs_fmri %>% 
  pivot_longer(cols = starts_with("RDM"), names_to = "region") %>% 
  group_by(sub, region) %>% 
  nest %>% 
  rename(fmri = data)

rdms_rats_nested <- RDMs_rats %>% 
  group_by(sub) %>% 
  nest %>% 
  rename(rat = data)

# we put a copy of the rating RDM for each atlas region's RDM
joint_fmri_rat_nested <- right_join(rdms_fmri_nested,rdms_rats_nested, by = "sub")
# # the following is just to check that the join went as expected
# ff %>% mutate(sum_rat = rat %>% map_dbl(sum)  )

joint_fmri_rat_nested

# Calculate all similarity
# This is effectively when we are calculating the RSA
all_similarities <- joint_fmri_rat_nested %>% 
  mutate(similarity = list(fmri,rat) %>% pmap_dbl(~ cor(.x,.y, method = dist_method_rsa)) ) %>%
  ungroup

# Produce estimates of RSA acros subjects
RSA <- all_similarities %>%
  select(region, similarity) %>% 
  group_by(region) %>% 
  reframe(
    mean_similarity = round(mean(similarity),2),
    sd_similarity = round(sd(similarity),2)
  ) %>% 
  arrange(desc(mean_similarity))

RSA %>% datatable()
```



# Write the map with the values
```{r}

results_to_map <- RSA %>% 
  mutate(region = as.integer(str_replace(region, "RDM_region_",""))) %>% 
  select(region, mean_similarity)

results_to_map

results_path <- "/data00/leonardo/RSA/analyses/rsa_emotion_predictors/rsa_emotion_predictors_results"

results_filename <- create_filename_results(
  atlas_filename, ratings_type, dist_method_rating, dist_method_fmri, dist_method_rsa
)


# Function to write the results to a nii.gz volume
# (purrr gives issues here)
write_results_to_nii <- function(results_to_map, filename_results_nii, atlas_nii) {
  
  results_nii = atlas_nii
  
  for (ROI_numba in region_labels) {
    cat(paste0(ROI_numba," "))
    row_region <- which(results_to_map$region == ROI_numba)
    value <- results_to_map$mean_similarity[row_region]
    
    idx <- which(results_nii == ROI_numba)
    results_nii[idx] = value
  }

  writeNifti(results_nii, filename_results_nii)
}


# Write the results nii.gz in the rsa_results dir
write_results_to_nii(
  results_to_map, 
  paste0(results_path, "/", results_filename), 
  atlas_nii
)

```









# Write the map with the values OLE
```{r}

results_to_map <- RSA %>% 
  mutate(region = as.integer(str_extract(region, "\\d+$"))) %>% 
  select(region, mean_similarity)

results_to_map

filename_results_nii <- paste0("rsa_emotion_predictors_results/rsa_results_",ratings_type,"_",atlas_filename)

write_results_to_nii <- function(results_to_map, filename_results_nii, atlas_nii) {
  
  results_nii = atlas_nii
  
  for (ROI_numba in region_labels) {
    cat(paste0(ROI_numba," "))
    row_region <- which(results_to_map$region == ROI_numba)
    value <- results_to_map$mean_similarity[row_region]
    
    idx <- which(results_nii == ROI_numba)
    results_nii[idx] = value
  }

  writeNifti(results_nii, filename_results_nii)
}

write_results_to_nii(results_to_map, filename_results_nii, atlas_nii)

```







