---
title: "Check Behavioral RDMs - Various Distance Metrics"
author: "LC"
date: "2026-02-01"
output: html_document
---

# Purpose
Compare behavioral RDMs calculated with different distance metrics:
- Euclidean
- Mahalanobis
- Cosine

This script loads behavioral ratings and visualizes how different distance metrics affect RDM structure.

# Load libraries
```{r load_libraries, message=FALSE}
library(tidyverse)
library(ComplexHeatmap)
library(circlize)
library(grid)

source("funs_V15_CROSSNOBIS_ROIs.R")
```

# Define paths and parameters
```{r}
bd <- "/data00/leonardo/RSA/analyses"
bd_ratings <- paste0(bd, "/RATINGS")

# Ratings types to analyze
ratings_type <- c("emotion", "arousal", "aroval")

# Distance methods to compare
distance_methods <- c("euclidean", "mahalanobis", "cosine")

# Subject list
subs_file <- "/data00/leonardo/RSA/sub_list.txt"
subs <- sprintf("%02d", readLines(subs_file) %>% as.numeric)

# Remove neutrals
remove_neutrals <- "YES"
```

# Load and prepare behavioral ratings
```{r, message=FALSE}
load_ratings <- function(ratings_type, remove_neutrals) {
  
  ratings_path <- paste0(bd_ratings, "/", ratings_type, "_ratings.csv")
  rats <- read_csv(ratings_path)
  
  # Average across movies for emotion predictors model
  rats <- rats %>% 
    select(sub, emotion, high_low_code, starts_with("r_")) %>% 
    mutate(emotion = paste0(emotion, "_", str_match(high_low_code, "high|low"))) %>% 
    select(-high_low_code) %>% 
    group_by(sub, emotion) %>% 
    reframe(across(starts_with("r_"), median, na.rm = TRUE)) %>% 
    ungroup() %>% 
    arrange(sub, emotion)
  
  if (remove_neutrals == "YES") {
    rats <- rats %>% filter(!str_detect(emotion, "neutral"))
  }
  
  return(rats)
}

# Load all rating types
ratings_list <- map(ratings_type, ~ load_ratings(.x, remove_neutrals)) %>%
  set_names(ratings_type)

cat("Loaded ratings for:", paste(ratings_type, collapse = ", "), "\n")
cat("Number of subjects:", length(subs), "\n")
cat("Number of conditions:", nrow(ratings_list$emotion %>% filter(sub == subs[1])), "\n")
```


# Correlations in the emotion ratings (justification for use of euclidean)

For the behavioral representational dissimilarity matrices (RDMs), distances were computed from six emotion ratings (happy, disgust, sad, fear, anger, pain). The emotion dimensions showed modest correlations, with most off-diagonal correlations being small to moderate (|r| < 0.3). These correlations are expected given the near-mutual exclusivity of emotions across stimuli and do not indicate substantial redundancy among rating dimensions.

Euclidean distance was used on z-scored ratings, reflecting absolute differences along each emotion dimension. This choice is appropriate when rating dimensions are psychologically meaningful and distinct, and when differences in magnitude are theoretically relevant. In this context, Euclidean distance provides a transparent and stable measure of dissimilarity without requiring estimation of a covariance structure.

Although Mahalanobis distance can account for correlations between dimensions, it was not adopted here because the observed covariance structure primarily reflects constraints imposed by the stimulus set (e.g., limited co-occurrence of emotions) rather than measurement noise or unreliability. Using Mahalanobis distance in this case would therefore reweight dimensions based on stimulus design rather than on properties of the behavioral measurements themselves.

Cosine distance was also not used, as it discards information about overall rating magnitude and emphasizes relative profiles, which was not aligned with the present hypotheses.

Behavioral RDMs were computed using Euclidean distance on z-scored emotion ratings. Although emotion dimensions showed modest correlations, Mahalanobis distance was not used because covariance primarily reflected stimulus design constraints rather than measurement noise.

```{r}
corr_emotion <- ratings_list$emotion %>% 
  select(starts_with("r_")) %>% 
  cor

corrplot::corrplot.mixed(corr_emotion)
# corrplot::corrplot(corr_emotion, method = 'number')

```



# Calculate RDMs for each distance method
```{r}
calculate_RDMs_all_methods <- function(rats, distance_methods, subs) {
  
  # Filter for relevant subjects
  rats_filtered <- rats %>% filter(sub %in% subs)
  
  # Calculate RDM for each distance method
  rdms <- map(distance_methods, function(dist_method) {
    
    cat("  Calculating with", dist_method, "distance...\n")
    
    rats_filtered %>%
      select(sub, starts_with("r_")) %>%
      group_by(sub) %>%
      nest() %>%
      mutate(rdm = map(data, ~ DDOS_vec(.x, method = dist_method))) %>%
      select(sub, rdm)
    
  }) %>%
    set_names(distance_methods)
  
  return(rdms)
}

# Calculate RDMs for all rating types and distance methods
cat("\nCalculating RDMs...\n")
all_rdms <- map(ratings_type, function(rt) {
  cat("Processing", rt, "ratings:\n")
  calculate_RDMs_all_methods(ratings_list[[rt]], distance_methods, subs)
}) %>%
  set_names(ratings_type)

cat("\nRDM calculation complete!\n")
```

# Compute mean RDMs across subjects
```{r}
compute_mean_rdm <- function(rdms_df) {
  # Extract all RDMs as matrix
  rdm_matrix <- map(rdms_df$rdm, ~ .x) %>%
    do.call(rbind, .)
  
  # Compute column-wise mean
  colMeans(rdm_matrix)
}

# Compute mean RDMs for all combinations
mean_rdms <- map(ratings_type, function(rt) {
  map(distance_methods, function(dm) {
    compute_mean_rdm(all_rdms[[rt]][[dm]])
  }) %>%
    set_names(distance_methods)
}) %>%
  set_names(ratings_type)

cat("\nMean RDMs computed for all combinations\n")
```

# Convert tril vector to full matrix for plotting
```{r}
tril_to_matrix <- function(tril_vec) {
  n <- (1 + sqrt(1 + 8 * length(tril_vec))) / 2
  mat <- matrix(0, n, n)
  mat[lower.tri(mat)] <- tril_vec
  mat <- mat + t(mat)
  diag(mat) <- NA  # Set diagonal to NA for visualization
  return(mat)
}

# Get condition labels
condition_labels <- ratings_list$emotion %>% 
  filter(sub == subs[1]) %>% 
  pull(emotion)
```


# Plot comparison for each rating type
```{r, fig.width=15, fig.height=5}
for (rt in ratings_type) {
  
  cat("\n=== Plotting", rt, "RDMs ===\n")
  
  # Create matrices for each distance method
  matrices <- map(distance_methods, ~ tril_to_matrix(mean_rdms[[rt]][[.x]]))
  
  # Find global min/max for consistent color scale
  all_values <- unlist(matrices)
  vmin <- min(all_values, na.rm = TRUE)
  vmax <- max(all_values, na.rm = TRUE)
  
  color_map <- colorRamp2(
    c(vmin, (vmin + vmax) / 2, vmax),
    c("blue", "white", "red")
  )
  
  # Create heatmaps
  ht_list <- NULL
  for (i in seq_along(distance_methods)) {
    dm <- distance_methods[i]
    mat <- matrices[[i]]
    rownames(mat) <- condition_labels
    colnames(mat) <- condition_labels
    
    ht <- Heatmap(
      mat,
      name = paste(rt, dm, sep = "\n"),
      col = color_map,
      cluster_rows = FALSE,
      cluster_columns = FALSE,
      show_row_dend = FALSE,
      show_column_dend = FALSE,
      column_title = dm,
      row_names_side = "left",
      column_names_side = "top",
      column_names_rot = 60,
      row_names_gp = gpar(fontsize = 8),
      column_names_gp = gpar(fontsize = 8),
      width = unit(40, "mm"),
      height = unit(40, "mm"),
      rect_gp = gpar(col = "white", lwd = 0.5),
      heatmap_legend_param = list(
        title = "Distance",
        direction = "vertical"
      )
    )
    
    if (is.null(ht_list)) {
      ht_list <- ht
    } else {
      ht_list <- ht_list + ht
    }
  }
  
  # Draw combined heatmap
  draw(ht_list, 
       column_title = paste("Mean", rt, "RDMs - Distance Method Comparison"),
       column_title_gp = gpar(fontsize = 14, fontface = "bold"))
}
```

# Correlation between distance methods
```{r}
cat("\n=== Correlations Between Distance Methods ===\n\n")

for (rt in ratings_type) {
  cat("Rating type:", rt, "\n")
  
  # Create correlation matrix
  cor_mat <- matrix(1, nrow = 3, ncol = 3)
  rownames(cor_mat) <- distance_methods
  colnames(cor_mat) <- distance_methods
  
  for (i in 1:(length(distance_methods) - 1)) {
    for (j in (i + 1):length(distance_methods)) {
      r <- cor(mean_rdms[[rt]][[i]], mean_rdms[[rt]][[j]])
      cor_mat[i, j] <- r
      cor_mat[j, i] <- r
    }
  }
  
  print(round(cor_mat, 3))
  cat("\n")
}
```

# Summary statistics
```{r}
cat("\n=== Summary Statistics ===\n\n")

for (rt in ratings_type) {
  cat("Rating type:", rt, "\n")
  
  for (dm in distance_methods) {
    rdm_vec <- mean_rdms[[rt]][[dm]]
    cat(sprintf("  %s: mean=%.3f, sd=%.3f, min=%.3f, max=%.3f\n",
                dm,
                mean(rdm_vec, na.rm = TRUE),
                sd(rdm_vec, na.rm = TRUE),
                min(rdm_vec, na.rm = TRUE),
                max(rdm_vec, na.rm = TRUE)))
  }
  cat("\n")
}
```

# Conclusions

This analysis compares three distance metrics for behavioral RDMs:

1. **Euclidean**: Standard L2 distance
2. **Mahalanobis**: Accounts for covariance structure (scaled by inverse covariance)
3. **Cosine**: Measures angle between vectors (insensitive to magnitude)

**Key Observations**:
- Correlations between methods show how similar the resulting RDM structures are
- Different metrics may emphasize different aspects of the data
- For scalar ratings (arousal), Mahalanobis ≈ standardized Euclidean
- Cosine distance is scale-invariant, focusing on pattern similarity

**For RSA**: Since we correlate RDMs, constant scaling factors don't affect results. However, different metrics can capture different aspects of dissimilarity that may relate differently to neural RDMs.

**Note**: Residualization (if needed) happens at the RDM level in the main RSA pipeline, not at the ratings level. This script shows the raw RDM structure for each distance metric.
