---
title: "do_RSA_v15_CROSSNOBIS_ROIs_EHLP"
author: "LC"
date: "2026-02-01"
output: html_document
params:
  atlas_filename: test_ROI_crossnobis.nii.gz
  remove_neutrals: "YES"
  minus_neutral: "NO"
  subs_set: "N26"
  RSA_on_residuals: FALSE
  filter_RDMs: FALSE
  crossnobis_cov_method: "shrinkage"  # shrinkage or pseudoinverse
---


# V15 CROSSNOBIS
- This version implements crossnobis distance for fMRI RDM computation
- Based on V15 but uses first-level run-wise copes instead of 2nd-level copes
- Crossnobis: leave-one-run-out Mahalanobis distance across runs
- **Data structure**: 8 runs × 14 copes = 112 files per subject
- **Only fMRI RDM changes** - behavioral RDMs and RSA method stay identical to V15

**IMPORTANT**: This is a work-in-progress version.
Currently implements:
- df_path_copes creation from first-level directories
- Metadata addition (emotion labels)

NOT YET IMPLEMENTED:
- load_sub_copes_crossnobis
- DDOS_crossnobis
- do_RDM_fmri_crossnobis
- Full RSA pipeline

# Load libraries
```{r load_libraries, message=F}

library(future)
library(tidyverse)
library(furrr)
library(tictoc)
library(RNifti)
library(proxy) # distances
library(DT)
library(formattable)
library(janitor)
library(ComplexHeatmap)
library(circlize)  # For color mapping functions
library(grid)  # For gpar()
library(papayaWidget)
library(ggstatsplot)
library(rstatix)
library(corpcor)  # For shrinkage covariance estimation

source("funs_V15_CROSSNOBIS_ROIs.R")

```


# Copes/emotion dictionary specific for emotion_prediction
```{r}
metadata_copes <- tribble(
  ~cope, ~emotion,
  1, "anger_high",
  2, "anger_low",
  3, "disgust_high",
  4, "disgust_low",
  5, "fear_high",
  6, "fear_low",
  7, "happy_high",
  8, "happy_low",
  9, "neutral_high",
  10, "neutral_low",
  11, "pain_high",
  12, "pain_low",
  13, "sad_high",
  14, "sad_low"
)
```


# Options for different flavours
```{r}

# Choose one set of ROIs - e.g. an atlas - from ${bd_atlases} (ROIS_REPO)
atlas_filename <- params$atlas_filename

# Do you want to remove neutral cope(s) from the RSA analysis?
remove_neutrals <- params$remove_neutrals

RSA_on_residuals <- params$RSA_on_residuals

filter_RDMs <- params$filter_RDMs

```


# Check the params the Rmd is running
```{r}
# Create a unique temporary file (using UUID)
tmpfile <- file.path(tempdir(), paste0("rsa_params_", uuid::UUIDgenerate(), ".txt"))

# Prepare lines to write: empty line + title + actual parameters
lines_to_write <- c(
  "",  # empty line
  "Params being used by the Rmd:",
  paste(names(params), params, sep = "=")
)

# Write lines to the temp file
writeLines(lines_to_write, con = tmpfile)

# Print the file to terminal
system(paste("cat", shQuote(tmpfile)))

# Delete the temporary file immediately
unlink(tmpfile)
```


# Define paths and analysis-specific options
```{r}

bd <- "/data00/leonardo/RSA/analyses"
bd_ratings <- paste0(bd, "/RATINGS")
bd_atlases <- paste0(bd, "/ROIS_REPO")
bd_results <- paste0(bd, "/RSA_ROI_APP/results_RSA_ROI")

# ********** NB VALENCE REMOVED: WE USE ONLY AROVAL NOW **********
ratings_type <- c("emotion", "arousal", "aroval")
# ********** NB VALENCE REMOVED: WE USE ONLY AROVAL NOW **********

rsa_flavour <- "rsa_emotion_high_low_predictors_ROI"

# Choose the distance metric to use
# For crossnobis, dist_method_fmri will be "crossnobis"
dist_method_rating <- "euclidean"
dist_method_fmri <- "crossnobis"  # NEW: crossnobis distance
dist_method_rsa <- "pearson"

# Vector of zeropadded sub_ids
subs_file <- "/data00/leonardo/RSA/sub_list.txt"
subs <- sprintf("%02d", readLines(subs_file) %>% as.numeric)

# ---------- ONLY TOP RATERS BELOW ------------
if (params$subs_set == "N14") {
  subs <- c("02", "03", "12", "11", "22", "09", "29", "28", "26", "32", "23", "15", "20", "19")
}

# ---------- N25 : remove sub 07 --------------------
if (params$subs_set == "N25") {
  subs <- setdiff(subs, "07")
}

# ---------- N23 : remove subs 07,34,16  ------------
if (params$subs_set == "N23") {
  subs <- setdiff(subs, c("07", "34", "16"))
}

```


# CROSSNOBIS: Create df_path_copes from first-level directories
This is the main difference from V15. We now scan first-level directories
for run-wise cope files instead of using 2nd-level aggregated copes.

Expected structure:
`/data01/.../sub-XX/fmri/.../sub-XX_run-N_preproc_reg.feat/reg_standard/stats/copeM.nii.gz`

The resulting `df_path_copes` will have columns: sub, run, cope, path

```{r, message=FALSE}

# Base directory for first-level copes
bd_copes_1st_level <- "/data01/7T_Emotion_Insula/Data_analysis/RSA/prep_data"

# Import df_path_copes using the crossnobis-specific function
# Pass subs vector so it can construct paths directly
df_path_copes <- import_df_path_copes_crossnobis(
  bd_copes = bd_copes_1st_level,
  subs = subs,
  runs = 1:8,
  copes = 1:14
)

# Add metadata (emotion labels)
df_path_copes <- df_path_copes %>%
  inner_join(metadata_copes, by = "cope") %>%
  relocate(sub, run, cope, emotion, path)

# If remove_neutrals is "YES", remove neutral copes
if (remove_neutrals == "YES") {
  df_path_copes <- df_path_copes %>%
    filter(!str_detect(emotion, "neutral")) %>%
    arrange(sub, run, cope)
}

# Add label column (same as emotion for this analysis)
df_path_copes$label <- df_path_copes$emotion

# Get vectors of unique runs and copes
runs_numba <- df_path_copes$run %>% unique() %>% sort()
copes_numba <- df_path_copes$cope %>% unique() %>% sort()

# Summary
cat("\n=== DF_PATH_COPES SUMMARY ===\n")
cat("Total files:", nrow(df_path_copes), "\n")
cat("Subjects:", length(unique(df_path_copes$sub)), "\n")
cat("Runs:", paste(runs_numba, collapse = ", "), "\n")
cat("Copes:", paste(copes_numba, collapse = ", "), "\n")
cat("Files per subject:", nrow(df_path_copes %>% filter(sub == subs[1])), "\n")

```


# Display df_path_copes for validation
```{r}
cat("\n=== FIRST 20 ROWS ===\n")
df_path_copes %>% 
  head(20) %>% 
  datatable()
```


# Validation: Check one subject's files
```{r}
test_sub <- subs[1]

cat("\n=== FILES FOR SUBJECT", test_sub, "===\n")

df_test_sub <- df_path_copes %>% filter(sub == test_sub)

cat("Total files for this subject:", nrow(df_test_sub), "\n")
cat("Expected:", length(runs_numba), "runs ×", length(copes_numba), "copes =", 
    length(runs_numba) * length(copes_numba), "files\n")

# Check if all files exist
cat("\nChecking if files exist...\n")
files_exist <- file.exists(df_test_sub$path)
cat("Files that exist:", sum(files_exist), "/", length(files_exist), "\n")

if (!all(files_exist)) {
  cat("\nWARNING: Some files do not exist!\n")
  cat("Missing files:\n")
  print(df_test_sub$path[!files_exist])
} else {
  cat("✓ All files exist!\n")
}

# Show breakdown by run and cope
cat("\n=== BREAKDOWN BY RUN ===\n")
df_test_sub %>%
  group_by(run) %>%
  summarise(n_copes = n(), .groups = "drop") %>%
  print()

```


# Read atlas and create a vector of atlas labels
```{r}

atlas_path <- paste0(bd_atlases, "/", atlas_filename)

atlas_nii <- readNifti(atlas_path)
region_labels <- atlas_nii[atlas_nii > 0] %>% unique %>% sort

cat("Atlas:", atlas_filename, "\n")
cat("Number of regions:", length(region_labels), "\n")

```


# Calculate FMRI RDMs using CROSSNOBIS
**IMPORTANT:** This uses crossnobis distance (leave-one-run-out Mahalanobis)
instead of the euclidean distance used in V15.

```{r}

cat("\n=== STARTING CROSSNOBIS RDM CALCULATION ===\n")
cat("This will take some time...\n\n")

# Create progress log file
log_file <- paste0("crossnobis_progress_", 
                   gsub("\\.nii\\.gz$", "", atlas_filename), 
                   "_", format(Sys.time(), "%Y%m%d_%H%M%S"), 
                   ".log")

cat("Progress will be logged to:", log_file, "\n\n")

# Write log header
cat("=== CROSSNOBIS RDM CALCULATION LOG ===\n", file = log_file)
cat("Started:", format(Sys.time()), "\n", file = log_file, append = TRUE)
cat("Atlas:", atlas_filename, "\n", file = log_file, append = TRUE)
cat("Subjects:", length(subs), "\n", file = log_file, append = TRUE)
cat("ROIs:", length(region_labels), "\n", file = log_file, append = TRUE)
cat("Cov method:", params$crossnobis_cov_method, "\n", file = log_file, append = TRUE)
cat(rep("-", 60), "\n\n", file = log_file, append = TRUE, sep = "")

# Verify environment variables are set
cat("BLAS Threading Environment Variables:\n")
cat("  OPENBLAS_NUM_THREADS =", Sys.getenv("OPENBLAS_NUM_THREADS"), "\n")
cat("  MKL_NUM_THREADS =", Sys.getenv("MKL_NUM_THREADS"), "\n")
cat("  OMP_NUM_THREADS =", Sys.getenv("OMP_NUM_THREADS"), "\n\n")

# Use sequential processing
plan(sequential)

tic("Total time for all subjects")

RDMs_fmri <- subs %>% map_dfr(
  ~ {
    # Progress message
    msg_start <- paste0("[", format(Sys.time(), "%H:%M:%S"), "] Processing subject ", .x, "...")
    cat(msg_start, "\n")
    cat(msg_start, "\n", file = log_file, append = TRUE)
    
    tic_start <- Sys.time()
    
    result <- do_RDM_fmri_crossnobis(
      sub_id = .x,
      runs_numba = runs_numba,
      copes_numba = copes_numba,
      df_path_copes = df_path_copes,
      atlas_nii = atlas_nii,
      region_labels = region_labels,
      cov_method = params$crossnobis_cov_method
    )
    
    elapsed <- as.numeric(difftime(Sys.time(), tic_start, units = "secs"))
    msg_done <- paste0("  ✓ Subject ", .x, " completed in ", 
                      round(elapsed, 1), " sec (", 
                      round(elapsed/60, 1), " min)")
    cat(msg_done, "\n")
    cat(msg_done, "\n", file = log_file, append = TRUE)
    
    return(result)
  }
)

total_time <- toc()

# Write completion to log
cat("\n", rep("=", 60), "\n", file = log_file, append = TRUE, sep = "")
cat("COMPLETED:", format(Sys.time()), "\n", file = log_file, append = TRUE)
cat("Total time:", round(total_time$toc - total_time$tic, 1), "sec (", 
    round((total_time$toc - total_time$tic)/60, 1), "min)\n", 
    file = log_file, append = TRUE)
cat(rep("=", 60), "\n", file = log_file, append = TRUE, sep = "")

cat("\n=== CROSSNOBIS RDM CALCULATION COMPLETE ===\n")
cat("Total subjects:", length(unique(RDMs_fmri$sub)), "\n")
cat("Total ROIs:", length(unique(RDMs_fmri$roi)), "\n")
cat("Total RDMs:", nrow(RDMs_fmri), "\n")

```


# Display RDMs_fmri structure
```{r}
cat("\n=== RDMs_fmri STRUCTURE ===\n")
cat("Dimensions:", nrow(RDMs_fmri), "rows ×", ncol(RDMs_fmri), "columns\n")
cat("RDM vector length:", length(RDMs_fmri$rdm_fmri[[1]]), "\n")
cat("Expected for", length(copes_numba), "copes:", 
    length(copes_numba) * (length(copes_numba) - 1) / 2, "\n\n")

RDMs_fmri %>% 
  head(10) %>%
  datatable()
```


# Noise ceiling and floor
```{r}

# ref: [Nili2014](https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1003553#pcbi.1003553.s007) and also Supplmentary materials

# Upper bound (tight ceiling):
# 1. Take each subject's RDM → apply the correlation-appropriate transform
#    (e.g., z-transform for Pearson, rank-transform for Spearman).
# 2. Compute the average (centroid) RDM across subjects.
# 3. Correlate this average RDM with each subject's RDM, then average these correlations.
# → This overestimates the true fit (since the centroid is "fitted" to all data) → upper bound.

# Lower bound (floor):
# 1. Leave-one-subject-out (LOSO). For each subject:
#    - Correlate that subject's RDM with the average of all other subjects' RDMs
#      (after the same transform).
# 2. Average these correlations across subjects.
# → This underestimates the true fit (since each subject has less data to "anchor" the average) → lower bound.

# So:
# Upper bound = correlation(sub, centroid) averaged across subs.
# Lower bound = correlation(sub, mean-of-others) averaged across subs.

# This matches the intuition:
# - If your model correlation is below the floor, it's worse than what's expected from noisy data.
# - If it's between the floor and ceiling, the model explains some meaningful structure.
# - If it reaches the ceiling, it's as good as possible given the experimental noise.



# Step 1. Group data per ROI
df_noise <- RDMs_fmri %>%
  group_by(roi) %>%
  nest() %>%
  mutate(
    noise = map(data, ~ {
      roi_data <- .x

      # Step 2. Z-transform all subject RDMs
      # Each rdm_fmri is a numeric vector → standardize to mean 0, sd 1
      z_rdms <- map(roi_data$rdm_fmri, ~ as.numeric(scale(.x, center = TRUE, scale = TRUE)))

      # Step 3. Compute noise ceiling and floor for each subject
      map_dfr(seq_along(z_rdms), function(i) {
        sub_id  <- roi_data$sub[i]
        sub_rdm <- z_rdms[[i]]

        # (a) Lower bound (noise floor): mean of all *other* subjects
        mean_others <- reduce(z_rdms[-i], `+`) / (length(z_rdms) - 1)

        # (b) Upper bound (noise ceiling): mean of *all* subjects (centroid)
        mean_all <- reduce(z_rdms, `+`) / length(z_rdms)

        tibble(
          sub = sub_id,
          noise_floor   = cor(sub_rdm, mean_others),  # conservative bound
          noise_ceiling = cor(sub_rdm, mean_all)      # optimistic bound
        )
      })
    })
  ) %>%
  select(roi, noise) %>%
  unnest(noise) %>%
  ungroup()

# Step 4. Average across subjects per ROI → final summary
df_noise_summary <- df_noise %>%
  group_by(roi) %>%
  summarize(
    noise_floor   = mean(noise_floor, na.rm = TRUE),
    noise_ceiling = mean(noise_ceiling, na.rm = TRUE)
  ) %>%
  ungroup()

df_noise_summary


```


# Calculate RATINGS RDMs
```{r, message=FALSE}
# NB: ratings_type is now a *vector*, e.g.: 
# ratings_type <- c("emotion","arousal","aroval")
# to test the fn below for one rating type : do_RDMs_rats("emotion")

do_RDMs_rats <- function(ratings_type, remove_neutrals) {

  ratings_path <- paste0(bd_ratings,"/",ratings_type,"_ratings.csv")
  rats <- read_csv(ratings_path)
  
  
  # ------- IMPORTANT : AVG ACROSS MOVIES FOR EMOTION PREDICTORS MODEL -----
  rats <- rats %>% 
    select(sub, emotion, high_low_code, starts_with("r_")) %>% 
    # ------ specific for emotion_high_low_predictors -----------
    mutate(emotion = paste0(emotion, "_", str_match(high_low_code, "high|low")) ) %>% 
    select(-high_low_code) %>% 
    # ------ specific for emotion_high_low_predictors -----------
    group_by(sub, emotion) %>% 
    reframe(across(starts_with("r_"), median, na.rm = TRUE)) %>% 
    ungroup %>% 
    arrange(sub, emotion)
  
  # ------- IMPORTANT : AVG ACROSS MOVIES FOR EMOTION PREDICTORS MODEL -----
  
  
  if (remove_neutrals == "YES") {
    rats <- rats %>% filter(!str_detect(emotion, "neutral"))
  }
  
  rdm <- rats %>%
    filter(sub %in% subs) %>%
    select(sub, starts_with("r_")) %>%
    group_by(sub) %>%
    nest %>%
    mutate(!!paste0("RDM_",ratings_type) := data %>% map(~ DDOS_vec(.x, dist_method_rating))) %>%
    ungroup %>% 
    # remove the sub with the prospect of merging rdms of different ratings/models
    select(!data)
  
  return(rdm)
  
}

# calculate rdms for all ratings_type and return one column for each ratings_type
RDMs_rats <- map_dfc(ratings_type, ~ do_RDMs_rats(.x, remove_neutrals)) %>%
  # remove the duplicated sub_ columns (generated by do_RDMs_rats) 
  janitor::clean_names() %>% 
  mutate(sub = sub_1) %>% 
  select(sub, starts_with("rdm"))

```


# RM-ANOVA and three-ttests on each element of the RDM_rats
These will be used to select the elements of (residualized) fmri and behavioural rdms

```{r}

library(rstatix)

# set the threshold for the p value below which we consider the results
# significant, both for RM-ANOVA and for the three paired ttest
p_thresh <- 0.001

# fn to transform the RDM values to z scores
zscore_RDMs <- function(RDMs) {
  RDMs %>%
    mutate(
      rdm_emotion = map(rdm_emotion, ~ as.numeric(scale(.x))),
      rdm_arousal = map(rdm_arousal, ~ as.numeric(scale(.x))),
      rdm_aroval  = map(rdm_aroval,  ~ as.numeric(scale(.x)))
    )
}

RDMs_rats_Z <- zscore_RDMs(RDMs_rats);


# reshape to long format for RM-ANOVA
RDMs_rats_long <- RDMs_rats %>%
  mutate(idx = map(rdm_emotion, ~ seq_along(.x))) %>%  # index for each element
  unnest(c(rdm_emotion, rdm_arousal, rdm_aroval, idx)) %>%   # expand all list-columns %>%
  pivot_longer(
    cols = starts_with("rdm_"),
    names_to = "RDM_type",
    values_to = "value"
  )


# ------------ RM-ANOVA ----------------

# nest by idx and calculate RM-ANOVA
RDMs_RM_ANOVA <- RDMs_rats_long %>%
  group_by(idx) %>%
  nest() %>% 
  
  # calculate RM-ANOVA per idx
  mutate(rm_anova = map(data, ~ anova_test(data = .x,
                                           dv = value,
                                           wid = sub,
                                           within = RDM_type)))


# Get the results
RDMs_RM_ANOVA_results <- RDMs_RM_ANOVA %>%
  mutate(rm_anova_tbl = map(rm_anova, ~ .x$ANOVA)) %>%  # extract the inner tibble
  select(idx, rm_anova_tbl) %>%
  unnest(rm_anova_tbl) %>%
  select(-c(Effect), RDM_element = idx)


RDMs_RM_ANOVA_results %>% 
  filter(p < p_thresh)


# ----------------- 3 paired ttests --------------------

RDMs_RM_ttest_results <- RDMs_rats_long %>%
  group_by(idx) %>%
  nest() %>%
  mutate(
    pvals = map(data, ~ tibble(
      emo_aro    = t.test(.x$value[.x$RDM_type=="rdm_emotion"], 
                           .x$value[.x$RDM_type=="rdm_arousal"], 
                           paired = TRUE)$p.value,
      emo_aroval = t.test(.x$value[.x$RDM_type=="rdm_emotion"], 
                           .x$value[.x$RDM_type=="rdm_aroval"], 
                           paired = TRUE)$p.value,
      aro_aroval = t.test(.x$value[.x$RDM_type=="rdm_arousal"], 
                           .x$value[.x$RDM_type=="rdm_aroval"], 
                           paired = TRUE)$p.value
    ))
  ) %>%
  select(RDM_element = idx, pvals) %>%
  unnest(pvals) %>% 
  # create a column showing where all three ttest are sig
  mutate(sig = if_else(emo_aro < p_thresh & emo_aroval < p_thresh & aro_aroval < p_thresh, 1, 0))

RDMs_RM_ttest_results %>% 
  filter(sig == 1)


idx_sig_RM_ANOVA = which(RDMs_RM_ANOVA_results$p < p_thresh)
idx_sig_ttest = which(RDMs_RM_ttest_results$sig == 1)

shared_idx <- intersect(idx_sig_RM_ANOVA, idx_sig_ttest)

cat(
  'After filtering at p < ', p_thresh, '  ', 
  length(shared_idx), '/', length(RDMs_rats$rdm_emotion[[1]]), 
  'rdm elements were retained'
)

```


# Calculate REGULAR RSA
```{r}
do_regular_RSA <- function(RDMs_fmri, RDMs_rats, ratings_type, dist_method_rsa = "spearman") {
  
  cat('Carrying out Regular RSA with original rdm \n')
  
  rats_rdm_colnames <- paste0("rdm_", ratings_type)
  
  RSA <- right_join(RDMs_fmri, RDMs_rats, by = "sub") %>%
    group_by(sub) %>%
    mutate(
      across(
        all_of(rats_rdm_colnames),
        ~ map2_dbl(.x, rdm_fmri, ~ cor(.x, .y, method = dist_method_rsa)),
        .names = "rsa_{.col}"
      )
    ) %>%
    mutate(across(starts_with("rsa_"), round, 3)) %>%
    ungroup()
  
  return(RSA)
}

```


# Calculate RSA RESIDUALS
```{r}
do_RSA_on_residuals <- function(RDMs_fmri, RDMs_rats, ratings_type, dist_method_rsa = "spearman") {
  
  cat('Carrying out RSA on residual rdm \n')
  
  rats_rdm_colnames <- paste0("rdm_", ratings_type)
  
  RSA <- right_join(RDMs_fmri, RDMs_rats, by = "sub") %>%
    group_by(sub, roi) %>% 
    mutate(
      across(
        all_of(rats_rdm_colnames),
        
        ~ map2_dbl(.x, rdm_fmri, function(behav_vec, fmri_vec) {
          
          regressors_colnames <- setdiff(rats_rdm_colnames, cur_column())
          
          regressors_mat <- map(regressors_colnames, ~ get(.x)[[1]]) %>% 
            set_names(regressors_colnames) %>% 
            bind_cols() %>% 
            as.matrix()
          
          fmri_res  <- resid(glm(fmri_vec ~ regressors_mat))
          behav_res <- resid(glm(behav_vec ~ regressors_mat))
          
          cor(fmri_res, behav_res, method = dist_method_rsa)
        }),
        .names = "rsa_{.col}"
      )
    ) %>%
    ungroup()
  
  return(RSA)
}

```


# Compute the RSA
```{r}

idx_sig <- idx_sig_ttest

if (params$filter_RDMs) {
  cat('Filtering the rdm according to the most discriminative ones. \n')
  
  RDMs_rats <- RDMs_rats %>%
    mutate(
      rdm_emotion = map(rdm_emotion, ~ .x[idx_sig]),
      rdm_arousal = map(rdm_arousal, ~ .x[idx_sig]),
      rdm_aroval  = map(rdm_aroval,  ~ .x[idx_sig])
    )
  
  RDMs_fmri <- RDMs_fmri %>%
    mutate(
      rdm_fmri = map(rdm_fmri, ~ .x[idx_sig])
    )
}


if (params$RSA_on_residuals) {
  
  cat("Running RSA on residuals\n")
  RSA <- do_RSA_on_residuals(
    RDMs_fmri, RDMs_rats, ratings_type, dist_method_rsa = "pearson"
  )
} else {
  
  cat("Running regular RSA\n")
  RSA <- do_regular_RSA(
    RDMs_fmri, RDMs_rats, ratings_type, dist_method_rsa = "pearson"
  )
}

```


# Get the mean RSA for each ratings_type
```{r}
RSA_mean <- RSA %>%
  select(roi, starts_with("rsa_")) %>%
  group_by(roi) %>%
  reframe(
    across(
      starts_with("rsa_"),
      list(mean = ~ round(mean(.x, na.rm = TRUE), 2))
    )
  ) %>% 
  rename_with(
    ~ str_replace(.x, "rsa_rdm_", ""),
    starts_with("rsa_")
  )

RSA_mean %>% datatable()
```


# Calculate mean tril for ratings and for fmri
```{r}
get_elementwise_mean <- function(mydf, nested_col, fun=mean) {
  m <- matrix(
    mydf[[nested_col]] %>% unlist,
    nrow = nrow(mydf),
    ncol = length(mydf[[nested_col]][[1]]),
    byrow = TRUE
  )
  
  m_summary <- apply(m, 2, fun)
  return(m_summary)
}


# mean RDMs_rats
mean_RDMs_rats <- RDMs_rats %>% select(starts_with("rdm_")) %>% colnames %>%  
  map_dfc(~{
    mean_tril <- get_elementwise_mean(RDMs_rats, .x, fun = mean)
    tibble(!!.x := mean_tril)
  })



# mean RDMs_fmri
plan(multisession, workers = 5)

mean_RDMs_fmri <- RDMs_fmri$roi %>% unique %>% 
  future_map_dfr(~{
    mean_tril <- get_elementwise_mean(RDMs_fmri %>% filter(roi == .x), "rdm_fmri")
    tibble(roi = .x, mean_tril = mean_tril)
  }) %>% 
  group_by(roi) %>% 
  nest

plan(sequential)
```


# Create the volumes with the RSA_mean
```{r}

RSA_volumes_df <- ratings_type %>% map_dfc(
  ~ {
    nii <- atlas_nii %>% as.vector()
    
    for (ROI_numba in RSA_mean$roi) {
      idx_ROI <- which(nii == ROI_numba)
      val_ROI <- RSA_mean %>% filter(roi == ROI_numba) %>% select(starts_with(.x)) %>% pull
      nii[idx_ROI] = val_ROI
    }
    return( tibble(!! .x := nii) )
  }
)

```


# Save results in RSA_ROI_APP/results_RSA_ROI
```{r}

results_file <- create_filename_results(
    atlas_filename, 
    subs, 
    RSA_on_residuals, 
    filter_RDMs,
    dist_method_rating,
    dist_method_fmri,
    dist_method_rsa,
    bd_results
)


cat("Results will be saved to:\n")
cat(results_file, "\n\n")

# save the results
save(
  RSA, 
  RSA_mean, 
  atlas_filename, 
  mean_RDMs_fmri, 
  mean_RDMs_rats, 
  df_path_copes, 
  subs, 
  RSA_volumes_df,
  df_noise_summary,
  file = results_file
)

cat("Results saved successfully!\n")

```
