---
title: "RSA v1"
output: html_notebook
---

```{r load_libraries, message=F}
library(tidyverse)
library(future)
library(furrr)
library(tictoc)
library(RNifti)
library(proxy) # distances
library(profvis)


model="one_ev_per_movie"
atlas_filename <-  "juelich_2mm.nii.gz"
ratings_type <- "emotion"

# base directory
bd = "/data00/leonardo/RSA/analyses"
bd_rsa = paste0(bd,"/rsa")

bd_atlases = paste0(bd_rsa,"/atlases")

bd_ratings = "/data00/leonardo/RSA/analyses/RATINGS"

bd_copes = paste0(bd, "/", model, "/results/2nd_level")

subs_file <- "/data00/leonardo/RSA/sub_list.txt"

ratings_path <- paste0(bd_ratings,"/",ratings_type,"_ratings.csv")

read_csv(ratings_path)

```


# Aux functions
```{r}

# plot_heatmap(D %>% as.matrix)
plot_heatmap <- function(M) {
  heatmap(M, Rowv = NA, Colv = NA, symm=T, revC = T)
}

```


# Pathnames of the cope files
Define a df with the pathname of all copes using the `list.files()` function.
NB: The cope numbers in the `cope` column are not-zeropadded since this is how they come out from FSL Feat

```{r create_dizio_data, message=FALSE}


subs <- sprintf("%02d",readLines(subs_file) %>% as.numeric)

# ----- EXPERIMENT WITH ONLY THREE SUBS ------------
# subs <- subs[1:3]
# ----- EXPERIMENT WITH ONLY THREE SUBS ------------

bd_copes = paste0(bd, "/", model, "/results/2nd_level")

df_path_copes <- list.files(bd_copes, recursive = T) %>% tibble()
names(df_path_copes) <- "fname"

df_path_copes <- df_path_copes %>%
  filter(str_detect(fname, "gfeat/cope[0-9]+.feat/stats/cope")) %>%
  tidyr::separate(
    fname, c("sub","cope","tmp1","tmp2"),
    sep = "/", fill = "right", remove = F
  ) %>% 
  select(!starts_with("tmp")) %>% 
  mutate(sub = str_extract(sub, "[0-9]+")) %>% 
  mutate(cope = (str_extract(cope, "[0-9]+") %>% as.numeric) ) %>% 
  mutate(path = paste0(bd_copes,"/",fname)) %>% 
  select(!fname) %>% 
  arrange(sub,cope)

df_path_copes

```


# Load copes
Very fast: reads 56 nifti at 2mm resolution in about 1 sec
```{r load_copes}
sub_id = "02"
copes <- df_path_copes$cope %>% unique

load_sub_copes <- function(sub_id, copes) {
  copes %>% map_dfc(~ {
    nii <- df_path_copes %>% 
      filter(sub == sub_id, cope == .x) %>% pull %>%  
      readNifti %>% as.vector
    tibble(!!paste0("cope_", .x) := nii)
  }) 
}

tic()
df_copes <- load_sub_copes(sub_id, copes)
toc()

```


# Read atlas and calculate fmri RDMs
```{r}

atlas_path <- paste0(bd_atlases,"/",atlas_filename)

calculate_fmri_RDMs <- function(df_copes, atlas_path) {
  
  # Load the atlas and define a vector with region_labels
  atlas_nii <- readNifti(paste0(bd_rsa,"/atlases/",atlas_filename))
  region_labels <- atlas_nii[atlas_nii > 0] %>% unique %>% sort
  
  # Calculate the RDM for each atlas region, 
  # i.e. distance matrix of the betas of each movie across all voxels in that ROI.
  # Returns a df of length(triu(D))-by-n_atlas_regions
  region_labels %>% map_dfc(~ {
    idx <- which(atlas_nii == .x)
    df_copes_region <- df_copes[idx, ]
    
    # The following line will be replaced by the calculate_distance() fn
    D <- dist( t(df_copes_region), method = "euclidean" )
    
    D <- D[!is.na(D)]
    tibble(!!paste0("RDM_region_", .x) := D)
  })
}

tic()
RDMs <- calculate_fmri_RDMs(df_copes, atlas_path)
toc()

# RDMs

```



# Calculate ratings RDM
```{r}

```






















# Try proxy for distances

```{r, eval=F}
library(proxy)
# pr_DB$get_entries() 
# summary(pr_DB)


# df_region <- df[idx,]
# D <- dist( t(df_region) )
# dim(as.matrix(D))
# # plot_heatmap(D %>% as.matrix)
# D[!is.na(D)]

df_region <- df[idx,]

D <- dist( t(df_region) )
D <- dist(t(df_region), method = "euclidean")

dim(as.matrix(D))
plot_heatmap(D %>% as.matrix)
# D[!is.na(D)]

heatmap(D %>% as.matrix, symm = T)

types = c("angular","correlation","cosine","Euclidean") 
par(mfrow=c(2,2))
for (i in types) {
  D = proxy::dist(t(df_region), method = i)
  image(D %>% as.matrix, axes = F, asp=1)
  title(i)
}

```


# Explore proxy distances package
```{r, eval=F}

library(proxy)
# pr_DB$get_entries() 
# summary(pr_DB)

set.seed(124)

M = matrix(runif(100*10), ncol = 100)

method <- "correlation"

Dstats <- dist(M, method = method)
Dproxy <- proxy::dist(M, method = method)

sum(Dstats-Dproxy)


par(mfrow = c(1,2))
image(Dstats %>% as.matrix()); title("pippo")
image(Dproxy %>% as.matrix()); title("topolino")



```


















