---
title: "do_RSA_v9"
author: "LC"
date: "2024-08-10"
output: html_document
---

# Load libraries and define parameters

NB: this Rmd can be used for all types of ratings: emotion, arousal, valence

Abbreviations:
- sub : subjects(s)
- rat : rating(s)

```{r load_libraries, message=F}
library(tidyverse)
library(future)
library(furrr)
library(tictoc)
library(RNifti)
library(proxy) # distances
library(profvis)
library(DT)
library(formattable)

source("funs_V9_ROIs.R")

bd="/data00/leonardo/RSA/analyses"

rsa_flavour="rsa_ROI"

# can be emotion, arousal, valence
ratings_type <- "emotion"

# the following is our contrast of interest and it's always the same
copes_type <- "one_ev_per_movie"


# Choose one atlas among:
# Schaefer100_2mm.nii.gz
# Schaefer200_2mm.nii.gz
atlas_filename <- "Schaefer100_2mm.nii.gz"


# Choose the distance metric to use for fmri RDM
# The metric for ratings RDM *should be* euclidean, since arousal and valence
# ratings have only one value
# Supported methods: 'pearson', 'spearman', 'euclidean', 'cosine' or 'mahalanobis'
dist_method_rating <- "euclidean"
dist_method_fmri = "euclidean"
dist_method_rsa = "pearson"

cat(paste0("Ratings RDM to be calculated with : ", dist_method_rating, "\n"))
cat(paste0("fMRI RDM to be calculated with : ", dist_method_fmri, "\n"))


# for RDM_type use:
# - tril for the lower triangular of the D matrix
RDM_type = "tril"

# Vector of zeropadded sub_ids
subs_file <- "/data00/leonardo/RSA/sub_list.txt"
subs <- sprintf("%02d",readLines(subs_file) %>% as.numeric)

# ---------- ONLY TOP RATERS BELOW ------------
subs <- c("02","03","12","11","22","09","29","28","26","32","23","15","20","19")

```




# Aux functions
```{r}
# plot_heatmap(D %>% as.matrix)
plot_heatmap <- function(M) {
  heatmap(M, Rowv = NA, Colv = NA, symm=T, revC = T)
}
```


# Create a df_path_copes with the location of the 56 cope niis from the one_movie_per_ev model
Extract the pathname of all copes using the `list.files()` function.
Also define a copes_numba vector with all the copes numbers.

NB: The cope numbers in the `cope` column are NOT zeropadded since this is how they come out from FSL Feat
```{r, message=FALSE}

copes_location_csv <- paste0(bd,"/",rsa_flavour,"/copes_location.csv")

df_path_copes <- if (file.exists(copes_location_csv) ) {
  read_csv(copes_location_csv)
} else {
  import_df_path_copes(bd, copes_type, rsa_flavour)
}

copes_numba <- df_path_copes$cope %>% unique

# df_path_copes

```


# Read atlas and create a vector of atlas labels
```{r}

bd_atlases = paste0("/data00/leonardo/RSA/analyses/",rsa_flavour,"/atlases")

atlas_path <- paste0(bd_atlases,"/",atlas_filename)

atlas_nii <- readNifti(paste0(bd_atlases,"/",atlas_filename))
region_labels <- atlas_nii[atlas_nii > 0] %>% unique %>% sort

```



# Calculate RATINGS RDMs -- NEEDS UPDATE TO CALCULATE FOR EVERY RATINGS_TYPE
```{r, message=FALSE}

bd_ratings = paste0(bd,"/RATINGS")
ratings_path <- paste0(bd_ratings,"/",ratings_type,"_ratings.csv")
rats <- read_csv(ratings_path)


RDMs_rats <- rats %>%
  filter(sub %in% subs) %>%
  select(sub, starts_with("r_")) %>%
  group_by(sub) %>%
  nest %>%
  mutate(!!paste0("RDM_",ratings_type) := data %>% map(~ DDOS_vec(.x, dist_method_rating))) %>%
  select(!data)


# one_tril <- RDMs_rats[1,]$RDM_emotion
# plot_tril(one_tril, length(copes_numba), reord = "YES")

```





# Calculate FMRI RDMs

## For one sub - for dev purposes
```{r, eval=FALSE}

sub_id = "03"  # subs[1]

# load all the copes nii.gz into a df
df_copes <- load_sub_copes(sub_id, copes_numba, df_path_copes)

# the code below does the calculation of the tril of RDM_fmri for all rois for one sub

one_sub_RDM_fmri <- tibble(
  sub = sub_id,
  roi = region_labels
) %>%

  # extract idx_roi for atlas voxels in that roi
  mutate(idx_roi = roi %>% map(~ which(atlas_nii == .x)) ) %>%
  
  # extract the df_copes for the voxels in that region (idx_roi)
  mutate(df_copes_region = idx_roi %>% map(~ df_copes[.x,]) ) %>%
  
  # calculate tril of RDM_roi (output as numeric vector)
  mutate(RDM_roi = df_copes_region %>% map(~ DDOS_vec(t(.x), dist_method_fmri)) ) %>% 
  
  # select only sub, roi (numba) and RDM_roi
  select(sub, roi, RDM_roi)
  
```



## For multiple subs - furrred
This blows my mind: ~10 seconds for 14 subs (!)
```{r}

plan(multisession, workers = 5)

RDMs_fmri <- future_map_dfr(
  
  subs, ~ {
    paste0("Calculating RDMs for sub ",.x,"\n") %>% cat
  
    # 1. load all the copes nii.gz into a df
    df_copes <- load_sub_copes(.x, copes_numba, df_path_copes)
    
    # 2. calculate RDM fmri for each roi
    tibble(
      sub = .x,
      roi = region_labels
    ) %>% 

      # extract idx_roi for atlas voxels in that roi
      mutate(idx_roi = roi %>% map(~ which(atlas_nii == .x)) ) %>%
      
      # extract the df_copes for the voxels in that region (idx_roi)
      mutate(df_copes_region = idx_roi %>% map(~ df_copes[.x,]) ) %>%
      
      # calculate tril of RDM_roi (output as numeric vector)
      mutate(RDM_roi = df_copes_region %>% map(~ DDOS_vec(t(.x), dist_method_fmri)) ) %>% 
      
      # select only sub, roi (numba) and RDM_roi
      select(sub, roi, RDM_roi)
    
  }
)

plan(sequential)

# RDMs_fmri %>%
#   group_by(sub) %>%
#   group_split()

```



# RSA
Basically right-join the RDMs_fmri with the RDMs_rating for each sub in order to create a copy of the
rating RDM for each roi.
Then calculate RSA using pmap
```{r}
# can be pearson, spearman
dist_method_rsa <- "spearman"

# join fmri and rats RDMs and create a copy of the rat RDM for each sub/roi
RSA <- right_join(RDMs_fmri,RDMs_rats,by = "sub") %>% 
  group_by(sub) %>% 
  # calculate RSA
  mutate(
    similarity = list(RDM_roi, RDM_emotion) %>% pmap_dbl(~ cor(.x,.y, method = dist_method_rsa))
  ) %>%
  mutate(similarity = round(similarity,3)) %>% 
  ungroup


# Get the mean RSA for each roi
RSA_mean <- RSA %>% 
  select(roi, similarity) %>% 
  group_by(roi) %>% 
  reframe(
    mean_similarity = round(mean(similarity),2),
    sd_similarity = round(sd(similarity),2)
  ) %>% 
  arrange(desc(mean_similarity))

RSA_mean %>% datatable()

```




# TOBEDONE -- Write the map with the values -- TOBEDONE
```{r, eval=FALSE}

results_to_map <- RSA %>% 
  mutate(region = as.integer(str_replace(region, "RDM_region_",""))) %>% 
  select(region, mean_similarity)

results_to_map

# results_path <- ("/data00/leonardo/RSA/analyses/rsa/rsa_results")
# filename_results_nii <- paste0("rsa_results/rsa_results_",ratings_type,"_",atlas_filename)


results_path <- "/data00/leonardo/RSA/analyses/rsa/rsa_results"

results_filename <- create_filename_results(
  atlas_filename, ratings_type, dist_method_rating, dist_method_fmri, dist_method_rsa
)


# Function to write the results to a nii.gz volume
# (purrr gives issues here)
write_results_to_nii <- function(results_to_map, filename_results_nii, atlas_nii) {
  
  results_nii = atlas_nii
  
  for (ROI_numba in region_labels) {
    cat(paste0(ROI_numba," "))
    row_region <- which(results_to_map$region == ROI_numba)
    value <- results_to_map$mean_similarity[row_region]
    
    idx <- which(results_nii == ROI_numba)
    results_nii[idx] = value
  }

  writeNifti(results_nii, filename_results_nii)
}


# Write the results nii.gz in the rsa_results dir
write_results_to_nii(
  results_to_map, 
  paste0(results_path, "/", results_filename), 
  atlas_nii
)

```








