---
title: "Prepare movie_encodings.csv"
output: html_notebook
---

Organization of the movies:

- 6 types of emotions are displayed, +1 with no apparent emotion (neutral).
- each of these 7 emotional states (including the neutral) is displayed in two movies, one with low and one with high intensity
- each combination of emotion/intensity is displayed by one of 4 actors
- so in total there are 7 * 2 * 4 = 56 movies

The initial list has two columns, one for the original movie code (`original_code` column - e.g. `AL_N_06`) and one which shows whether the emotion in that movie was displayed with high or low intensity (`high_low_code` column - e.g. `AL_N_high`). 

The correspondence of each movie with the relative high/low intensity is evident when watching the movie. This is how the `movie_encodings_INITIAL.csv` was generated.

We now enrich this initial table with 3 more columns. 

- one for the `movie_cope_number` : this corresponds to the row number after alphabetically sorting the table by the `high_low_code` column. This column was chosen wrt to the `original_code` column to yield easier interpretability of the fmri results

- one for the `actor` code : for potential future use

- one for the `emotion` : where the emotion (or neutral) code in the original movie code is spelled out. Again we might need this for easier interpretability of the fmri results and for a preliminary analysis of the correspondence between the movie type and the max rating across emotions. 


```{r, message=FALSE}
library(tidyverse)
```


```{r, message=FALSE}
df_initial <- read_csv("movie_encodings_INITIAL.csv")

df <- df_initial %>%
  
  # Assign a movie number according to the alphabetical order of the high_low column
  # This will be used to number the copes in the fmri analysis
  arrange(high_low_code) %>% 
  mutate(movie_cope_number = row_number()) %>% 
  
  # Create a column where the emotion represented in that movie is spelled out
  mutate(tmp_original_code = original_code) %>% 
  separate(
    tmp_original_code, into = c("actor", "emotion", "movie_recording_number"), sep = "_"
  ) %>% 
  mutate(emotion = case_when(
   emotion == "A" ~ "anger",
   emotion == "D" ~ "disgust",
   emotion == "F" ~ "fear",
   emotion == "H" ~ "happy",
   emotion == "N" ~ "neutral",
   emotion == "P" ~ "pain",
   emotion == "S" ~ "sad",
   TRUE ~ as.character(emotion)  # If none of the above, keep the original value
  )) %>% 
  select(-movie_recording_number)

df

```

```{r}
write_csv(df, "movie_encodings.csv")
```

