---
title: "do_RSA_v8_Searchlight"
author: "LC"
date: "2024-04-30"
output: html_document
---

# Load libraries and define parameters

This script is a modification to use a searchlight, instead of a parcellation/atlas.

NB: this Rmd can be used for all types of ratings: emotion, arousal, valence

Abbreviations:
- sub : subjects(s)
- rat : rating(s)

```{r load_libraries, message=F}
library(tidyverse)
library(future)
library(furrr)
library(tictoc)
library(RNifti)
library(proxy) # distances
library(profvis)
library(DT)
library(formattable)

source("funs_V8_Searchlight.R")

bd="/data00/leonardo/RSA/analyses"

# can be emotion, arousal, valence
ratings_type <- "emotion"

# the following is our contrast of interest and it's always the same
copes_type <- "one_ev_per_movie"



# Point to which gray matter mask to use for searchlight
# GM_clean
# insula_HO_GM
gm_file <- "GM_clean"
gm_path <- paste0(bd,"/rsa/masks/", gm_file, ".nii.gz")

# Define searchlight radius in mm
r_mm <- 4


# Choose the distance metric to use for fmri RDM
# The metric for ratings RDM *should be* euclidean, since arousal and valence
# ratings have only one value
# Supported methods: 'pearson', 'spearman', 'euclidean', 'cosine' or 'mahalanobis'
dist_method_rating <- "euclidean"
dist_method_fmri = "euclidean"
dist_method_rsa = "pearson"

cat(paste0("Ratings RDM to be calculated with : ", dist_method_rating, "\n"))
cat(paste0("fMRI RDM to be calculated with : ", dist_method_fmri, "\n"))


# for RDM_type use:
# - tril for the lower triangular of the D matrix
# - svd for the first 3 pc from svd of the D matrix
# - in case of svd, set the number of components to return with ncomp_svd
RDM_type = "tril"  # svd or tril. Must be the same for both rating and fmri
ncomp_svd = 3      

# Vector of zeropadded sub_ids
subs_file <- "/data00/leonardo/RSA/sub_list.txt"
subs <- sprintf("%02d",readLines(subs_file) %>% as.numeric)

# ---------- ONLY TOP RATERS BELOW ------------
subs <- c("02","03","12","11","22","09","29","28","26","32","23","15","20","19")

```




# Aux functions
```{r}
# plot_heatmap(D %>% as.matrix)
plot_heatmap <- function(M) {
  heatmap(M, Rowv = NA, Colv = NA, symm=T, revC = T)
}
```


# Create a df_path_copes with the location of the 56 cope niis from the one_movie_per_ev model
Extract the pathname of all copes using the `list.files()` function.
Also define a copes_numba vector with all the copes numbers.

NB: The cope numbers in the `cope` column are NOT zeropadded since this is how they come out from FSL Feat
```{r, message=FALSE}

copes_location_csv <- paste0(bd,"/rsa/copes_location.csv")

df_path_copes <- if (file.exists(copes_location_csv) ) {
  df_path_copes <- read_csv(copes_location_csv)
} else {
  df_path_copes <- import_df_path_copes(bd, copes_type)
}

copes_numba <- df_path_copes$cope %>% unique

# df_path_copes

```





# Calculate RATINGS RDMs
We do this first since the RDM for the ratings will be the same for whatever
atlas will be used. Also, it's much faster than the RDM_fmri and can be done
for all the sub_ids at once.


- **tril** : The final RDMs_rats has 40040 nrows : 26 sub_ids * 1540, where the latter
derives from ((56^2)-56)/2 = 1540, i.e. tril from the D matrix of 56 movies
- **svd** : The final RDMs_rats has 4368 nrows : 26 sub_ids * 168, where the latter
derives from 56*3, i.e. the first three components from svd of the D matrix
```{r, message=FALSE}

bd_ratings = paste0(bd,"/RATINGS")
ratings_path <- paste0(bd_ratings,"/",ratings_type,"_ratings.csv")
rats <- read_csv(ratings_path)

# cols are subs
# rows are elements of the tril
RDMs_rats_wider <- subs %>% map_dfc(~ {
  do_RDM_ratings_one_sub(.x, dist_method_rating, rats, RDM_type =  RDM_type)
})

# pivot the df above so that there is only one col for subs
# and one for the values of the tril for each sub
RDMs_rats <- RDMs_rats_wider %>%
  pivot_longer(cols = everything(), names_to = "sub") %>%
  arrange(sub)

RDMs_rats
```


# Calculate fMRI RDMs

## Create SEARCHLIGHT regions
```{r}

# 1. Read the gm mask and get voxel size, get idx of voxels in gm and their
#    xyz coordinates
gm <- readNifti(gm_path)
voxel_size <- pixdim(gm) %>% mean   # voxel size of gm
idx <- which(gm != 0)   # index of nonzero voxels
coords <- which(gm != 0, arr.ind = TRUE)   # get coordinates


# 2. Define a searchlight function returning the volume (92x109x92) idx of the 
#    voxels within the searchlight
searchlight <- function(center_coord, r_mm, coords, idx, voxel_size) {

  # Calculate distance of the given voxel from every other voxel, considering voxel size
  distances <- sqrt(rowSums((t(t(coords) - center_coord) * voxel_size)^2))
  
  # # TEST ONLY: return the coords of the voxels in the searchlight
  # return(coords[distances <= r_mm, , drop = FALSE])

  # return the idx of the voxels in the searchlight
  return(idx[distances <= r_mm])
}


# # Testing the searchlight fn with one coordinate 
# # See also the nb : earchlight_development.Rmd
# mycoord <- coords[1,]
# idx_within_radius <- searchlight(mycoord, r_mm=6, coords, idx, voxel_size)

# # Convert indices to coordinates (just to check)
# mycoord
# coords_within_radius <- coords[match(idx_within_radius, idx), , drop = FALSE]
# coords_within_radius


# 3. Get the idx of the vxls within radius for each voxel in gm (~ 20 sec)
plan(multisession, workers = 15)

tic()

idx_within_radius <- 1:length(idx) %>% future_map(
  ~ searchlight(coords[.x,], r_mm, coords, idx, voxel_size)
)

toc()

plan(sequential)

idx_within_radius %>% map_dbl(length) %>% hist


```



## Fn to carry out the RMDs Searchlight calculation for one sub

```{r, message=FALSE}
sub_id = "02"
df_copes <- load_sub_copes(sub_id, copes_numba, df_path_copes)


calculate_fmri_RDMs_Searchlight <- function(df_copes, idx, idx_within_radius) {
  
  one_rdm <- idx_within_radius %>% map_dfc(~ {
    idx_spot <- unlist(.x)
    df_copes_region <- df_copes[idx_spot, ]
    
    # center (i.e. demean) variables
    # (treat also the case of single voxels)
    if(nrow(df_copes_region) > 1) {
      df_copes_region_demean <- df_copes_region %>% 
        mutate(across(everything(), ~ .x - mean(.x)))  
    } else {
      df_copes_region_demean <- df_copes_region
    }
    
    # calculate distance / similarity
    D_fmri <- DDOS( t(df_copes_region_demean), method = dist_method_fmri )
    D_fmri[is.na(D_fmri)] <- 0 
    
    # flat onto a vector
    D_feature_vector <- D_fmri[!is.na(D_fmri)]
  })
  
  colnames(one_rdm) = idx
  
  return(one_rdm)
}

# ~ 90 seconds, no need to use furrr
N <- 10 #length(idx)
one_RDM_fmri <- calculate_fmri_RDMs_Searchlight(df_copes, idx[1:N], idx_within_radius[1:N])

one_RDM_fmri

```



## Furrr the RDMs Searchlight calculation across subs
```{r, message=FALSE}

# for testing on a subset of idx
N = length(idx)

# 5 workers give the best performance
plan(multisession, workers = 10)

tic()
RDMs_fmri <- subs %>% future_map_dfr(~{
  
  paste0("Calculating RDMs Searchlight for sub ",.x,"\n") %>% cat
  
  df_copes <- load_sub_copes(.x, copes_numba, df_path_copes)
  
  one_RDM_fmri <- calculate_fmri_RDMs_Searchlight(
    df_copes, 
    idx[1:N], 
    idx_within_radius[1:N]
  )
  
  one_RDM_fmri %>% mutate(sub = .x) %>% relocate(sub)
  
})
toc()

plan(sequential)
```



# RSA Searchlight

```{r}

rdms_fmri_nested <- RDMs_fmri %>% 
  pivot_longer(cols = !starts_with("sub"), names_to = "region") %>% 
  group_by(sub,region) %>% 
  nest() %>% 
  rename(fmri = data)


rdms_rats_nested <- RDMs_rats %>% 
  group_by(sub) %>% 
  nest %>% 
  rename(rat = data)


# we put a copy of the rating RDM for each atlas region's RDM
joint_fmri_rat_nested <- right_join(rdms_fmri_nested,rdms_rats_nested, by = "sub")
# # the following is just to check that the join went as expected
# ff %>% mutate(sum_rat = rat %>% map_dbl(sum)  )

joint_fmri_rat_nested

# Calculate all similarity
# This is effectively when we are calculating the RSA
all_similarities <- joint_fmri_rat_nested %>% 
  mutate(
    similarity = list(fmri,rat) %>% pmap_dbl(~ cor(.x,.y, method = dist_method_rsa))
  ) %>%
  ungroup

# Produce estimates of RSA acros subjects
RSA <- all_similarities %>%
  select(region, similarity) %>% 
  group_by(region) %>% 
  reframe(
    mean_similarity = round(mean(similarity),2),
    sd_similarity = round(sd(similarity),2)
  ) %>% 
  arrange(desc(mean_similarity))

RSA %>% datatable()
```


# Write the map with the values Searchlight

```{r}

results_to_map <- RSA %>% 
  mutate(region = as.integer(region)) %>% 
  select(region, mean_similarity)

results_path <- "/data00/leonardo/RSA/analyses/rsa/rsa_results"

# add the Searchlight
results_filename <- create_filename_results(
  paste0("Searchlight_",gm_file), ratings_type, dist_method_rating, dist_method_fmri, dist_method_rsa
)


# function to write the results to a nii.gz volume
write_results_to_nii <- function(results_to_map, filename_results_nii, gm) {
  
  results_nii = gm
  
  idx_to_write <- results_to_map$region
  results_nii[idx_to_write] = results_to_map$mean_similarity

  writeNifti(results_nii, filename_results_nii)  
}


# Write the results nii.gz in the rsa_results dir
write_results_to_nii(
  results_to_map,
  paste0(results_path, "/", results_filename),
  gm
)

```









