---
title: "Emotion Ratings Analysis V2"
author: "Leonardo Cerliani"
date: "2024-11-27"
output:
  html_document:
    self_contained: true
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
    theme: cerulean 
---

```{r, message=FALSE}
library(tidyverse)
library(RColorBrewer)

bd <- "/data00/leonardo/RSA/analyses/RATINGS"

subs_file <- "/data00/leonardo/RSA/sub_list.txt"
subs <- sprintf("%02d",readLines(subs_file) %>% as.numeric)

# Load the preprocessed emotion ratings file
# which was prepared with the 02_prepare_ratings_EMOTION.Rmd
# located in /data00/leonardo/RSA/prep_scripts/05_prepare_ratings
df <- read_csv(paste0(bd,"/","emotion_ratings.csv"))

```


# Examine congruence

We need to do the following for each movie, but let's explain the procedure
for one movie only since it's simpler:

- given one movie, for each subject take the mean of the ratings of all other subs
- then correlate the ratings of that sub with the mean of all other sub for that movie
- repeat for each sub
- now apply to each movie

In the end, you will have 56 values (one per movie) for each sub.
Each value shows the congruence (expressed as correlation) of the ratings of that
sub for a specific movie, with respect to the mean rating of all other subs for
that movie.



```{r}
# Let's start with only one movie
df_one_movie <- df %>% 
  select(sub, video, starts_with("r_"), emotion) %>% 
  filter(video == "AK_A_32.mp4")

df_one_movie

```


```{r}
# Filter out the columns with ratings
rating_cols <- df_one_movie %>% select(starts_with("r_")) %>% colnames

# Function to calculate correlation for one subject
calculate_correlation <- function(sub_id, data, rating_cols) {
  # Separate the target subject
  target <- data %>% filter(sub == sub_id) %>% select(starts_with("r_"))
  
  # Calculate the mean ratings excluding the target subject
  others_mean <- data %>%
    filter(sub != sub_id) %>%
    summarise(across(all_of(rating_cols), mean, na.rm = TRUE))
  
  # Calculate the correlation
  corr <- cor(
    unlist(target),
    unlist(others_mean),
    use = "complete.obs"
  )
  
  # Return as a tibble
  tibble(
    sub = sub_id,
    # movie = unique(data$video),
    corr = corr
  )
}


# # Try the function with one sub (and one movie)
# sub_id = "03"
# data = df_one_movie
# calculate_correlation(sub_id = "03", data = df_one_movie, rating_cols)


# Map the function to all subjects
unique(df$sub) %>% map_dfr( ~ calculate_correlation(.x, df_one_movie, rating_cols))


```




```{r}
# Now use the full dataset of ratings with all movies.
# We will therefore need to write two nested loops
#
# NB: in the loop below, x = nested_df for each row of 'data'

congruence <- df %>%
  
  # Remove neutral first!
  filter(emotion != "neutral") %>% 
  
  select(sub, video, starts_with("r_")) %>%
  group_by(video) %>%
  nest() %>%
  mutate(
    correlations = map(data, function(x) {
      # Extract the column names for ratings
      rating_cols <- x %>% select(starts_with("r_")) %>% colnames()
      
      # Iterate over the subjects in the nested data and calculate correlation
      map_dfr(x$sub, function(sub_id) {
        calculate_correlation(sub_id, x, rating_cols)
      })
    })
  ) %>%
  select(video, correlations) %>%
  unnest(correlations) %>% 
  ungroup 
  
  # NB: in some cases the rating was all zeros, thereby leading to no sd and
  # NA in the corr calculation. Replace these with zeros
  congruence <- congruence %>% 
    mutate(corr = ifelse(is.na(corr), 0, corr))


congruence  

```

# Visualization of congruence

```{r}
# Calculate the median correlation for each subject
median_corrs <- congruence %>%
  group_by(sub) %>%
  summarise(median_corr = median(corr, na.rm = TRUE)) %>%
  arrange(median_corr)

# Create the plot
congruence %>%
  ggplot(aes(x = sub, y = corr)) +
  geom_boxplot(outlier.shape = NA) +  # Boxplot for each subject
  geom_jitter(width = 0.2, alpha = 0.6) +  # Dots inside the boxplot
  scale_x_discrete(limits = median_corrs$sub) +  # Order subjects by their median correlation
  geom_line(data = median_corrs, aes(x = sub, y = median_corr, group = 1), color = "red", size = 1.2) +  # Corrected geom_line
  labs(
    title = "Correlation of Subject Ratings with Mean (Leave-One-Out)",
    x = "Subject ID",
    y = "Correlation"
  ) +
  theme_minimal() +  # Use a minimal theme for a cleaner look
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for better readability
    plot.title = element_text(hjust = 0.5)  # Center the title
  )

```





# The correlation is not good in this case. Let's try to use euclidean distance
```{r}
calculate_euclidean_distance <- function(sub_id, data, rating_cols) {
  # Separate the target subject's ratings
  target <- data %>% filter(sub == sub_id) %>% select(starts_with("r_"))
  
  # Calculate the mean ratings excluding the target subject
  others_mean <- data %>%
    filter(sub != sub_id) %>%
    summarise(across(all_of(rating_cols), mean, na.rm = TRUE))
  
  # Calculate the Euclidean distance
  euclidean_dist <- sqrt(sum((unlist(target) - unlist(others_mean))^2))
  
  # Return as a tibble
  tibble(
    sub = sub_id,
    # movie = unique(data$video),  # Uncomment if you want to include the movie column
    euclidean_dist = euclidean_dist
  )
}


congruence_euclidean <- df %>%
  # Remove neutral first!
  filter(emotion != "neutral") %>% 
  
  select(sub, video, starts_with("r_")) %>%
  group_by(video) %>%
  nest() %>%
  mutate(
    correlations = map(data, function(x) {
      # Extract the column names for ratings
      rating_cols <- x %>% select(starts_with("r_")) %>% colnames()
      
      # Iterate over the subjects in the nested data and calculate correlation
      map_dfr(x$sub, function(sub_id) {
        calculate_euclidean_distance(sub_id, x, rating_cols)
      })
    })
  ) %>%
  select(video, correlations) %>%
  unnest(correlations) %>% 
  ungroup 
  

# Viz

# Calculate the median D (euclidean distance) for each subject
median_D <- congruence_euclidean %>%
  group_by(sub) %>%
  summarise(median_D = median(euclidean_dist, na.rm = TRUE)) %>%
  arrange(median_D)

# Create the plot
congruence_euclidean %>%
  ggplot(aes(x = sub, y = euclidean_dist)) +
  # geom_boxplot(outlier.shape = NA) +  # Boxplot for each subject
  geom_jitter(width = 0.2, alpha = 0.6) +  # Dots inside the boxplot
  scale_x_discrete(limits = median_D$sub) +  # Order subjects by their median correlation
  labs(
    title = "Correlation of Subject Ratings with Mean (Leave-One-Out)",
    x = "Subject ID",
    y = "Euclidean Distance"
  ) +
  theme_minimal() +  # Use a minimal theme for a cleaner look
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for better readability
    plot.title = element_text(hjust = 0.5)  # Center the title
  )



# Create the plot
congruence_euclidean %>%
  ggplot(aes(x = sub, y = euclidean_dist)) +
  # Dots inside the plot
  geom_jitter(width = 0.2, alpha = 0.6) +  
  scale_x_discrete(limits = median_D$sub) +  # Order subjects by their median Euclidean distance
  # Red line connecting the median values of each subject
  geom_line(data = median_D, aes(x = sub, y = median_D, group = 1), color = "red", size = 1.2) +
  labs(
    title = "Euclidean Distance of Subject Ratings from Mean (Leave-One-Out)",
    x = "Subject ID (Ordered by Median Euclidean Distance)",
    y = "Euclidean Distance"
  ) +
  theme_minimal() +  # Use a minimal theme for a cleaner look
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for better readability
    plot.title = element_text(hjust = 0.5)  # Center the title
  )


```




```{r}

# Viz with euclidean distances colored by emotion

congruence_euclidean_emotion_name <- left_join(
  congruence_euclidean, 
  df %>% filter(sub == "02") %>%  select(video, emotion),
  by = "video"
)



# Load RColorBrewer for predefined palettes
library(RColorBrewer)

# Create the plot
congruence_euclidean_emotion_name %>%
  ggplot(aes(x = sub, y = euclidean_dist, color = emotion)) +  # Color dots by emotion
  # geom_boxplot(outlier.shape = NA) +  # Boxplot for each subject
  geom_jitter(width = 0.2, alpha = 0.7) +  # Dots inside the boxplot
  scale_x_discrete(limits = median_D$sub) +  # Order subjects by their median Euclidean distance
  labs(
    title = "Euclidean Distance of Subject Ratings from Mean (Leave-One-Out)",
    x = "Subject ID (Ordered by Median Euclidean Distance)",
    y = "Euclidean Distance"
  ) +
  theme_minimal() +  # Use a minimal theme for a cleaner look
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for better readability
    plot.title = element_text(hjust = 0.5)  # Center the title
  ) +
  scale_color_manual(values = brewer.pal(n = 6, name = "Set1"))  # Use the Set11 palette from RColorBrewer





```


# Previous version of congruence : congruence within participant
```{r}

calculate_congruence_initial <- function(df, highlow_flag = "ALL") {
  
  cat("highlow_flag = ", highlow_flag, "\n")

  if (highlow_flag %in% c("high","low")) {
    df <- df %>% filter(str_detect(high_low_code, highlow_flag))
  }
      
  #   -------- IMPORTANT ----------
  # The neutral movies were included in the first version of the congruence
  # calculation, which determines the groun N14. Therefore it is kept.
  df_congruence <- df %>%
    # Remove neutral first!
    # filter(emotion != "neutral") %>%
    select(sub,original_code, emotion, starts_with("r_")) %>%
    pivot_longer(starts_with("r_"), names_to = "rating") %>%
    group_by(sub, original_code, emotion) %>%
    filter(value == max(value)) %>%
    filter(value != 0) %>%
    mutate(rating = str_remove(rating,"r_")) %>%
    mutate(consistent = ifelse(emotion==rating,1,0)) %>%
    ungroup %>% 
    group_by(sub) %>%  # here you could also group by emotion
    summarise(
      congruence = round(sum(consistent)/n(),2)
    )   
  
  return(df_congruence)
}


# calculate congruence for ALL, high, low, using the fn above
congruence_initial <- calculate_congruence_initial(df,"ALL") %>% 
  rename(ALL = congruence) %>% 
  inner_join(calculate_congruence_initial(df,"high"), by="sub") %>% 
  rename(high = congruence) %>% 
  inner_join(calculate_congruence_initial(df,"low"), by="sub") %>% 
  rename(low = congruence)


# OLE plot, suitable only for one of ALL, high, low at the time
# NB: the cut at 0.82 is due to the fact that this is the mean value of congruence
# obtained from the validation data from Rune (check Emotion_Ratings_Analysis_V1.Rmd)
congruence_initial %>%
  ggplot(aes(x = reorder(sub, ALL), y = ALL)) +
  geom_bar(stat = "identity", color = "grey", fill = "lightblue") +
  theme_minimal() +
  labs(
    title = "Congruence between displayed emotion and max rated emotion",
    subtitle = paste0("mean : ", congruence_initial$ALL %>% mean %>% round(2)),
    x = "sub"
  ) +
  # geom_hline(yintercept = congruence_initial$ALL %>% mean, linetype = "dashed", color = "blue")
  geom_hline(yintercept = 0.82, linetype = "dashed", color = "blue", size = 1)


congruence_initial %>%
  pivot_longer(cols = c("ALL", "high", "low"), names_to = "intensity", values_to = "value") %>%
  ggplot(aes(x = reorder(sub, value), y = value, fill = intensity)) +
  geom_bar(stat = "identity", position = "dodge") + 
  theme_minimal() +
  theme(legend.position = "top") +
  labs(x = "Subject", y = "Value", title = "Congruence across emotion intensity levels")


# IMPORTANT!!! THE SLIGHT DISCREPANCY WRT TO THE PREVIOUS PLOT IS DUE TO THE FACT
# THAT I DO NOT CONSIDER CONGRUENCE IN NEUTRAL MOVIES

congruence_initial %>% 
  pivot_longer(
    cols = c("ALL","high","low"), names_to = "intensity", values_to = "congruence"
  ) %>% 
  ggstatsplot::ggwithinstats(
    x = intensity,
    y = congruence,
    type = "parametric"
  )

```

```{r}
congruence_threshold <- congruence_initial$ALL %>% mean %>% round(2)

# tmp <- df_congruence %>%
#   count(is_congruent = ALL >= congruence_threshold)

n_above <- which(congruence_initial$ALL >= congruence_threshold) %>% length
n_below <- length(congruence_initial$sub) - n_above

df %>%
  select(sub, starts_with("r_"), emotion) %>% 
  inner_join(congruence_initial %>% select(sub, ALL), by = "sub") %>% 
  mutate(is_congruent = ifelse(ALL >= congruence_threshold, "ABOVE", "BELOW")) %>%
  mutate(is_congruent = factor(is_congruent)) %>% 
  select(-ALL) %>% 
  group_by(sub, emotion, is_congruent) %>%
  reframe(
    across(starts_with("r_"), ~ round(mean(.),2))
  ) %>% 
  pivot_longer(cols = starts_with("r_"), names_to = "rating_type", values_to = "rating") %>%
  ggplot(aes(x = rating_type, y = rating, color = is_congruent, group = sub)) +
  geom_line(linewidth = 0.5, alpha = 0.7) +
  geom_point(color = "grey") +
  scale_y_continuous(limits = c(0, 10), breaks = seq(0, 10, by = 2)) +  # Setting y-axis limits and breaks
  facet_grid(emotion ~ is_congruent, axes = "all", axis.labels = "all") +  
  theme_minimal() +
  labs(
    title = paste0("Ratings for participants above or below congruence of ", congruence_threshold),
    subtitle = paste0("n_above: ", n_above, " n_below: ", n_below)
  )
```




# Compare initial and actual calculation 
To assess whether the participants with high congruence are the same

```{r}
# Define the median values
median_cong_OLD <- congruence_initial$ALL %>% mean
median_cong_NEW <- median_D$median_D %>% mean

# Create the plot
congruence_initial %>% 
  select(sub, ALL) %>%
  inner_join(median_D, by = "sub") %>% 
  rename(cong_ole = ALL, cong_new = median_D) %>% 
  ggplot(aes(x = cong_ole, y = cong_new)) +
  geom_point() +  # Scatter plot
  geom_text(aes(label = sub), vjust = 1.5, hjust = 0.5) +  # Text labels
  # Add dotted vertical line at median_cong_OLD
  geom_vline(xintercept = median_cong_OLD, linetype = "dashed", color = "red") +
  # Add dotted horizontal line at median_cong_NEW
  geom_hline(yintercept = median_cong_NEW, linetype = "dashed", color = "blue") +
  theme_minimal() +
  labs(
    x = "Old Congruence (high=good)", 
    y = "New Congruence (high=bad)", 
    title = "Scatter Plot with Median Lines"
  )

```


# Alternatively, plot only the sub number. It is clear enough and avoids clutter
```{r}
# Define the median values
median_cong_OLD <- congruence_initial$ALL %>% mean
median_cong_NEW <- median_D$median_D %>% mean

# Create the plot
congruence_initial %>% 
  select(sub, ALL) %>%
  inner_join(median_D, by = "sub") %>% 
  rename(cong_ole = ALL, cong_new = median_D) %>% 
  ggplot(aes(x = cong_ole, y = cong_new)) +
  # geom_point() +  # Scatter plot
  geom_text(aes(label = sub), vjust = 0, hjust = 0) +  # Text labels
  # Add dotted vertical line at median_cong_OLD
  geom_vline(xintercept = median_cong_OLD, linetype = "dashed", color = "red") +
  # Add dotted horizontal line at median_cong_NEW
  geom_hline(yintercept = median_cong_NEW, linetype = "dashed", color = "blue") +
  theme_minimal() +
  labs(
    x = "Old Congruence (high=good)", 
    y = "New Congruence (high=bad)", 
    title = "Scatter Plot with Median Lines"
  )

cat(paste0(
    "Congruence within: mean : ", congruence_initial$ALL %>% mean %>% round(2),
    " sd : ", congruence_initial$ALL %>% sd %>% round(2)
  ))

cat("\n")
cat(paste0(
    "Congruence between: mean : ", median_D$median_D %>% mean %>% round(2),
    " sd : ", median_D$median_D %>% sd %>% round(2)
  ))



```


# Compare congruence_initial with ISC
```{r}
# Define the median values
median_cong_OLD <- congruence_initial$ALL %>% mean
median_cong_NEW <- median_corrs$median_corr %>% mean

# Create the plot
congruence_initial %>% 
  select(sub, ALL) %>%
  # filter(sub == 11) %>% # NB: 11 overlaps with 12 in the plot
  inner_join(median_corrs, by = "sub") %>% 
  rename(cong_ole = ALL, cong_new = median_corr) %>% 
  ggplot(aes(x = cong_ole, y = cong_new)) +
  geom_point() +  # Scatter plot
  # geom_point() +  # Scatter plot
  geom_text(aes(label = sub), vjust = 0, hjust = 0) +  # Text labels
  # Add dotted vertical line at median_cong_OLD
  geom_vline(xintercept = median_cong_OLD, linetype = "dashed", color = "red") +
  # Add dotted horizontal line at median_cong_NEW
  geom_hline(yintercept = median_cong_NEW, linetype = "dashed", color = "blue") +
  theme_minimal() +
  labs(
    x = "Congruence within subjects : high=good", 
    y = "Congruence between subjects R : high=good", 
    title = "Scatter Plot with Median Lines"
  )

cat(paste0(
    "Congruence within: mean : ", congruence_initial$ALL %>% mean %>% round(2),
    " sd : ", congruence_initial$ALL %>% sd %>% round(2)
  ))

cat("\n")
cat(paste0(
    "Congruence between: mean : ", median_D$median_D %>% mean %>% round(2),
    " sd : ", median_D$median_D %>% sd %>% round(2)
  ))

```

























