---
title: "Generate design.mat"
author: "LC"
date: "2024-11-29"
output: html_document
---



```{r, message=FALSE}
library(tidyverse)

# NB: remember to change the number of ALLMOVIES cope (to be removed)
#     according to the model! For emotion, it is cope 7

model="emotion"
bd <- paste0("/data00/leonardo/RSA/analyses/parametric_modulation/",model)

F_results_dir <- paste0(bd,"/results_F_tests")
ifelse(!dir.exists(F_results_dir),dir.create(F_results_dir),"Directory Exists")

subs_file <- "/data00/leonardo/RSA/sub_list.txt"
subs <- sprintf("%02d",readLines(subs_file) %>% as.numeric)
```


# Generate df with file_path of the cope1.nii.gz images
```{r}

bd_2nd_level = paste0(bd,"/results/2nd_level")

df <- list.files(
  bd_2nd_level, 
  pattern = "^cope1.nii.gz", recursive = T
) %>% 
  
  # read all the *cope1.nii.gz in a tibble
  as_tibble() %>% 
  rename(file_path = value) %>% 
  
  # full path of file_path
  mutate(file_path = paste0(bd_2nd_level,"/",file_path)) %>%
  
  # add columns for sub and for cope#
  mutate(
    sub = str_extract(file_path, paste(subs, collapse = "|")),
    cope = str_extract(file_path, "cope\\d+\\.feat") %>% str_remove_all("[^\\d]")
  ) %>% 
  
  # REMOVE ALLMOVIES COPE (number 7)
  filter(cope != 7)
  

df
  
```

# Write the list of files to pass to the Feat GUI
```{r}

writeLines(
  df$file_path,
  con = paste0(F_results_dir,"/list_cope_nii_gz.txt")
)

```




# Generate design.mat
```{r}

n_subs = length(subs)
n_copes = length(unique(df$cope))

# create design.mat in a ff dataframe
ff <- df %>% select(sub,cope)

for (i in 1:n_copes) {
  ev_column <- paste0("EV", i)  # Create the name for the EV column (e.g., EV1, EV2, etc.)
  
  # Add the column where the cope number equals i will be 1, others will be 0
  ff <- ff %>%
    mutate(!!ev_column := ifelse(cope == i, 1, 0))
}

ff
```



# Viz the design matrix
```{r}

ff_matrix <- ff %>%
  # arrange(cope) %>%
  mutate(across(everything(), as.numeric)) %>%
  mutate(row_label = paste0("s",sub,"_cope",cope)) %>% 
  select(-sub) %>% 
  column_to_rownames("row_label") %>% 
  as.matrix()

# Plot the heatmap (after reversing the order of the rows)
heatmap(ff_matrix[nrow(ff_matrix):1, ], Rowv = NA, Colv = NA)

heatmap(ff_matrix[nrow(ff_matrix):1, 2:7], Rowv = NA, Colv = NA)
```


# Write design.mat

Header first:

```
/NumWaves       n_copes
/NumPoints      dim(df)[1]
/PPheights      1.000000 1.000000 ... [n_copes]
/Matrix
```

```{r}

ff2write <- ff %>% 
  mutate(row_label = paste0("s",sub,"_cope",cope)) %>% 
  relocate(row_label) %>% 
  select(!c(sub,cope)) %>% 
  column_to_rownames("row_label") %>% 
  mutate(across(everything(), as.numeric)) %>%
  as.matrix

ff2write

# Write the header lines
num_waves <- n_copes  # Set number of copes
num_points <- dim(ff)[1]  # Set number of subjects (rows)

header_lines_design_mat <- c(
  paste("/NumWaves", num_waves),
  paste("/NumPoints", num_points),
  paste("/PPheights", paste0(rep("1.000000",num_waves), collapse = " ")),
  paste(" "),
  "/Matrix"
)

design_mat_location <- paste0(F_results_dir,"/design.mat")
writeLines(header_lines_design_mat, design_mat_location)

# Append the matrix to the file
write.table(
  ff2write, 
  design_mat_location, 
  append = TRUE, 
  sep = " ", 
  col.names = FALSE, 
  row.names = FALSE, 
  quote = FALSE
)


```


# Write the design.grp
```{r}

num_waves <- 1
num_points = dim(ff)[1]

design_grp = c(
  paste0("/NumWaves ", num_waves),
  paste0("/NumPoints ", num_points),
  "/Matrix",
  rep("1.000000000000", num_points)
)

design_grp_location <- paste0(F_results_dir,"/design.grp")

write.table(
  design_grp, 
  design_grp_location, 
  append = FALSE, 
  sep = " ", 
  col.names = FALSE, 
  row.names = FALSE, 
  quote = FALSE
)

```



# Manually write the other matrices

## design.con
```
/NumWaves       6
/NumContrasts   11
/PPheights  1.000000e+00    1.000000e+00    1.000000e+00    1.000000e+00    1.000000e+00    1.000000e+00 
/RequiredEffect		1000	1000    1000    1000    1000    1000

/Matrix
1 0 0 0 0 0
0 1 0 0 0 0
0 0 1 0 0 0
0 0 0 1 0 0
0 0 0 0 1 0
0 0 0 0 0 1
1 -1 0 0 0 0
0 1 -1 0 0 0
0 0 1 -1 0 0
0 0 0 1 -1 0
0 0 0 0 1 -1
```

## design.fcon
```
/NumWaves       11
/NumContrasts   2

/Matrix
1 1 1 1 1 1 0 0 0 0 0
0 0 0 0 0 0 1 1 1 1 1
```



https://bookdown.org/yihui/rmarkdown-cookbook/eng-bash.html

```{bash, engine.opts='-l'}
model=emotion

sub_list=/data00/leonardo/RSA/sub_list.txt

dest=/data00/leonardo/RSA/analyses/parametric_modulation/${model}/results_F_tests
orig=/data00/leonardo/RSA/analyses/parametric_modulation/${model}/results/2nd_level

# Create the merge *within* sub across copes first

# Read each subject number from the sub_list.txt file
while read -r sub; do

    subpad=$(printf "%02d" ${sub})  # Zero-pad the subject number
    echo $subpad

    # fslmerge copes and write list_sub_copes
    fslmerge -t  ${dest}/sub-${subpad}_copes \
        ${orig}/sub-${subpad}_${model}.gfeat/cope{1..6}.feat/stats/cope1.nii.gz

    # fslmerge varcopes and write list_sub_copes
    fslmerge -t  ${dest}/sub-${subpad}_varcopes \
        ${orig}/sub-${subpad}_${model}.gfeat/cope{1..6}.feat/stats/varcope1.nii.gz


done < "$sub_list"


echo "merging copes across subs..."
fslmerge -t ${dest}/copes $(ls ${dest}/*_copes.nii.gz)

echo "merging varcopes across subs..."
fslmerge -t ${dest}/varcopes $(ls ${dest}/*_varcopes.nii.gz)

rm ${dest}/sub*
```



# Also make sure you have the additional required files

- `mask.nii.gz` (e.g. MNI mask)
- `tdof_filtered_func_data.nii.gz` = (num_points - 1) 4D of mask.nii.gz 

```{bash, engine.opts='-l'}

model=emotion

dest=/data00/leonardo/RSA/analyses/parametric_modulation/${model}/results_F_tests

mask_nii_path="${dest}/mask.nii.gz"

tdof_nii_path="${dest}/tdof_filtered_func_data.nii.gz"

num_points=$(fslinfo ${dest}/copes.nii.gz | grep ^dim4 | awk '{print $2}')
tdof=$((${num_points} - 1))

time fslmerge -t ${tdof_nii_path} $(for i in $(seq 1 ${tdof}); do echo ${mask_nii_path}; done)

```

# Finally run flameo
```{bash, engine.opts='-l'}

model="emotion"

bd="/data00/leonardo/RSA/analyses/parametric_modulation/${model}/results_F_tests"
cd ${bd}

dest=${bd}/flameo_output

[ -d ${dest} ] && rm -rf ${dest}
# mkdir ${dest}

# fe (fixed effects), ols (mixed effects - OLS), 
# flame1 (mixed effects - FLAME stage 1), flame12 (mixed effects - FLAME stage 1+2)


flameo --cope=copes.nii.gz \
       --vc=varcopes.nii.gz \
       --dvc=tdof_filtered_func_data \
       --mask=mask.nii.gz \
       --ld=${dest} \
       --dm=design.mat \
       --tc=design.con \
       --cs=design.grp \
       --fc=design.fcon \
       --runmode=ols
      
```


# Thresholding
Do it manually. Sample code here:

```
fsl-cluster -i zfstat1 \
	-t 3.1 \
	--othresh=thresh_zstat1 \
	--connectivity=26 \
	--mm \
	-p 0.05 \
	-d 0.06 \
	--volume=213207
	```















